---
title: "Chapter 12: Model Assessment and Predictive Model Checking"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Chapter 12: Model Assessment and Predictive Model Checking}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6
)
knitr::opts_knit$set(global.par = TRUE)
pdfplots <- FALSE # default: FALSE; set this to TRUE only if you like pdf figures
seed <- 42
```

# Section 12.1: Beyond classical Bayesian model comparison

## Section 12.1.1: Schwarz Criterion and BIC

### Example 12.1. Labor Market Data: Bayesian Variable Selection using BIC

We return to the probit analysis of Example 11.11 and run variable selection 
using the Schwarz criterion instead of the marginal likelihood. 
We first implement this approach.

```{r}
modsel_probit_BIC <- function(y,X,burnin = 1000L, M = 5000L){
  n <- dim(X)[1]
  p <- dim(X)[2] # change to X later

  gamma_post <- matrix(ncol = p, nrow = M)
  acc <- numeric(length = M)

  gamma <- matrix(rep(1,p),nrow=1)
  mod0<-glm(y~X, family=binomial(link = "probit"))
  SC_old <- logLik(mod0) - log(n)*(p+1)/2

  for (m in seq_len(burnin + M)) {
    gamma_proposed <- gamma_old <- gamma

    j <- sample(1:p, size=1)
    gamma_proposed[j] <- 1-gamma_old[j]

    X_gamma <- X[,gamma_proposed==1]
    model<-glm(y~X_gamma, family=binomial(link = "probit"))
    SC_proposed <- logLik(model) - log(n)*(sum(gamma_proposed)+1)/2
    
    # compute acceptance probability and decide on acceptance
    log_acc <- SC_proposed - SC_old

    if (log(runif(1)) < log_acc) {
      gamma <- gamma_proposed
      SC_old <- SC_proposed
      accept <- 1
    } else {
      gamma <- gamma_old
      accept <- 0
    }

    if (m > burnin) {
      gamma_post[m-burnin, ] <- gamma
      acc[m-burnin] <- accept
    }
  }
  return(list(gamma_post=gamma_post,acc=acc))
}
```

We next load the data and as in exercise 11.11. add 5 irrelevant, standard 
normal regressors. Then we run the algorithm. 

```{r}
library("BayesianLearningCode")
data("labor", package = "BayesianLearningCode")
y <- labor$income_1998 == "zero"
N <- length(y)
X_unemp <- with(labor, cbind(female = female,
                             age18 = 1998 - birthyear - 18,
                             wcollar = wcollar_1997,
                             unemp97 = income_1997 == "zero")) # regressor matrix
p_irrel <- 5
set.seed(seed)
X <- cbind(X_unemp,matrix(rnorm(p_irrel*N),ncol=p_irrel))
colnames(X)[4+(1:p_irrel)]<-c("irrel1","irrel2","irrel3","irrel4","irrel5")

p<-dim(X)[2]
M <-20000 
res <- modsel_probit_BIC(y, X, M=M)
```

We determine the number of models visited during MCMC, the average model 
complexity, i.e. the average number of covariates included in the model and show 
the posterior inclusion probabilities.
```{r}
if (pdfplots) {
  pdf("12-1_1.pdf", width = 5, height = 6)
}

gammas_unique <- unique(res$gamma_post)
num_mod <- dim(gammas_unique)[1]
print(num_mod)

k_gamma=rowSums(res$gamma_post)
print(mean(k_gamma))

# PIPs
print(colMeans(res$gamma_post))

par(mfrow = c(1, 1), mar = c(2.5, 2.5, 1.5, .5), mgp = c(1.5, .5, 0))
barplot(colMeans(res$gamma_post), col="blue",names.arg=1:p,
        xlab="Covariate",ylab="PIP")
```
To determine the model visited most often we re-use a function which we defined
in Chapter 11.
```{r}
number_draws<- function(gamma_post, models){
  
  nmod <- dim(models)[1]
  freq <- rep(NA,nmod)
  
  for (j in (1:nmod)){
    freq[j] <- sum(apply(gamma_post, 1,
                         function(x) identical(x, models[j,])))
  }
  return(freq)
}
freq_gammas<- number_draws(res$gamma_post,gammas_unique)
io <- order(freq_gammas, decreasing=TRUE)

knitr::kable(cbind(gammas_unique,freq_gammas/M)[io[1:5],],
             digits=cbind(rep(0, p),4))
```
