---
title: "Chapter 10: Bayesian Model Selection"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Chapter 10: Bayesian Model Selection}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6
)
knitr::opts_knit$set(global.par = TRUE)
pdfplots <- FALSE # default: FALSE; set this to TRUE only if you like pdf figures
```

# Section 10.1: The Foundations of Bayesian Model Selection
## Table 10.1: (Log) Bayes factor and model probabilities

```{r}
logBF <- -7:7
BF <- exp(logBF)
PrM1 <- BF / (1 + BF)
PrM2 <- 1 - PrM1
knitr::kable(cbind(BF, logBF, PrM1, PrM2))
```

## Example 10.5: CHF exchange rate data - Testing the null hypothesis $\mu = 0$

Before producing Figure 10.1, we begin with a concrete example.
We load the data, compute the required parameters, and evaluate the two 
marginal likelihoods.

```{r}
data("exrates", package = "stochvol")
y <- 100 * diff(log(exrates$USD / exrates$CHF))

c0 <- 1
C0 <- 1
N0 <- 10^(-5:3)

N <- length(y)
cN <- c0 + N  / 2
CN_M1 <- C0 + sum(y^2) / 2
CN_M2 <- C0 + 0.5 * N * (mean((y - mean(y))^2) + N0 / (N0 + N) * mean(y)^2)

(logmarglikM1 <- lgamma(cN) + c0 * log(C0) -
   lgamma(c0) - cN * log(CN_M1) - 0.5 * N * log(2 * pi))
(logmarglikM2 <- lgamma(cN) + c0 * log(C0) + 0.5 * log(N0) -
   lgamma(c0) - cN * log(CN_M2) - 0.5 * N * log(2 * pi) - 0.5 * log(N0 + N))
```

The "direct" formula for the log Bayes factor is even simpler, and we can check
whether we get the same answer.

```{r}
(logBF <- 0.5 * (log(N0 + N) - log(N0)) + cN * (log(CN_M2) - log(CN_M1)))
all.equal(logBF, logmarglikM1 - logmarglikM2)
```

Now we are ready to investigate the sensitivity of the log Bayes factor with
respect to prior hyperparameter choices.

```{r, echo = -c(1:2)}
if (pdfplots) {
  pdf("10-1_1.pdf", width = 8, height = 4)
}
par(mar = c(2.5, 2.5, 1.5, .5), mgp = c(1.6, .6, 0))
c0 <- c(-1, 0, 0.001, 1, 100)
C0 <- c(0, 0, 0.001, 1, 100)
N0 <- 10^seq(1, 10, by = 0.1)

logBF <- matrix(NA_real_, nrow = length(N0), ncol = length(c0))
for (i in seq_along(c0)) {
  cN <- c0[i] + N  / 2
  CN_M1 <- C0[i] + sum(y^2) / 2
  CN_M2 <- C0[i] + 0.5 * N * (mean((y - mean(y))^2) + N0 / (N0 + N) * mean(y)^2)
  logBF[, i] <- 0.5 * (log(N0 + N) - log(N0)) + cN * (log(CN_M2) - log(CN_M1))
}
matplot(N0, logBF, log = "x", type = "l",
        xlab = expression(paste(N[0], " [log scale]")), ylab = "log BF")
abline(h = 0, lty = 3)
legend("topright", paste0("IG(", c0, ",", C0, ")"), col = seq_along(c0),
       lty = seq_along(c0))
title("Log Bayes factor in favor of the zero-mean model")
```

# Section 10.2: Bayesian Testing of Hypotheses
## Example 10.6: Labor market data - Testing for heterogeneity of the no-income risk

First, we re-load the data from Chapter 3.

```{r}
data("labor", package = "BayesianLearningCode")
labor <- subset(labor,
                income_1997 != "zero" & female,
                c(income_1998, wcollar_1986))
labor <- with(labor,
              data.frame(unemployed = income_1998 == "zero",
                         wcollar = wcollar_1986))
```

Next, we compute the marginal likelihoods for the homogeneity model.

```{r}
N <- length(labor$unemployed)
SN <- sum(labor$unemployed)
hN <- SN / N
N0 <- c(2, 10)
a0 <- c(1, 0.5, hN * N0) 
b0 <- c(1, 0.5, N0 * (1 - hN))

aN <- a0 + sum(labor$unemployed)
bN <- b0 + N - sum(labor$unemployed)

logmarglikM1 <- lbeta(aN, bN) - lbeta(a0, b0)
```

And for the heterogeneity model.

```{r}
N1 <- with(labor, sum(wcollar))
SN1 <- with(labor, sum(wcollar & unemployed))
N2 <- with(labor, sum(!wcollar))
SN2 <- with(labor, sum(!wcollar & unemployed))
hN <- (SN1 + SN2) / (N1 + N2) # redundant, same as above
a01 <- a02 <- c(1, 0.5, hN * N0) # redundant, same as a0
b01 <- b02 <- c(1, 0.5, N0 * (1 - hN)) # redundant, same as b0

aN1 <- a01 + SN1
bN1 <- b01 + N1 - SN1
aN2 <- a02 + SN2
bN2 <- b02 + N2 - SN2

logmarglikM2 <- lbeta(aN1, bN1) + lbeta(aN2, bN2) -
   lbeta(a01, b01) - lbeta(a02, b02)
```

We can now compute the model probabilities and the log BFs.

```{r}
logBF <- logmarglikM1 - logmarglikM2
PrM2 <- 0.5 / (0.5 + 0.5 * exp(logBF))
PrM1 <- 1 - PrM2
#print(xtable::xtable(res, digits = c(0, 2, 2, 5, 5, 5), row.names = FALSE))
knitr::kable(res <- cbind(logmarglikM1, logmarglikM2, logBF, PrM1, PrM2))
```

## Example 10.7: Stomach cancer data - Testing for heterogeneity of the mortality rate

We proceed exactly as above, just with different data.

```{r}
N1  <- 1668 # number at risk in City A
SN1 <- 2    # cancer deaths in City A
N2  <- 583  # number at risk in City B
SN2 <- 1    # cancer deaths in City B
```

We again begin with the homogeneity model.

```{r}
N <- N1 + N2
SN <- SN1 + SN2
hN <- SN / N
N0 <- c(2, 10)
a0 <- c(1, 0.5, hN * N0) 
b0 <- c(1, 0.5, N0 * (1 - hN))

aN <- a0 + SN
bN <- b0 + N - SN

logmarglikM1 <- lbeta(aN, bN) - lbeta(a0, b0)
```

Followed by the heterogeneity model.

```{r}
a01 <- a02 <- a0
b01 <- b02 <- b0

aN1 <- a01 + SN1
bN1 <- b01 + N1 - SN1
aN2 <- a02 + SN2
bN2 <- b02 + N2 - SN2

logmarglikM2 <- lbeta(aN1, bN1) + lbeta(aN2, bN2) -
   lbeta(a01, b01) - lbeta(a02, b02)
```

We can now compute the model probabilities and the log BFs.

```{r}
logBF <- logmarglikM1 - logmarglikM2
PrM2 <- 0.5 / (0.5 + 0.5 * exp(logBF))
PrM1 <- 1 - PrM2
#print(xtable::xtable(res, digits = c(0, 2, 2, 3, 3, 3), row.names = FALSE))
knitr::kable(res <- cbind(logmarglikM1, logmarglikM2, logBF, PrM1, PrM2))
```

## Example 10.8: Testing for a structural break in the road safety data

We load the data.

```{r}
data("accidents", package = "BayesianLearningCode")
```

We define the priors hyperparameters and compute the log marginal likelihood
for the exchangable model.

```{r}
y <- accidents[, "children_accidents"]
a0_tmp <- c(0.01, 0.1, 0.5, 1, 2)
m0_tmp <- c(1, mean(y), 5, 7, 10)
grid <- expand.grid(a0 = a0_tmp, m0 = m0_tmp)
a0 <- grid$a0
b0 <- grid$a0 / grid$m0
aN <- sum(y) + a0
bN <- length(y) + b0
logmarglikM1 <- matrix(a0 * log(b0) + lgamma(aN) -
                       aN * log(bN) - lgamma(a0) -
                       sum(lgamma(y + 1)),
                       nrow = length(a0_tmp), ncol = length(m0_tmp),
                       dimnames = list(a0 = a0_tmp, m0 = round(m0_tmp, 2)))
knitr::kable(round(logmarglikM1, 2))
```

And now the same for the model with structural break.

```{r}
accidents1 <- window(accidents, end = c(1994, 9))
accidents2 <- window(accidents, start = c(1994, 10))

a01 <- a02 <- a0
b01 <- b02 <- b0

aN1 <- a01 + sum(accidents1[, "children_accidents"])
aN2 <- a02 + sum(accidents2[, "children_accidents"])
bN1 <- b01 + length(accidents1[, "children_accidents"])
bN2 <- b02 + length(accidents2[, "children_accidents"])

logmarglikM2 <- matrix(a01 * log(b01) + lgamma(aN1) +
                       a02 * log(b02) + lgamma(aN2) -
                       aN1 * log(bN1) - lgamma(a01) -
                       aN2 * log(bN2) - lgamma(a02) -
                       sum(lgamma(y + 1)),
                       nrow = length(a0_tmp), ncol = length(m0_tmp),
                       dimnames = list(a0 = a0_tmp, m0 = round(m0_tmp, 2)))
knitr::kable(round(logmarglikM2, 2))
```

We can now compute log Bayes factors and corresponding model probabilities
(under uniform prior probabilities).

```{r}
logBF <- logmarglikM1[, 2] - logmarglikM2[, 2]
PrM2 <- 0.5 / (0.5 + 0.5 * exp(logBF))
PrM1 <- 1 - PrM2
knitr::kable(round(rbind(PrM1, PrM2), 3))
```

## Example 10.9: CHF exchange rate data - Testing normal versus Student t

After loading the data, we define the degrees of freedom $\nu$ and the prior
hyperparameters.

```{r}
y <- 100 * diff(log(exrates$USD / exrates$CHF))
N <- length(y)
nu <- 7
c10 <- c30 <- 3
C10 <- 3
C30 <- (nu - 2) / nu * C10
```

We now compute log marginal likelihoods. This is straightforward for the normal
model.

```{r}
c1N <- c10 + N / 2
C1N <- C10 + sum(y^2) / 2

logmarglikM1 <- lgamma(c1N) + c10 * log(C10) -
   lgamma(c10) - c1N * log(C1N) - 0.5 * N * log(2 * pi)
```

For the Student $t$ model, we need, e.g., numerical integration. Note that we
apply the "log-sum-exp" trick here to normalize the integrand so that its
maximum is 1.

```{r}
integrand_nonvec <- function(sigma2, y, c0, C0, nu, const = 0, log = FALSE) {
  N <- length(y)
  logint <- -(N/2 + c0 + 1) * log(sigma2) -
            0.5 * (nu + 1) * sum(log(1 + y^2 / (nu * sigma2))) -
            C0 / sigma2
  if (log) logint + const else exp(logint + const)
}
integrand <- Vectorize(integrand_nonvec, "sigma2")

resolution <- 1000
grid <- seq(0.1, 0.7, length.out = resolution + 1)

logtmp <- integrand(grid, y, c30, C30, nu, log = TRUE)
const <- -max(logtmp)
tmp <- exp(logtmp + const)
plot(grid, tmp, type = 'l')
logarea <- log(sum(diff(grid) * .5 * (head(tmp, -1) + tail(tmp, -1)))) - const

logmarglikM3 <- c30 * log(C30) +
                N * lgamma((nu + 1) / 2) -
                lgamma(c30) -
                N * lgamma(nu / 2) -
                N / 2 * log(nu * pi) +
                logarea
```

We can now compute log Bayes factors and corresponding model probabilities
(under equal prior model probabilities).

```{r}
logBF <- logmarglikM1 - logmarglikM3
PrM3 <- 0.5 / (0.5 + 0.5 * exp(logBF))
PrM1 <- 1 - PrM3
knitr::kable(cbind(logML1 = logmarglikM1, logML3 = logmarglikM3,
                   BF = exp(logBF), logBF = logBF, Pr1 = PrM1, Pr3 = PrM3))
```

## Example 10.10: Eye tracking data - Testing homogeneity against unobserved heterogeneity

We begin by loading the data and specifying the hyperparameters.

```{r}
data("eyetracking", package = "BayesianLearningCode")
y <- eyetracking$anomalies
N <- length(y)

a0_tmp <- c(0.1, 0.5, 1, 2)
m0_tmp <- c(1, mean(y), 5, 10, 20)
grid <- expand.grid(a0 = a0_tmp, m0 = m0_tmp)
a0 <- grid$a0
m0 <- grid$m0
b0 <- a0 / m0
```

Now we can compute and print the log marginal likelihoods.

```{r}
logmarglikM1 <- a0 * log(b0) + lgamma(a0 + sum(y)) -
                lgamma(a0) - (a0 + sum(y)) * log(b0 + N) -
                sum(lgamma(y + 1))

logmarglikM2 <- rep(NA_real_, length(a0))
for (i in seq_along(a0)) {
  logmarglikM2[i] <- N * a0[i] * log(b0[i]) + sum(lgamma(a0[i] + y)) -
                     N * lgamma(a0[i]) - sum(a0[i] + y) * log(b0[i] + 1) -
                     sum(lgamma(y + 1))
}

resM1 <- matrix(logmarglikM1, nrow = length(a0_tmp), ncol = length(m0_tmp),
                dimnames = list(a0 = a0_tmp, m0 = round(m0_tmp, 2)))
resM2 <- matrix(logmarglikM2, nrow = length(a0_tmp), ncol = length(m0_tmp),
                dimnames = list(a0 = a0_tmp, m0 = round(m0_tmp, 2)))

knitr::kable(resM1, digits = 2)
knitr::kable(resM2, digits = 2)
```

## Example 10.11: Eye tracking data - Testing Poisson vs. negative binomial

In the negative binomial case, if $a_0$ is fixed and $b_0$ follows a beta prime
prior, we have a closed-form expression for the marginal likelihood.

```{r}
alpha_b0 <- rep(6, length(a0))
beta_b0 <- (alpha_b0 - 1) * m0 / a0
sdb0 <- sqrt(alpha_b0 * (alpha_b0 + beta_b0 - 1) /
             ((beta_b0 - 2) * (beta_b0 - 1)^2))
alpha_bN <- alpha_b0 + N * a0
beta_bN <- beta_b0 + sum(y)

logmarglikM3 <- rep(NA_real_, length(a0))
for (i in seq_along(a0)) {
  logmarglikM3[i] <- lbeta(alpha_bN[i], beta_bN[i]) - lbeta(alpha_b0[i], beta_b0[i]) -
                     N * lgamma(a0[i]) + sum(lgamma(a0[i] + y)) - sum(lgamma(y + 1))
}

resM3 <- matrix(logmarglikM3, nrow = length(a0_tmp), ncol = length(m0_tmp),
                dimnames = list(a0 = a0_tmp, m0 = round(m0_tmp, 2)))

knitr::kable(resM3, digits = 2)
```


## Example 10.12: CHF exchange rate data - Savage-Dickey density ratio 

```{r}
library("BayesianLearningCode")
y <- 100 * diff(log(exrates$USD / exrates$CHF))

c0R <- 1
c0U <- c0R - 1/2 # This is needed for the SD theorem to be valid!
C0 <- 1
N0s <- N0 <- 10^(seq(2.5, 7, by = .1))

N <- length(y)
bNs <- sum(y) / (N0s + N)
cNR <- c0R + N / 2
cNU <- c0U + N / 2
CNs <- C0 + 0.5 * sum((y - mean(y))^2) + 0.5 * N * N0s * mean(y)^2 / (N0s + N)

(logSD <- dstudt(0, bNs, sqrt(CNs / (cNU * (N0s + N))), df = 2 * cNU, log = TRUE) -
          dstudt(0, 0, sqrt(C0 / (c0U * N0s)), df = 2 * c0U, log = TRUE))

plot(N0s, logSD, type = "l", log = "x")
```

Let us double-check this:

```{r}
CN_M1 <- C0 + sum(y^2) / 2

logmarglikM1 <- lgamma(cNR) + c0R * log(C0) -
   lgamma(c0R) - cNR * log(CN_M1) - 0.5 * N * log(2 * pi)
logmarglikM2 <- lgamma(cNU) + c0U * log(C0) + 0.5 * log(N0) -
   lgamma(c0U) - cNU * log(CNs) - 0.5 * N * log(2 * pi) - 0.5 * log(N0 + N)

(logBF <- logmarglikM1 - logmarglikM2)
all.equal(logSD, logBF)
```

Here is a visualization.

```{r, echo = -1}
if (pdfplots) {
  pdf("10-1_2.pdf", width = 8, height = 4)
}
par(mar = c(2.5, 1.5, .5, .5), mgp = c(1.6, .6, 0))
mus <- seq(-.01, .02, 0.0001)
plot(NULL, xlim = range(mus), log = "", xlab = expression(mu), ylab = "",
     ylim = range(dstudt(mus,
                         bNs[length(N0s)],
                         sqrt(CNs[length(N0s)] / (cNU * (N0s[length(N0s)] + N))),
                         df = 2 * cNU)))
abline(v = 0, lty = 3)
abline(h = 0, lty = 3)
for (i in seq_along(N0s)) {
  lines(mus, dstudt(mus, 0, sqrt(C0 / (c0R * N0s[i])), df = 2 * c0R),
        lty = 2, col = i)
  lines(mus, dstudt(mus, bNs[i], sqrt(CNs[i] / (cNU * (N0s[i] + N))), df = 2 * cNU),
        lty = 1, col = i)
  points(c(0, 0),
         c(dstudt(0, 0, sqrt(C0 / (c0R * N0s[i])), df = 2 * c0R),
           dstudt(0, bNs[i], sqrt(CNs[i] / (cNU * (N0s[i] + N))), df = 2 * cNU)),
         col = i, pch = c(1, 16))
}
legend("topright",
       paste0(rep(c("Posterior (N0 = ", "Prior (N0 = "), each = length(N0)),
              rep(N0, 2), ")"),
       lty = rep(c(1, 2), each = length(N0)),
       pch = rep(c(16, 1), each = length(N0)), 
       col = rep(seq_along(N0), 2))
```

## Example 10.13: Labor market data - Savage-Dickey density ratio for the no-income-risk homogeneity test

We only need to re-estimate the heterogeneity model which we use to simulate
the prior and the posterior of the no-income-risk difference.

```{r, echo = -1}
if (pdfplots) {
  pdf("10-1_3a.pdf", width = 5, height = 4)
}
par(mar = c(2.5, 1.5, .5, .8), mgp = c(1.6, .6, 0))
set.seed(42)
M <- 10000000

a01 <- a02 <- 1
b01 <- b02 <- 1
N1 <- with(labor, sum(wcollar))
SN1 <- with(labor, sum(wcollar & unemployed))
N2 <- with(labor, sum(!wcollar))
SN2 <- with(labor, sum(!wcollar & unemployed))

aN1 <- a01 + SN1
bN1 <- b01 + N1 - SN1
aN2 <- a02 + SN2
bN2 <- b02 + N2 - SN2

psi <- rbeta(M, aN1, bN1) - rbeta(M, aN2, bN2)
mybreaks <- seq(floor(200 * min(psi)) / 200 - 0.0025,
                ceiling(200 * max(psi)) / 200 + 0.0025,
                by = 0.005)
hist(psi, breaks = mybreaks, freq = FALSE, xlab = expression(psi),
     main = "", ylab = "", border = NA)
lines(c(-1, 0, 1), c(0, 1, 0), lty = 2, lwd = 2)
abline(v = 0, lty = 3)
abline(h = 0, lty = 3)

# Estimate density
lines(d <- density(psi), lwd = 2)
```

We can approximate the Bayes factor through the estimated Savage-Dickey
density ratio. Note, though, that this can be a very poor approximation.

```{r}
# Evaluate as close a possible to zero (with linear interpolation)
x1 <- sum(d$x < 0) # Find largest x below 0
x2 <- x1 + 1L      # Find smallest x above 0
pos <- -d$x[x1] / (d$x[x2] - d$x[x1])
dpost <- (1 - pos) * d$y[x1] + pos * d$y[x2]

# The prior is a symmetric triangular distribution on [-1, 1], thus p(0) = 1
dprior <- 1

(logSD <- log(dpost) - log(dprior))
```

## Example 10.14: Stomach cancer data - Savage-Dickey density ratio for the mortality rate homogeneity test

As above, with different data.

```{r, echo = -1}
if (pdfplots) {
  pdf("10-1_3b.pdf", width = 5, height = 4)
}
par(mar = c(2.5, 1.5, .5, .8), mgp = c(1.6, .6, 0))
N1  <- 1668 # number at risk in City A
SN1 <- 2    # cancer deaths in City A
N2  <- 583  # number at risk in City B
SN2 <- 1    # cancer deaths in City B

aN1 <- a01 + SN1
bN1 <- b01 + N1 - SN1
aN2 <- a02 + SN2
bN2 <- b02 + N2 - SN2

psi <- rbeta(M, aN1, bN1) - rbeta(M, aN2, bN2)
mybreaks <- seq(floor(1000 * min(psi)) / 1000 - 0.0005,
                ceiling(1000 * max(psi)) / 1000 + 0.0005,
                by = 0.001)
d <- density(psi)
hist(psi, breaks = mybreaks, freq = FALSE, xlab = expression(psi),
     main = "", ylab = "", ylim = c(0, max(d$y)), border = NA)
lines(d, lwd = 2)
lines(c(-1, 0, 1), c(0, 1, 0), lty = 2, lwd = 2)
abline(v = 0, lty = 3)
abline(h = 0, lty = 3)
```

We can approximate the Bayes factor through the estimated Savage-Dickey
density ratio. Note, though, that this can be a very poor approximation.

```{r}
x1 <- sum(d$x < 0) # Find largest x below 0
x2 <- x1 + 1L      # Find smallest x above 0
pos <- -d$x[x1] / (d$x[x2] - d$x[x1])
dpost <- (1 - pos) * d$y[x1] + pos * d$y[x2]
dprior <- 1
(logSD <- log(dpost) - log(dprior))
```
## Example 10.15: Road safety data - Lindley's paradoxon

```{r}
a0 <- c(0.01, 0.1, 0.5, 1:5)
res <- rbind(gamma(a0))
colnames(res) <- a0
knitr::kable(res)
```

# Section 10.4: Addressing Model Uncertainty

We re-use the marginal likelihoods from above to compute posterior
probabilities under a uniform prior for the individual models.

```{r}
res <- rbind(resM1, resM3)
probs_unnormalized <- exp(res - max(res))
probs <- probs_unnormalized / sum(probs_unnormalized)

knitr::kable(probs, digits = 2)
```

We now evaluate the posteriors for $\xi$ for the different hyperparameter
values.

```{r}
posterior_xi <- function(xi, y, a0, m0, alpha_b0, log = FALSE) {
  N <- length(y)
  alpha_bN <- alpha_b0 + N * a0
  beta_bN  <- (alpha_b0 - 1) * m0 / a0 + sum(y)
  logdens <- -a0 - lbeta(alpha_bN, beta_bN) +
             (alpha_bN / a0 - 1) * log(xi) +
             (beta_bN - 1) * log(1 - xi^(1 / a0))
  if (log) logdens else exp(logdens)
}

xis <- seq(0, 1, length.out = 700)
post_xis <- array(NA_real_,
                  dim = c(length(a0_tmp), length(m0_tmp), length(xis)),
                  dimnames = list(a0 = a0_tmp, m0 = round(m0_tmp, 2), xi = xis))
for (i in seq_along(a0_tmp)) {
  for (j in seq_along(m0_tmp)) {
    post_xis[i, j, ] <- posterior_xi(xis, eyetracking$anomalies,
                                     a0_tmp[i], m0_tmp[j], 6)
  }
}
```

Now we visualize, first the individual (model-specific) posteriors, then the BMA
density and (again) the one for the model with highest posterior probability.

```{r, echo = -c(1:2)}
if (pdfplots) {
  pdf("10-4_1.pdf", width = 8, height = 4)
}
par(mfrow = c(1, 2), mar = c(2.5, 1.5, 1, .5), mgp = c(1.6, .6, 0))
plot(NULL, xlim = c(0, 1), ylim = range(post_xis), xlab = expression(xi),
     ylab = "", main = "Model-specific posterior densities")
legend("topright",
       legend = c(paste("a0 = ", a0_tmp), "", paste("m0 = ", round(m0_tmp, 2))),
       lty = c(rep(1, length(a0_tmp)), NA, seq_along(m0_tmp)),
       col = c(seq_along(a0_tmp) + 1, NA, rep(1, length(m0_tmp))))
for (i in seq_along(a0_tmp)) {
  for (j in seq_along(m0_tmp)) {
    lines(xis, post_xis[i, j, ], col = i + 1, lty = j)
  }
}
abline(h = 0, lwd = 1.5)

weighted_post_xis <- apply(as.numeric(probs[5:8, ]) * post_xis, 3, sum)
argmax <- which(probs[5:8, ] == max(probs[5:8, ]), arr.ind = TRUE)

plot(xis, post_xis[argmax[1, 1], argmax[1, 2], ], type = "l", lwd = 3,
     xlim = c(0.27, 0.43), col = argmax[1, 1] + 1, lty = argmax[1, 2],
     main = "Most probable and BMA posterior", xlab = expression(xi))
lines(xis, weighted_post_xis, lwd = 2)
legend("topright", legend = c("BMS", "BMA"),
       col = c(argmax[1, 1] + 1, 1), lty = c(argmax[1, 2], 1), lwd = c(3, 2))
abline(h = 0, lwd = 1.5)
```

