---
title: "Chapter 12: Model Assessment and Predictive Model Checking"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Chapter 12: Model Assessment and Predictive Model Checking}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6
)
knitr::opts_knit$set(global.par = TRUE)
pdfplots <- FALSE # default: FALSE; set this to TRUE only if you like pdf figures
mcmcspeedup <- 10L # default: 1L; set this to something higher for fewer draws
seed <- 42
```

# Section 12.1: Beyond classical Bayesian model comparison

## Section 12.1.1: Schwarz criterion and BIC

### Example 12.1: Labor market data: Bayesian variable selection using BIC

We return to the probit analysis of Example 11.11 and run variable selection 
using the Schwarz criterion instead of the marginal likelihood. 
We first implement this approach.

```{r}
modsel_probit_BIC <- function(y, X, burnin = 1000L, M = 5000L){
  n <- dim(X)[1]
  p <- dim(X)[2] # change to X later

  gamma_post <- matrix(ncol = p, nrow = M)
  acc <- numeric(length = M)

  gamma <- matrix(rep(1, p), nrow = 1)
  mod0 <- glm(y ~ X, family = binomial(link = "probit"))
  SC_old <- logLik(mod0) - log(n) * (p + 1) / 2

  for (m in seq_len(burnin + M)) {
    gamma_proposed <- gamma_old <- gamma

    j <- sample(1:p, size = 1)
    gamma_proposed[j] <- 1 - gamma_old[j]

    X_gamma <- X[, gamma_proposed == 1]
    model <- glm(y ~ X_gamma, family = binomial(link = "probit"))
    SC_proposed <- logLik(model) - log(n) * (sum(gamma_proposed) + 1) / 2
    
    # compute acceptance probability and decide on acceptance
    log_acc <- SC_proposed - SC_old

    if (log(runif(1)) < log_acc) {
      gamma <- gamma_proposed
      SC_old <- SC_proposed
      accept <- 1
    } else {
      gamma <- gamma_old
      accept <- 0
    }

    if (m > burnin) {
      gamma_post[m - burnin, ] <- gamma
      acc[m - burnin] <- accept
    }
  }
  return(list(gamma_post = gamma_post, acc = acc))
}
```

We next load the data and as in exercise 11.11. add 5 irrelevant, standard 
normal regressors. Then we run the algorithm. 

```{r}
library("BayesianLearningCode")
data("labor", package = "BayesianLearningCode")
y <- labor$income_1998 == "zero"
N <- length(y)
X_unemp <- with(labor, cbind(female = female,
                             age18 = 1998 - birthyear - 18,
                             wcollar = wcollar_1997,
                             unemp97 = income_1997 == "zero")) # regressor matrix
p_irrel <- 5
set.seed(seed)
X <- cbind(X_unemp, matrix(rnorm(p_irrel * N), ncol = p_irrel))
colnames(X)[4 + (1:p_irrel)] <-
  c("irrel1", "irrel2", "irrel3", "irrel4", "irrel5")

p <- dim(X)[2]
M <- 20000 / mcmcspeedup
res <- modsel_probit_BIC(y, X, M = M)
```

We determine the number of models visited during MCMC, the average model 
complexity, i.e. the average number of covariates included in the model and show 
the posterior inclusion probabilities.

```{r}
if (pdfplots) {
  pdf("12-1_1.pdf", width = 5, height = 6)
}
par(mfrow = c(1, 1), mar = c(2.5, 2.5, 1.5, .5), mgp = c(1.5, .5, 0))

gammas_unique <- unique(res$gamma_post)
num_mod <- dim(gammas_unique)[1]
print(num_mod)

k_gamma <- rowSums(res$gamma_post)
print(mean(k_gamma))

# PIPs
print(colMeans(res$gamma_post))

barplot(colMeans(res$gamma_post), col = "blue", names.arg = 1:p,
        xlab = "Covariate", ylab = "PIP")
```

To determine the model visited most often we re-use a function which we defined
in Chapter 11.

```{r}
number_draws <- function(gamma_post, models) {
  nmod <- dim(models)[1]
  freq <- rep(NA, nmod)
  
  for (j in (1:nmod)) {
    freq[j] <- sum(apply(gamma_post, 1, function(x) identical(x, models[j,])))
  }
  freq
}

freq_gammas <- number_draws(res$gamma_post, gammas_unique)
io <- order(freq_gammas, decreasing = TRUE)

knitr::kable(cbind(gammas_unique, freq_gammas / M)[io[1:5], ],
             digits = cbind(rep(0, p), 4))
```

## Section 12.1.2: Perspectives on Bayesian model comparison

[no code]

# Section 12.2: Comparative model assessment through predictive information criteria

## Section 12.2.1: Measuring predictive loss

[no code]

## Section 12.2.2: Akaike information criterion (AIC)

### Example 12.2: U.S. GDP data: Choosing the model order via AIC and BIC

We need to re-load the function which creates the design matrix.

```{r}
ARdesignmatrix <- function(dat, p = 1, conditioninglength = p) {
  d <- p + 1
  N <- length(dat) - p

  Xy <- matrix(NA_real_, N, d)
  Xy[, 1] <- 1
  for (i in seq_len(p)) {
    Xy[, i + 1] <- dat[(p + 1 - i) : (length(dat) - i)]
  }
  Xy[(1 + conditioninglength - p):N, , drop = FALSE]
}
```

Let's load the data and compute log returns.

```{r}
data(gdp, package = "BayesianLearningCode")
logret <- diff(log(gdp))
```

Now we compute five OLS estimates and the corresponding AICs and BICs.

```{r}
AIC <- BIC <- rep(NA_real_, 5)
for (i in seq_along(AIC)) {
  y <- tail(logret, -4)
  p <- i - 1     # lag length
  d <- i + 1     # total number of parameters
  N <- length(y) # number of data points used for estimation
  Xy <- ARdesignmatrix(logret, p, conditioninglength = 4)
  fit <- lm(y ~ Xy - 1)
  betahat <- coef(fit)
  sigma2hat <- sum(resid(fit)^2) / length(y)
  minus2timesloglikmax <- N * (1 + log(2 * pi)) + N * log(sigma2hat)
  AIC[i] <- minus2timesloglikmax + 2 * d      # equals bult-in AIC(fit)
  BIC[i] <- minus2timesloglikmax + log(N) * d # equals built-in BIC(fit)
}
knitr::kable(rbind(AIC, BIC), digits = 2)
```


### Example 12.6: CHF exchange rate data: Testing normal vs. t using DIC

Let's load the data first.

```{r}
data("exrates", package = "stochvol")
y <- 100 * diff(log(exrates$USD / exrates$CHF))
X <- 1
N <- length(y)
```

Under the zero-mean Gaussian model, the posterior of $\sigma^2$ is inverse
gamma, and we can easily draw from it.

```{r}
c0 <- 2.5
C0 <- 1.5

SSR <- sum(y^2)
cN <- c0 + N / 2
CN <- C0 + SSR / 2

M <- 10000 / mcmcspeedup
sigma2normal <- rinvgamma(M, cN , CN)
```

Under the t distribution with 7 degrees of freedom, the posterior of $\sigma^2$
is not available in closed form, but we can draw from it using data
augmentation.

```{r}
set.seed(1)
burnin <- 100

# fix df
nu <- 7

# allocate space for storing the draws
sigma2t7 <- rep(NA_real_, M)

# starting value for w
w <- rep(1, N)

# pre-compute cN
cN <- c0 + N / 2

# log likelihood
loglik <- function(y, sigma2, nu) {
  length(y) * (lgamma((nu + 1) / 2) - lgamma(nu / 2) - .5 * log(nu)) -
  (nu + 1) / 2 * sum(log(1 + y^2 / nu / sigma2))
}

for (m in 1:(burnin + M)) {
  # sample sigma^2 from its full conditional (S-DA)
  eps <- sqrt(w) * y
  CN <- C0 + crossprod(eps) / 2
  sigma2 <- rinvgamma(1, cN, CN)

  # sample w (W-DA)
  r <- eps^2 / (w * sigma2)
  w <- rgamma(length(eps), (nu + 1) /2, (nu + r) / 2)

  # store the results
  if (m > burnin) sigma2t7[m - burnin] <- sigma2
}
```


For sampling also the degrees of freedom under an exponential prior, 
we re-run the sampler from Example 8.16, but this time without an intercept.


```{r}
set.seed(1)

# fix tuning parameter for MH, the number of draws M, and burn-in length
cnu <- 0.3
burnin <- 1000

# fix prior hyperparameters
lambda <- 1 / 7

# allocate space for storing the draws
nus <- mus <- sigma2s <- rep(NA_real_, M)
ws <- matrix(NA_real_, nrow = M, ncol = N)

# starting value for sigma2, log(nu), and w
sigma2 <- var(y) / 2
w <- rep(1, N)
nu <- 10

# pre-compute cN
cN <- c0 + N / 2

# log likelihood
loglik <- function(y, sigma2, nu) {
  length(y) * (lgamma((nu + 1) / 2) - lgamma(nu / 2) - .5 * log(nu)) -
  (nu + 1) / 2 * sum(log(1 + y^2 / nu / sigma2))
}

accepts <- 0L
for (m in 1:(burnin + M)) {
  # sample sigma^2 from its full conditional (S-DA)
  eps <- sqrt(w) * y
  CN <- C0 + crossprod(eps) / 2
  sigma2 <- rinvgamma(1, cN, CN)

  # sample nu (N-MH)
  nuprop <- exp(rnorm(1, log(nu), cnu))
  logR <- loglik(y, sigma2, nuprop) -
          loglik(y, sigma2, nu) +
          dexp(nuprop, lambda, log = TRUE) -
          dexp(nu, lambda, log = TRUE) +
          log(nuprop) -
          log(nu)
  
  if (log(runif(1)) < logR) {
    nu <- nuprop
    if (m > burnin) accepts <- accepts + 1L
  }
  
  # sample w (W-DA)
  r <- eps^2 / (w * sigma2)
  w <- rgamma(length(eps), (nu + 1) /2, (nu + r) / 2)

  # store the results
  if (m > burnin) {
    sigma2s[m - burnin] <- sigma2
    nus[m - burnin] <- nu
    ws[m - burnin, ] <- w
  }
}
```
