---
title: "Chapter 8: Bayesian Learning Beyond Standard Regression Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Chapter 8: Bayesian Learning Beyond Standard Regression Analysis}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6
)
knitr::opts_knit$set(global.par = TRUE)
pdfplots <- FALSE # default: FALSE; set this to TRUE only if you like pdf figures
```

```{r, include = FALSE}
par(mgp = c(1.6, .6, 0), mar = c(2.6, 2.6, 2.6, .4), lwd = 1)
```

# Section 8.1.1: Probit model 
## Example 8.1: Labour market data

We illustrate probit regression analysis for the labor market data.

```{r}
library("BayesianLearningCode")
data("labor", package = "BayesianLearningCode")
```

We model the binary variable unemployment as dependent variable and
use as covariates the variables female (binary), wcollar (binary),
age18 (quantitative, centered at 18 years) and unemployed in 1997
(binary). The baseline person is hence an 18 year old male blue collar
worker who was employed in 1997.

```{r}
y.unemp <- labor$unemp98
N.unemp <- length(y.unemp)  # number of observations 
X.unemp <- with(labor, cbind(intercept = rep(1, N.unemp),
                             female = female,
                             age18 = age-18,
                             wcollar = wcollar,
                             unemp97 = unemp97)) # regressor matrix
```

The regression coefficients are estimated using data augmentation and
Gibbs sampling. We define a function yielding posterior draws using
the algorithm detailed in Chapter 8.1.1.


```{r}
probit <- function(y, X, b0 = 0, B0 = 10000,
                   burnin = 1000L, M = 5000L) {
  N <- length(y)
  d <- ncol(X) # number regression effects 

  B0.inv <- diag(rep(1/B0, length.out = d), nrow = d) 
  b0 <- rep(b0, length.out = d) 
  B0inv.b0 <- B0.inv %*% b0

  betas <- matrix(NA_real_, nrow = M, ncol = d)
  colnames(betas) <- colnames(X)

  # define quantities for the Gibbs sampler
  BN <- solve(B0.inv + crossprod(X))
  ind0 <- (y == 0) # indicators for zeros
  ind1 <- (y == 1) # indicators for ones

  # starting values
  beta <- rep(0, d)
  z <- rep(NA_real_, N)

  for (m in seq_len(burnin + M)) {
    # Draw z conditional on y and beta
    u <- runif(N)
    eta <- X %*% beta
    pi <- pnorm(eta) 
   
    z[ind0] <- eta[ind0] + qnorm(u[ind0] * (1 - pi[ind0]))
    z[ind1] <- eta[ind1] + qnorm(1 - u[ind1] * pi[ind1])
  
    # sample beta from the full conditional 
    bN <- BN %*% (B0inv.b0 + crossprod(X, z))
    beta <- t(mvtnorm::rmvnorm(1, mean = bN, sigma = BN))
    
    # Store the beta draws
    if (m > burnin) {
      betas[m-burnin, ] <- beta
    }
 }
 return(betas)
}
```

We specify the prior on the regression effects as a rather flat
Normal independence prior and estimate the model.

```{r}
set.seed(1234)
betas <- probit(y.unemp, X.unemp, b0 = 0, B0 = 10000)
```

To compute summary statistics from the  posterior we use the following function. 
```{r}
res.mcmc <- function(x, lower = 0.025, upper = 0.975) {
  y <- c(quantile(x, lower), mean(x), quantile(x, upper))
  names(y) <- c(paste0(lower * 100, "%"), "Mean", paste0(upper * 100, "%"))
  y
}    
```


```{r}
res_beta <- apply(betas, 2, res.mcmc)
knitr::kable(round(res_beta, 3))

(p_unemploy_base <- pnorm(res_beta[1, 2]))
```
The estimated risk of unemployment for a baseline person is low and it
is even lower for a white collar worker. It is higher for females,
older persons and particularly for those unemployed in 1997.

```{r, echo = -c(1:2)}
if (pdfplots) {
  pdf("8-1_1.pdf", width = 8, height = 5)
}
par(mfrow = c(2, 3), mar = c(2.5, 1.5, 1.5, .1), mgp = c(1.5, .5, 0), lwd = 1.5)
for (j in seq_len(ncol(betas))) {
  hist(betas[, j], freq = FALSE, main = "", xlab = colnames(betas)[j], 
       ylab = "")
}
```

A plot of the autocorrelation of the draws shows that although there is some autocorrelation, it vanishes after a few lags.

```{r, echo = -c(1:2)}
if (pdfplots) {
  pdf("8-1_2.pdf", width = 8, height = 5)
}
par(mfrow = c(2, 3), mar = c(2.5, 1.5, 1.5, .1), mgp = c(1.5, .5, 0), lwd = 1.5)
for (j in seq_len(ncol(betas))) {
  acf(betas[, j], main = "", xlab = colnames(betas)[j], ylab = "")
}
```
The sampler is easy to implement, however there  might be problems  when the response variable  contains either only few or very many successes.
To illustrate this issue, we use data where in $N=500$ trials  only  1 success  or only 1 failure is observed.

```{r}
N <- 500
X <- matrix(1, nrow = N)

y1 <- c(0, rep(1, N-1))
betas1 <- probit(y1, X, b0 = 0, B0 = 10000)

y2 <- c(rep(0, N-1), 1)
betas2 <- probit(y2, X, b0 = 0, B0 = 10000) 
```

In both cases the autocorrelation of the draws decreases very slowly.
```{r, echo = -c(1:2)}
if (pdfplots) {
  pdf("8-1_3.pdf", width = 8, height = 5)
}
par(mfrow = c(2, 2), mar = c(2.5, 1.5, 1.5, .1), mgp = c(1.5, .5, 0), lwd = 1.5)
plot(betas1, type = "l", main = "", xlab = "", ylab = "")
acf(betas1)
plot(betas2, type = "l", main = "", xlab = "", ylab = "")
acf(betas2)
```

High autocorrelated draws in probit models not only occur if successes
or failures are very rare, but also when a covariate (or a linear
combination of covariates) perfectly allows to predict successes
and/or failures.  Complete separation means that both successes and
failures can be perfectly predicted by a covariate, whereas with
quasi-complete separation only either successes or failures can be
predicted perfectly.

To illustrate posterior sampling, in the case of complete separation,
we simulate $N=500$ observations with 10 successes and 490 failures.
We add a binary predictor $x$ where for $x=1$ we observe only
successes and for $x=0$ only failures.

```{r}
N <- 500
ns <- 10
x <- rep(c(1, 0), c(ns, N-ns))
y <- rep(c(1, 0), c(ns, N-ns))

table(y, x)
```
We estimate the  model parameters and plot the ACF of the draws. Again the autocorrelations remain high even for lag 35.

```{r, echo = -c(1:2)}
if (pdfplots) {
  pdf("8-1_4.pdf", width = 8, height = 5)
}
par(mfrow = c(2, 2), mar = c(2.5, 1.5, 1.5, .1), mgp = c(1.5, .5, 0), lwd = 1.5)
X <- cbind(rep(1, N), x)
betas <- probit(y, X, b0 = 0, B0 = 10000)

plot(betas[, 1], type = "l", main = "", xlab = "", ylab = "")
acf(betas[, 1])
plot(betas[, 2], type = "l", main = "", xlab = "", ylab = "")
acf(betas[, 2])
```

Finally we simulate another data set of $N=500$ observations with 10
successes and 490 failures. In this data for $x=1$ only successes but
for $x=0$ successes as well as failures are observed and hence there
is quasi-separation.

```{r}
N <- 500
x <-  rep(c(1, 0), c(ns, N-ns))
y <- rep(c(1, 0), each = N/2)

table(y, x)
```
A plot of the acf of the draws shows low autocorrelation for the
intercept but autocorrelations for the covariate effect are again
high.

```{r, echo = -c(1:2)}
if (pdfplots) {
  pdf("8-1_5.pdf", width = 8, height = 5)
}
par(mfrow = c(2, 2), mar = c(2.5, 1.5, 1.5, .1), mgp = c(1.5, .5, 0), lwd = 1.5)
X <- cbind(rep(1, N), x)
betas <- probit(y, X, b0 = 0, B0 = 10000)

plot(betas[, 1], type = "l", main = "", xlab = "", ylab = "")
acf(betas[, 1])
plot(betas[, 2], type = "l", main = "", xlab = "", ylab = "")
acf(betas[, 2])
```

High autocorrelations typically indicate problems with the sampler. If
there is complete or quasi-complete separation in the data, the
likelihood is monotone and the maximum likelihood estiamte does not
exist.  In a Bayesian approach using a flat, improper prior on the
regression effects will result in an improper posterior
distribution. Hence, a proper prior is required to avoid improper
posteriors in case of separation.

In the examples above we used a proper prior which is rather
flat. With a more informative prior, the autocorrelations of the draws
are lower. This can be seen in the next figure, where the simulated
data under quasi-separation are re-analyzed with a Normal prior that
is tighter around zero.

```{r, echo = -c(1:2)}
if (pdfplots) {
  pdf("8-1_6.pdf", width = 8, height = 5)
}
par(mfrow = c(2, 2), mar = c(2.5, 1.5, 1.5, .1), mgp = c(1.5, .5, 0), lwd = 1.5)
betas <- probit(y, X, b0 = 0, B0 = 2.5^2)

plot(betas[, 1], type = "l", main = "", xlab = "", ylab = "")
acf(betas[, 1])
plot(betas[, 2], type = "l", main = "", xlab = "", ylab = "")
acf(betas[, 2])
```

# Section 8.1.2

We now estimate a logistic regression model for the labor market data
using the two-block Polya-Gamma sampler.

```{r}
logit <- function(y, X, b0 = 0, B0 = 10000,
                  burnin = 1000L, M = 5000L) {
  
  N <- length(y)
  d <- ncol(X) # number regression effects 

  B0.inv <- diag(rep(1/B0, length.out = d), nrow = d) 
  b0 <- rep(b0, length.out = d) 
  B0inv.b0 <- B0.inv %*% b0

  betas <- matrix(NA_real_, nrow = M, ncol = d)
  colnames(betas) <- colnames(X)

  # define quantities for the Gibbs sampler
  ind0 <- (y == 0) # indicators for zeros
  ind1 <- (y == 1) # indicators for ones
  
  # starting values
  beta <- rep(0, d)
  z <- rep(NA_real_, N)
  omega <-rep(NA_real_, N)

  for (m in seq_len(burnin+M)) {
    # Draw z conditional on y and beta
    eta <- X %*% beta
    pi <- plogis(eta) 
      
    u <- runif(N)
    z[ind0] <- eta[ind0] + qlogis(u[ind0]*(1-pi[ind0]))
    z[ind1] <- eta[ind1] + qlogis (1-u[ind1]*pi[ind1])
      
    # Draw omega conditional on y, beta and z
    omega <- pgdraw::pgdraw(b = 1, c = z-eta)
  
    # sample beta from the full conditional 
    Xomega <- matrix(omega, ncol = d, nrow = N) * X
    BN <- solve(B0.inv + crossprod(Xomega, X))
    bN <- BN %*% (B0inv.b0 + crossprod(Xomega, z))
    beta <- t(mvtnorm::rmvnorm(1, mean = bN, sigma = BN))
    
    # Store the beta draws
    if (m > burnin) {
      betas[m-burnin, ] <- beta
    }
 }
 return(betas)
}
```

We again use the flat independence Normal prior on the regression
effects and estimate the model.

```{r}
betas_logit <- logit(y.unemp, X.unemp, b0 = 0, B0 = 10000)
print(str(betas_logit))

res_beta_logit <- apply(betas_logit, 2, res.mcmc)
knitr::kable(round(res_beta_logit, 3))

(p_unemploy_base <- plogis(res_beta_logit[1, 2]))
```

Note that the logistic distribution has a variance of $\pi^2/3$ and
hence the regression effects are absolutely larger than in the probit
model. However any probability computed from the two models will be
very close, e.g., compare the probability to be unemployed for the
baseline person.

We can compare the posterior estimates of the coefficients in the
probit model $\beta$ those of of logit model by multiplying them with
$\pi/\sqrt{3}$ and see that there is not much difference.

```{r}
knitr::kable(round(res_beta * pi/sqrt(3), 3))
```

# Section 8.2.1:Poisson Regression models
## Example 8.3: Road Safety Data

We fit two different Poisson regression models:

1. a small model with intercept, intervention effect and holiday dummy
   (activated in July/August);

2. a large model with intercept, intervention effect, linear trend,
   and a seasonal pattern captured by monthly dummies.

The sampler performance for these two models is assessed to study how
the acceptance rate deteroriates, if $d$ increases.

We load the data and extract the observations for the children in
Linz. 

```{r}
data("accidents", package = "BayesianLearningCode")
y <- accidents[, "children_accidents"]
e <- rep(1, length(y))
```

Then, we define the regressor matrix.
```{r}
X <- cbind(intercept = rep(1, length(y)),
           intervention = rep(c(0, 1), c(7*12+9, 8*12+3)),
           holiday = rep(rep(c(0, 1, 0), c(6, 2, 4)), 16))
```

To compute the parameters of the Normal proposal density, we use the
Newton-Raphson estimator described in Section 8.2.1.

```{r}
gen.proposal.poisson <- function(y, X, e, b0 = 0, B0 = 100, t.max = 30) {
  N <- length(y)
  d <- ncol(X)
  betas <- matrix(NA_real_, ncol = t.max, nrow = d)
  beta.old <- c(log(mean(y)), rep(0, d-1))

  b0 <- rep(b0, length.out = d)
  B0.inv <- diag(rep(1/B0, length.out = d), nrow = d) 

  for (t in 1:t.max) {
    rate <- e * exp(X %*% beta.old)
    score <- t(crossprod(y - rate, X) - (beta.old - b0) %*% B0.inv)

    H <- -B0.inv
    for (i in 1:N) {
      H <- H - rate[i] * tcrossprod(X[i, ])
    }
    betas[, t] <- beta.old - solve(H, score)
  }
  qmean <- betas[, t.max]
  
  # determine the variance matrix  
  rate <- e * exp(X %*% qmean)
  H <- -B0.inv
  for (i in 1:N) {
     H <- H - rate[i] * tcrossprod(X[i, ])
  }
  qvar <- -solve(H)
  return(parms.proposal = list(mean = qmean,
                               var = qvar))
}
```

We use a flat Normal prior on the regression effects.

```{r}
parms.proposal <- gen.proposal.poisson(y, X, e, b0 = 0, B0 = 100)
print(parms.proposal)
```

Next we set up the independence Metropolis-Hastings algorithm and
estimate the model parameters.

```{r}
poisson <- function(y, X, e, b0 = 0, B0 = 100, qmean, qvar,
                    burnin = 1000L, M = 5000L) {
  d <- ncol(X)
  beta.post <- matrix(ncol = d, nrow = M)
  colnames(beta.post) <- colnames(X)
  
  acc <- numeric(length = M)
  b0 <- rep(b0, length.out = d)
  B0 <- diag(rep(B0, length.out = d), nrow = d)
  
  beta <- as.vector(mvtnorm::rmvnorm(1, mean = qmean, sigma = qvar))
  
  for (m in seq_len(burnin + M)) {
    beta.old <- beta
    beta.proposed <- as.vector(mvtnorm::rmvnorm(1, mean = qmean, sigma = qvar))
    
    # compute log proposal density at proposed and old value
    lq_proposed <- mvtnorm::dmvnorm(beta.proposed, mean = qmean, sigma = qvar,
                                    log = TRUE)
    lq_old  <- mvtnorm::dmvnorm(beta.old, mean = qmean, sigma = qvar,
                                log = TRUE)
    
    # compute log prior of proposed and old value
    lpri_proposed <- mvtnorm::dmvnorm(beta.proposed, mean = b0, sigma = B0,
                                      log = TRUE)
    lpri_old  <- mvtnorm::dmvnorm(beta.old, mean = b0, sigma = B0,
                                  log = TRUE)
    
    # compute loglikelihood of proposed and old value
    lh_proposed <- dpois(y, e * exp(X %*% beta.proposed), log = TRUE)
    lh_old  <- dpois(y, e * exp(X %*% beta.old), log = TRUE)
    
    maxlik <- max(lh_old, lh_proposed)
    ll <- sum(lh_proposed - maxlik) - sum(lh_old - maxlik)
    
    # compute acceptance probability and accept or not
    log_acc <- min(0, ll + lpri_proposed - lpri_old + lq_old - lq_proposed)
    
    if (log(runif(1)) < log_acc) {
      beta <- beta.proposed
      accept <- 1
    } else {
      beta <- beta.old
      accept <- 0
    }

    # Store the beta draws
    if (m > burnin) {
        beta.post[m-burnin, ] <- beta
        acc[m-burnin] <- accept
    }
  }
  return(res = list(beta.post = beta.post, accept = mean(acc)))
}
```

```{r}
res1 <- poisson(y, X, e, b0 = 0, B0 = 100,
                qmean = parms.proposal$mean, qvar = parms.proposal$var)
res.poisson1 <- t(rbind(apply(res1$beta.post, 2, res.mcmc),
                        `exp(Mean)` = exp(colMeans(res1$beta.post))))
knitr::kable(round(res.poisson1, 3))
print(res1$accept)
```

We then fit an alternative model with intercept, intervention effect,
linear trend and saisonal dummy variables.

```{r}
seas <- rbind(diag(1, 11), rep(-1, 11)) 
seas.dummies <- matrix(rep(t(seas), 16), ncol = 11, byrow = TRUE)
colnames(seas.dummies) <- c("Jan", "Feb", "Mar", "Apr","May", "Jun", "Jul",
                            "Aug", "Sep", "Oct", "Nov")
X.large <- cbind(X,
                 lin.trend = 1:length(y),
                 seas.dummies)
```

We set the prior parameters and compute parameters of the proposal distribution 
```{r}
parms.proposal2 <- gen.proposal.poisson(y, X.large, e, b0 = 0, B0 = 100)
```

and fit the model.
```{r}
res2 <- poisson(y, X.large, e, b0 = 0, B0 = 100,
                qmean = parms.proposal2$mean, qvar = parms.proposal2$var)
res.poisson2 <- t(rbind(apply(res2$beta.post, 2, res.mcmc),
                        `exp(Mean)` = exp(colMeans(res2$beta.post))))
knitr::kable(round(res.poisson2, 3))
print(res2$accept)
```
```{r}
# xtable(res.poisson1, caption= "Count Data: estimation results for Model 1", label="res.count1", digits=3)
 
 # xtable(res.poisson2, caption= "Count Data: estimation results for Model 2", label="res.count2", digits=3)
```

# Section 8.2.2.: Negative binomial regression
# Example 8.4.

Now we analyse the road safety data allowing for unobserved heterogeneity.
We first set up both the two versions of the  three-block MH-within-Gibbs sampler 

```{r}
negbin <- function(y, X, e, b0 = 0, B0 = 100, qmean, qvar, pri.alpha,
                    full.gibbs= FALSE, burnin = 1000L, M = 50000L) {
  
  N=length(y)
  d <- ncol(X)
  beta.post <- matrix(ncol = d, nrow = M)
  colnames(beta.post) <- colnames(X)
  
  b0 <- rep(b0, length.out = d)
  B0 <- diag(rep(B0, length.out = d), nrow = d)
  
  acc.beta <- numeric(length = M)
  
  alpha.post<- rep(NA,M)
  acc.alpha <-rep(NA,M)
  c_alpha <- 0.1
  
  # set starting values
  beta <- as.vector(mvtnorm::rmvnorm(1, mean = qmean, sigma = qvar))
  alpha <- pri.alpha$shape/pri.alpha$rate
  phi <- rep(1.,N)
  
  for (m in seq_len(burnin + M)){
     # Step 1: draw beta  
     beta.old <- beta
     beta.proposed <- as.vector(mvtnorm::rmvnorm(1, mean=qmean, sigma=qvar))

     # compute log proposal density at proposed and old value
     lq_proposed <- mvtnorm::dmvnorm(beta.proposed,mean=qmean, sigma=qvar, 
                                     log=TRUE)
     lq_old  <- mvtnorm::dmvnorm(beta.old, mean=qmean, sigma=qvar, log=TRUE)
            
     # compute log prior  of proposed and old value
     lpri_proposed <- mvtnorm::dmvnorm(beta.proposed, mean=b0, sigma=B0, 
                                       log=TRUE )
     lpri_old  <- mvtnorm::dmvnorm(beta.old,  mean=b0, sigma=B0, log=TRUE )

     # compute log likelihood of proposed and old value
     lh_proposed <- dpois(y,exp(X%*%beta.proposed),  log=TRUE)
     lh_old  <- dpois(y,exp(X%*%beta.old), log=TRUE)

     maxlik <- max(lh_old,lh_proposed)
     ll <- sum(lh_proposed-maxlik)-sum(lh_old-maxlik)

     # compute acceptance probability and accept or not
     log_acc <- min(0,ll+lpri_proposed-lpri_old+lq_old-lq_proposed)

     if (log(runif(1)) < log_acc){
        beta <-  beta.proposed
        acc.b <- 1
     }else{
        beta <-  beta.old
        acc.b <- 0
     }
     linpred <- X%*%beta

     # Step 2:  Sample alpha
     alpha.old <- alpha
     alpha.proposed <- alpha.old*exp(c_alpha*rnorm(1))
     
     if (full.gibbs){
        llik_alpha.proposed <- sum(dgamma(phi, shape=alpha.proposed, rate=alpha.proposed,
                                 log=TRUE))
        llik_alpha.old      <- sum(dgamma(phi, shape=alpha.old, rate=alpha.old,
                                 log=TRUE)) 
     }else{
       llik_alpha.proposed <- sum(dnbinom(y, size=alpha.proposed, 
                                          mu=e*exp(linpred), log = TRUE))
       llik_alpha.old      <- sum(dnbinom(y, size=alpha.old, 
                                     mu=e*exp(linpred), log = TRUE))
     }
     log_acc_alpha <- llik_alpha.proposed - llik_alpha.old+
                       dgamma(alpha.proposed, shape=pri.alpha$shape ,
                           rate=pri.alpha$rate,log=TRUE) -
                       dgamma(alpha.old, shape=pri.alpha$shape ,
                           rate=pri.alpha$rate,log=TRUE)+
                       log(alpha.proposed) - log(alpha.old)

     if (log(runif(1)) < log_acc_alpha){
        alpha <- alpha.proposed
        acc.a<- 1
     }else{
        alpha <- alpha.old
        acc.a <- 0
     }
   
    # Step 3:  sample phi from its full conditional
    phi <- rgamma (N,shape=alpha+y, rate=alpha+e*exp(linpred))
    
    # Save the draws
    if (m > burnin) {
        beta.post[m-burnin, ] <- beta
        acc.beta[m-burnin] <- acc.b
        
        alpha.post[m-burnin] <- alpha
        acc.alpha[m-burnin] <- acc.a
    }
  }
  return(res = list(beta.post = beta.post, acc.beta=acc.beta,
                    alpha.post=alpha.post,acc.alpha = acc.alpha))
}
```

We specify the prior on $\alpha$ as the Gamma distribution with shape 2 and rate 
0.5 and use both samplers to estimate the model parameters.
```{r}
pri.alpha <- data.frame(shape=2,rate=0.5)

res1 <- negbin(y, X, e,  qmean=parms.proposal$mean, 
               qvar=parms.proposal$var, pri.alpha=pri.alpha, full.gibbs = TRUE)

res.negbin.full <- cbind(apply(res1$beta.post, 2, res.mcmc), 
                         res.mcmc(res1$alpha.post))
colnames(res.negbin.full)[4]="alpha"
knitr::kable(round(res.negbin.full, 3))

               
res2 <- negbin(y, X, e,  qmean=parms.proposal$mean, 
               qvar=parms.proposal$var, pri.alpha=pri.alpha, full.gibbs = FALSE)

res.negbin.partial <- cbind(apply(res2$beta.post, 2, res.mcmc), 
                         res.mcmc(res2$alpha.post))
colnames(res.negbin.partial)[4]="alpha"
knitr::kable(round(res.negbin.partial, 3))
```
As expected estimation results using  both samplers are rather similar.
```{r}
require(xtable)
xtable(cbind(t(res.negbin.full), t(res.negbin.partial)), label="exam:negbin_est")
```

## Section 8.2.3 
```{r}
print(c(mean(res1$acc.beta), mean(res1$acc.alpha)))
print(c(mean(res2$acc.beta), mean(res2$acc.alpha)))

```

```{r}
if (pdfplots) {
  pdf("8-2_1.pdf", width = 8, height = 5)
}
par(mfrow = c(1,2), mar = c(2.5, 2.5, 1.5, .1), mgp = c(1.5, .5, 0), lwd = 1.5)

qqplot(res1$beta.post[,1], res2$beta.post[,1], xlab="Full Gibbs",
       ylab="Partial Gibbs", main="Intercept")
abline(a = 0, b = 1)

qqplot(res1$alpha.post, res2$alpha.post, xlab="Full Gibbs",
       ylab="Partial Gibbs", main="Alpha")
abline(a = 0, b = 1)
```
