---
title: "Chapter 8: Beyond Standard Regression Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Chapter 8: Beyond Standard Regression Analysis}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6
)
knitr::opts_knit$set(global.par = TRUE)
pdfplots <- FALSE # default: FALSE; set this to TRUE only if you like pdf figures
```

```{r, include = FALSE}
par(mgp = c(1.6, .6, 0), mar = c(2.6, 2.6, 2.6, .4), lwd = 1)
```

# Section 8.1: Binary response variables

## Section 8.1.1: Probit model

### Figure 8.1: Latent utility and outcome in the probit model

We start by visualizing a latent utility for a linear predictor
$\mathbf{x}\boldsymbol{\beta}$ with a value of 1.

```{r, echo = -c(1:2)}
if (pdfplots) {
  pdf("8-1_0.pdf", width = 8, height = 4)
}
par(mfrow = c(1, 3), mar = c(2.5, 2.5, 2.5, .1), mgp = c(1.5, .5, 0), lwd = 1.5)
curve(dnorm(x, mean = 1), from = -3, to = 5, col = "blue", 
      xlab = expression(z[i]), ylab = "")
abline(v = 0, col = "red")
 
dens <- curve(dnorm(x, mean = 1), from = -3, to = 5, n = 161, col = "blue",
              xlab = expression(paste(z[i], "|", y[i], "=0")), ylab = "")
abline(v = 0, col = "red")
polygon(c(dens$x[dens$x <= 0], 0), c(dens$y[dens$x <= 0], 0),
        col = "red", border = NA)

dens <- curve(dnorm(x, mean = 1), from = -3, to = 5, n = 161, col = "blue",
              xlab = expression(paste(z[i], "|", y[i], "=1")), ylab = "" )
abline(v = 0, col = "red")
polygon(c(dens$x[dens$x >= 0], 0), c(dens$y[dens$x >= 0], 0),
        col = "red", border = NA)
```

### Example 8.1: Labor market data

We now perform probit regression analysis for the labor market data.

```{r}
library("BayesianLearningCode")
data("labor", package = "BayesianLearningCode")
```

We model the income in 1998, binarized into unemployed (zero income)
and employed, as dependent variable, and use as covariates the
variables female (binary), age18 (quantitative, centered at 18 years),
wcollar (binary), and unemployed in 1997 (binary). The baseline person
is hence an 18 year old male blue collar worker who was employed in
1997.

```{r}
y.unemp <- labor$income_1998 == "zero"
N.unemp <- length(y.unemp)  # number of observations 
X.unemp <- with(labor, cbind(intercept = rep(1, N.unemp),
                             female = female,
                             age18 = 1998 - birthyear - 18,
                             wcollar = wcollar_1997,
                             unemp97 = income_1997 == "zero")) # regressor matrix
```

### Example 8.2: Fitting a probit model to the labor market data

The regression coefficients are estimated using data augmentation and
Gibbs sampling. We define a function yielding posterior draws using
the algorithm detailed in Section 8.1.1.

```{r}
probit <- function(y, X, b0 = 0, B0 = 10000,
                   burnin = 1000L, M = 20000L) {
  N <- length(y)
  d <- ncol(X) # number of regression effects 

  b0 <- rep(b0, length.out = d) 
  B0.inv <- diag(rep(1 / B0, length.out = d), nrow = d)
  B0inv.b0 <- B0.inv %*% b0

  betas <- matrix(NA_real_, nrow = M, ncol = d)
  colnames(betas) <- colnames(X)
  
  z <- rep(NA_real_, N)

  # Define quantities for the Gibbs sampler
  BN <- solve(B0.inv + crossprod(X))
  ind0 <- (y == 0) # indicators for zeros
  ind1 <- (y == 1) # indicators for ones

  # Set starting values
  beta <- c(qnorm(mean(y)), rep(0, d-1))

  for (m in seq_len(burnin + M)) {
    # Draw z conditional on y and beta
    u <- runif(N)
    eta <- X %*% beta
    pi <- pnorm(eta) 
   
    z[ind0] <- eta[ind0] + qnorm(u[ind0] * (1 - pi[ind0]))
    z[ind1] <- eta[ind1] + qnorm(1 - u[ind1] * pi[ind1])
  
    # Sample beta from the full conditional 
    bN <- BN %*% (B0inv.b0 + crossprod(X, z))
    beta <- t(mvtnorm::rmvnorm(1, mean = bN, sigma = BN))
    
    # Store the beta draws
    if (m > burnin) {
      betas[m - burnin, ] <- beta
    }
 }
 return(betas)
}
```

We specify the prior on the regression effects as a rather flat normal
independence prior and estimate the model parameters.
 
```{r}
set.seed(1234)
M <- 20000
betas <- probit(y.unemp, X.unemp, b0 = 0, B0 = 1000L, burnin = 1000, M = M)
```

To compute summary statistics from the posterior we use the following function. 

```{r}
res.mcmc <- function(x, lower = 0.025, upper = 0.975) {
  res <- c(quantile(x, lower), mean(x), quantile(x, upper))
  names(res) <- c(paste0(lower * 100, "%"), "Posterior mean", 
                  paste0(upper * 100, "%"))
  res
}
```

We show posterior means and
equal-tailed 95\% credible intervals of the regression effects.

```{r}
res_probit.labor <- t(apply(betas, 2, res.mcmc))
knitr::kable(round(res_probit.labor, 3))
```

Next, we determine the estimated risk of unemployment for a baseline
person, i.e., a 18 year old male blue collar worker who was employed in
1997, using the posterior mean estimate of the intercept.

```{r}
(p_unemploy_base_probit <- round(pnorm(res_probit.labor[1, 2]), 4))
```

The estimated risk to be unemployed in 1998 for a baseline person is
very low with a value of `r p_unemploy_base_probit` and even lower for
a white collar worker. This risk is higher for a female, an older
person and particularly high if the person was unemployed in 1997.

Next, we visualize the estimated posterior distributions for the
regression effects by histograms.

```{r, echo = -c(1:2)}
if (pdfplots) {
  pdf("8-1_1.pdf", width = 8, height = 5)
}
par(mfrow = c(2, 3), mar = c(2.5, 1.5, 1.5, .1), mgp = c(1.5, .5, 0), lwd = 1.5)
for (j in seq_len(ncol(betas))) {
  hist(betas[, j], freq = FALSE, main = "", xlab = colnames(betas)[j], 
       ylab = "")
}
```

A plot of the autocorrelation of the draws shows that although there
is some autocorrelation, it vanishes after a few lags.

```{r, echo = -c(1:2)}
if (pdfplots) {
  pdf("8-1_2.pdf", width = 8, height = 5)
}
par(mfrow = c(2, 3), mar = c(2.5, 2.5, 2.5, .1), mgp = c(1.5, .5, 0), lwd = 1.5)
for (j in seq_len(ncol(betas))) {
    acf(betas[, j], main = "", xlab = "Lag",
        ylab = "empirical ACF")
    title(colnames(betas)[j])
}
```

We also determine the estimated effective sample sizes (ESSs) to
assess the efficiency of the sampler. 

```{r}
ESS <- coda::effectiveSize(betas)
IF <- M / ESS
res_eff <- cbind(ESS = round(ESS, digits = 1),
                 IF = round(IF, digits = 2))
knitr::kable(res_eff)
```

The estimated effective sample size is at least 
`r min(round(ESS, digits = 1))` for every regression effect, and hence
all estimated inefficiency factors (IFs) are below 
`r max(round(IF, digits = 2))`.

The sampler is easy to implement, however there might be problems when
the response variable contains either only very few or very many
successes.

### Example 8.3: Imbalanced data

To illustrate this issue, we use data where in $N = 500$ trials only 1
 failure or only 1 success is observed.

```{r}
set.seed(1234)

N <- 500
X <- matrix(1, nrow = N)

y1 <- c(0, rep(1, N-1))
betas1 <- probit(y1, X, b0 = 0, B0 = 10000, burnin = 1000, M = M)

y2 <- c(rep(0, N-1), 1)
betas2 <- probit(y2, X, b0 = 0, B0 = 10000, burnin = 1000, M = M)
```

In both cases the empirical autocorrelation of the draws decreases
very slowly and remains high even for a lag of 40.

```{r, echo = -c(1:3)}
if (pdfplots) {
  pdf("8-1_3.pdf", width = 8, height = 5)
}
par(mfrow = c(2, 2), mar = c(2.5, 2.5, 1.5, .1), mgp = c(1.5, .5, 0), lwd = 1.5)

labels <- expression(beta[0])
plot(betas1, type = "l", main = "N=500, 1 failure", xlab = "Draws after burnin",
     ylab = labels)
acf(betas1, ylab = "empirical ACF")

(ESS1 <- coda::effectiveSize(betas1))
(IF1 <- round(M/ESS1, 2))

plot(betas2, type = "l", main = "N=500, 1 success", xlab = "Draws after burnin",
     ylab = labels)
acf(betas2, ylab = "empirical ACF")

(ESS2 <- coda::effectiveSize(betas2))
(IF2 <- round(M/ESS2, 2))
```

Hence for these data sets the estimated ESS of the intercept has a
value of at most `r ceiling(max(ESS1, ESS2) / 10) * 10`, 
yielding an estimated IF of at most `r ceiling(max(c(IF1, IF2) / 10) * 10)`.

High autocorrelation in MCMC draws for probit models occurs not only
when either successes or failures are rare, but also when a covariate
(or a linear combination of covariates) perfectly allows to predict
successes and/or failures.  Complete separation means that both
successes and failures can be perfectly predicted by a covariate,
whereas quasi-complete separation means that only either successes or
failures can be predicted perfectly.

### Example  8.4: Complete separation

To illustrate the effect of complete separation on the estimates, we
generate $N = 500$ observations where half of them are successes and
the other half are failures. We add a binary predictor $x$ where for
$x = 1$ we observe only successes and for $x = 0$ only failures.

```{r}
N <- 500
ns <- 250
x.sep <- rep(c(0, 1), c(ns, N - ns))
y <- rep(c(0, 1), c(ns, N - ns))

table(x.sep, y)
```

We estimate the model parameters $\boldsymbol{\beta}= (\beta_0,
\beta_1)'$ under the Normal prior with mean $\mathbf{0}$ and variance
matrix $10000\mathbf{I}$ and run the sampler for $M=20000$ iterations
after a burn-in of 1000.

From the plot of the empirical ACF of the draws we see that
autocorrelations are close to 1 even at lag 40.

```{r, echo = -c(1:3)}
if (pdfplots) {
  pdf("8-1_4.pdf", width = 8, height = 5)
}
par(mfrow = c(2, 2), mar = c(2.5, 2.5, 1.5, .1), mgp = c(1.5, .5, 0), lwd = 1.5)

set.seed(1234)
X.sep <- cbind(rep(1, N), x.sep)
betas.sep <- probit(y, X.sep, b0 = 0, B0 = 10000, burnin = 1000, M = M)

labels <- expression(beta[0], beta[1])
plot(betas.sep[, 1], type = "l", xlab = "Draws after burn-in", ylab = labels[1])
acf(betas.sep[, 1], ylab = "empirical ACF")

plot(betas.sep[, 2], type = "l", xlab = "Draws after burn-in", ylab = labels[2])
acf(betas.sep[, 2], ylab = "empirical ACF")

(ESS.sep <- coda::effectiveSize(betas.sep))
(IF.sep <- M/ESS.sep)
```

Hence the estimated ESSs are very low with a value of around 
`r round(mean(ESS.sep))`, resulting in estimated IFs of about 
`r round(mean(IF.sep), -2)`.

### Example 8.5: Quasi-complete separation

To illustrate quasi-separation we use the same responses as in Example
8.4, but now set $x=1$ for all successes and additionally for 100
failures. Hence for $x=0$ always a failure is observed, whereas for
$x=1$ both successes and failures occur.

```{r}
x.qus1 <- rep(c(0, 1), c(ns-100, N - ns+100))
table(x.qus1, y)
```

We again estimate the regression effects using data augmentation and Gibbs sampling.

```{r, echo = -c(1:3)}
if (pdfplots) {
  pdf("8-1_5.pdf", width = 8, height = 5)
}
par(mfrow = c(2, 2), mar = c(2.5, 2.5, 1.5, .1), mgp = c(1.5, .5, 0), lwd = 1.5)

set.seed(1234)
X.qus1 <- cbind(rep(1, N), x.qus1)
betas.qus1 <- probit(y, X.qus1, b0 = 0, B0 = 10000, burnin = 1000, M = M)

plot(betas.qus1[, 1], type = "l", xlab = "Draws after burn-in", ylab = labels[1])
acf(betas.qus1[, 1], ylab = "empirical ACF")

plot(betas.qus1[, 2], type = "l", xlab = "Draws after burn-in", ylab = labels[2])
acf(betas.qus1[, 2], ylab = "empirical ACF")

(ESS.qus1 <- coda::effectiveSize(betas.qus1))
(IF.qus1 <- M/ESS.qus1)
```

Again autocorrelations are very high for both the intercept as well as
the covariate effect resulting in high estimated IFs of
about `r round(mean(IF.qus1), -1)`.

We now change the setting so that $x$ takes values of $0$ not only for
failures but also for some successes, whereas $x=1$ for all successes.

```{r, echo = -c(1:3)}
if (pdfplots) {
  pdf("8-1_5a.pdf", width = 8, height = 5)
}
par(mfrow = c(2, 2), mar = c(2.5, 2.5, 1.5, .1), mgp = c(1.5, .5, 0), lwd = 1.5)

x.qus2 <- rep(c(0, 1), c(ns+100, N - ns-100))
table(x.qus2, y)

set.seed(1234)
X.qus2 <- cbind(rep(1, N), x.qus2)
betas.qus2 <- probit(y, X.qus2, b0 = 0, B0 = 10000, burnin = 1000, M = M)

plot(betas.qus2[, 1], type = "l", xlab = "Draws after burn-in", ylab = labels[1])
acf(betas.qus2[, 1], ylab = "empirical ACF")

plot(betas.qus2[, 2], type = "l", xlab = "Draws after burn-in", ylab = labels[2])
acf(betas.qus2[, 2], ylab = "empirical ACF")

(ESS.qus2 <- coda::effectiveSize(betas.qus2))
(IF.qus2 <- M/ESS.qus2)
```

Autocorrelations of the intercept are low and close to zero for small
lags but remain very high even at lag 40 for the covariate
effect. Hence we have a high ESS for the intercept 
(`r round(ESS.qus2[1])`) and a low for the covariate effect 
(`r round(ESS.qus2[2])`), resulting in an estimated IF of 
`r round(IF.qus2[1])` for the intercept, but of `r round(IF.qus2[2])` 
for the effect of the covariate.

High autocorrelations typically indicate problems with the sampler. If
there is complete or quasi-complete separation in the data, the
likelihood is monotone and the maximum likelihood estimate does not
exist.  In a Bayesian approach using a flat, improper prior on the
regression effects will result in an improper posterior
distribution. Hence, a proper prior is required to avoid improper
posteriors in case of separation and with a tighter prior we can
shrink coefficients to zero.


# Example 8.6: Complete separation: analysis under an informative prior

We now analyze the data of example 8.4. under the more informative
prior $\mathcal{N}(\mathbf{0}, \mathbf{I})$. This prior distribution
encodes the prior believe that $\beta_0$ and $\beta_1$ are in the
interval $(-1.96, 1.96)$$ with probability $0.95.
We compare the estimation results to those from example 8.4, where the
prior variance  was larger by a factor of 10000.

```{r}
set.seed(1234)
betas.sep1 <- probit(y, X.sep, b0 = 0, B0 = 1, burnin = 1000, M = M)

# compare results to the less informative prior
res_betas.sep <- t(apply(betas.sep, 2, res.mcmc))
rownames(res_betas.sep) <- c("Intercept", "X")

ESS.sep <- coda::effectiveSize(betas.sep)
IF.sep <- M/ESS.sep
res.sep <- round(cbind(res_betas.sep, ESS.sep, IF.sep), 2)
colnames(res.sep)[4:5] <- c("Estimated ESS", "Estimated IF")

knitr:: kable(res.sep)

res_betas.sep1 <- t(apply(betas.sep1, 2, res.mcmc))
rownames(res_betas.sep1) <- c("Intercept", "X")

ESS.sep1 <- coda::effectiveSize(betas.sep1)
IF.sep1 <-  M/ESS.sep1
res.sep1 <- round(cbind(res_betas.sep1, ESS.sep1, IF.sep1), 2)
colnames(res.sep1)[4:5] <- c("Estimated ESS", "Estimated IF")

knitr:: kable(res.sep1)
```
We see that the tighter prior shrinks the estimates to zero, estimated ESSs are higher and estimated IFs are lower.

```{r, echo = -c(1:3)}
if (pdfplots) {
  pdf("8-1_6.pdf", width = 8, height = 5)
}
par(mfrow = c(2, 2), mar = c(2.5, 2.5, 1.5, .1), mgp = c(1.5, .5, 0), lwd = 1.5)

plot(betas.sep1[, 1], type = "l", xlab = "Draws after burn-in", ylab = labels[1])
acf(betas.sep1[, 1], ylab = "empirical ACF")

plot(betas.sep1[, 2], type = "l", xlab = "Draws after burn-in", ylab = labels[2])
acf(betas.sep1[, 2], ylab = "empirical ACF")
```

Correspondingly the autocorrelation of the draws are much lower under the tighter prior. 


## Section 8.1.2: Logit model

### Example 8.7: Labor market data

We now estimate a logistic regression model for the labor market data
using the two-block Polya-Gamma sampler.

```{r}
logit <- function(y, X, b0 = 0, B0 = 10000,
                  burnin = 1000L, M = 5000L) {
    
    N <- length(y)
    d <- ncol(X) # number regression effects 

    b0 <- rep(b0, length.out = d) 
    B0.inv <- diag(rep(1 / B0, length.out = d), nrow = d)
    B0inv.b0 <- B0.inv %*% b0

    betas <- matrix(NA_real_, nrow = M, ncol = d)
    colnames(betas) <- colnames(X)

                                        # Define quantities for the Gibbs sampler
    ind0 <- (y == 0) # indicators for zeros
    ind1 <- (y == 1) # indicators for ones
    
                                        # Set starting values
    beta <- rep(0, d)
    z <- rep(NA_real_, N)
    omega <-rep(NA_real_, N)

    for (m in seq_len(burnin + M)) {
                                        # Draw z conditional on y and beta
        eta <- X %*% beta
        pi <- plogis(eta) 
        
        u <- runif(N)
        z[ind0] <- eta[ind0] + qlogis(u[ind0] * (1 - pi[ind0]))
        z[ind1] <- eta[ind1] + qlogis (1 - u[ind1] * pi[ind1])
        
                                        # Draw omega conditional on y, beta and z
        omega <- pgdraw::pgdraw(b = 1, c = z - eta)
        
                                        # Sample beta from the full conditional 
        Xomega <- matrix(omega, ncol = d, nrow = N) * X
        BN <- solve(B0.inv + crossprod(Xomega, X))
        bN <- BN %*% (B0inv.b0 + crossprod(Xomega, z))
        beta <- t(mvtnorm::rmvnorm(1, mean = bN, sigma = BN))
        
                                        # Store the beta draws
        if (m > burnin) {
            betas[m - burnin, ] <- beta
        }
    }
    return(betas)
}
```

We again use the Normal prior with mean $\mathbf{0}$ and covariance
matrix $10000 \mathbf{I}$ on the regression effects and estimate the
model. We summarize the posterior effect estimates and determine the
risk of unemployment for a baseline person using the fitted logit
model.

```{r}
set.seed(1234)
betas_logit <- logit(y.unemp, X.unemp, b0 = 0, B0 = 10000)

res_logit.labor <- t(apply(betas_logit, 2, res.mcmc))
knitr::kable(round(res_logit.labor, 3))

(p_unemploy_base_logit <- round(plogis(res_logit.labor[1, 2]), 4))

```

The risk of being unemployed 1998 for a male blue collar worker of age
18, who was employed 1997 is very low with a value of 
`r p_unemploy_base_logit` and the risk is even lower for a white collar
worker. It is higher for females, increases with age and is
particularly high for persons who were unemployed 1997.

While the signs of the covariate effects can be interpreted in the
same way for the probit and the logit model, their numerical value
will differ due to the different scale of the link function.
 
As the logistic distribution has a variance of $\pi^2/3$ compared to 1
for the standard Normal distribution, the regression effects in the
logit model are absolutely larger than those in the probit
model. However any probability computed from the two models will be
very close, e.g., the estimated probability to be unemployed for a
baseline person is `r p_unemploy_base_probit` in the probit model
(compared to `r p_unemploy_base_logit` in the logit model).

By multiplying the estimated coefficients in the probit model by
$\pi/\sqrt{3}$ we can compare them to the estimates of the logit model
and we see that there is not much difference.

```{r}
knitr::kable(round(res_probit.labor * pi / sqrt(3), 3))
```

# Section 8.2: Count response variables

## Section 8.2.1: Poisson regression models

### Example 8.8: Road safety data

We fit two different Poisson regression models to the series of
monthly deadly and seriously injured children aged 6-10 in Linz
introduced in Example 2.1:

1. a small model with intercept, intervention effect and holiday dummy
   (activated in July/August);

2. a larger model with intercept, intervention effect, 
   and a seasonal dummy variables for all months except for December.

The sampling performance for these two models is assessed to study how
the acceptance rate deteriorates, when the dimension of regression effects $d$ 
increases.

We load the data and extract the observations for children in Linz. 

```{r}
data("accidents", package = "BayesianLearningCode")
y <- accidents[, "children_accidents"]
e <- accidents[, "children_exposure"]
```

Then we define the regressor matrix for Model 1.

```{r}
X <- cbind(intercept = rep(1, length(y)),
           intervention = rep(c(0, 1), c(7 * 12 + 9, 8 * 12 + 3)),
           holiday = rep(rep(c(0, 1, 0), c(6, 2, 4)), 16))
```

To compute the parameters of the normal proposal density, we use the
Newton-Raphson estimator described in Section 8.2.1.

```{r}
gen.proposal.poisson <- function(y, X, e, b0 = 0, B0 = 100, t.max = 20) {
  N <- length(y)
  d <- ncol(X)
  
  betas <- matrix(NA_real_, ncol = t.max, nrow = d)
  beta.new <- matrix(c(log(mean(y)), rep(0, d - 1)), nrow = d) 
  
  b0 <- matrix(rep(b0, length.out = d), nrow = d) 
  B0.inv <- diag(rep(1 / B0, length.out = d), nrow = d)

  for (t in seq_len(t.max)) {
    beta.old <- beta.new
    
    rate <- e * exp(X %*% beta.old)
    score <- t(crossprod(y - rate, X) - t(beta.old - b0) %*% B0.inv)

    H <- -B0.inv
    for (i in seq_len(N)) {
      H <- H - rate[i] * tcrossprod(X[i, ])
    }
    beta.new <- beta.old - solve(H, score)
  }
  qmean <- beta.new
  
  # Determine the variance matrix  
  rate <- e * exp(X %*% qmean)
  H <- -B0.inv
  for (i in seq_len(N)) {
     H <- H - rate[i] * tcrossprod(X[i, ])
  }
  qvar <- -solve(H)
  return(parms.proposal = list(mean = qmean,
                               var = qvar))
}
```

We use a rather flat normal independence prior
$\mathcal{N}(\mathbf{0}, 100 \mathbf{I})$ on the regression
effects. First we determine the parameters of the proposal
distribution.

```{r}
parms.proposal <- gen.proposal.poisson(y, X, e, b0 = 0, B0 = 100)
parms.proposal
```

Next we set up the independence Metropolis-Hastings algorithm to
estimate the model parameters.

```{r}
poisson <- function(y, X, e, b0 = 0, B0 = 100, qmean, qvar,
                    burnin = 1000L, M = 10000L) {
  d <- ncol(X)

  b0 <- rep(b0, length.out = d)
  B0 <- diag(rep(B0, length.out = d), nrow = d)

  beta.post <- matrix(ncol = d, nrow = M)
  colnames(beta.post) <- colnames(X)
  acc <- numeric(length = M)
  
  beta <- as.vector(mvtnorm::rmvnorm(1, mean = qmean, sigma = qvar))
  
  for (m in seq_len(burnin + M)) {
    beta.old <- beta
    beta.proposed <- as.vector(mvtnorm::rmvnorm(1, mean = qmean, sigma = qvar))
    
    # Compute log proposal density at proposed and old value
    lq_proposed <- mvtnorm::dmvnorm(beta.proposed, mean = qmean, sigma = qvar,
                                    log = TRUE)
    lq_old  <- mvtnorm::dmvnorm(beta.old, mean = qmean, sigma = qvar,
                                log = TRUE)
    
    # Compute log prior of proposed and old value
    lpri_proposed <- mvtnorm::dmvnorm(beta.proposed, mean = b0, sigma = B0,
                                      log = TRUE)
    lpri_old  <- mvtnorm::dmvnorm(beta.old, mean = b0, sigma = B0,
                                  log = TRUE)
    
    # Compute log-likelihood of proposed and old value
    lh_proposed <- dpois(y, e * exp(X %*% beta.proposed), log = TRUE)
    lh_old  <- dpois(y, e * exp(X %*% beta.old), log = TRUE)
    
    maxlik <- max(lh_old, lh_proposed)
    ll <- sum(lh_proposed - maxlik) - sum(lh_old - maxlik)
    
    # Compute acceptance probability and accept or not
    log_acc <- min(0, ll + lpri_proposed - lpri_old + lq_old - lq_proposed)
    
    if (log(runif(1)) < log_acc) {
      beta <- beta.proposed
      accept <- 1
    } else {
      beta <- beta.old
      accept <- 0
    }

    # Store the beta draws
    if (m > burnin) {
        beta.post[m - burnin, ] <- beta
        acc[m - burnin] <- accept
    }
  }
  return(list(beta.post = beta.post, accept = mean(acc)))
}
```

We perform MCMC and report the results.

```{r}
set.seed(1234)
res1 <- poisson(y, X, e, b0 = 0, B0 = 100,
                qmean = parms.proposal$mean, qvar = parms.proposal$var)

res.poisson1 <- cbind(t(round(apply(res1$beta.post, 2, res.mcmc), 3)),
                      "exp(beta)" = round(exp(colMeans(res1$beta.post)), 5))
knitr::kable(res.poisson1)
```

We see that the risk for a child to be killed or seriously injured is
lower during holiday months as well as after the intervention.

```{r}
(base_risk <- res.poisson1[1, "exp(beta)"]*10^4)
res1$accept
```

The baseline risk is 2.7 per 10000 children months and it is reduced
to less than half during holiday months. After the intervention the
risk of being killed or seriously injured of a child in Linz is
reduced to 70% of the baseline risk.

Next, we fit the alternative model with intercept, intervention
effect, and seasonal dummy variables for all months except for
December. Hence the intercept models the risk in December before the
intervention.

```{r}
seas <- rbind(diag(1, 11), rep(0, 11)) 
seas.dummies <- matrix(rep(t(seas), 16), ncol = 11, byrow = TRUE)
colnames(seas.dummies) <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul",
                            "Aug", "Sep", "Oct", "Nov")
X.large <- cbind(X[, -3],
                 seas.dummies)
```

We set the prior parameters and compute the parameters of the proposal
distribution.

```{r}
parms.proposal2 <- gen.proposal.poisson(y, X.large, e, b0 = 0, B0 = 100)
```
Next we fit the model.

```{r}
set.seed(1234)
res2 <- poisson(y, X.large, e, b0 = 0, B0 = 100,
                qmean = parms.proposal2$mean, qvar = parms.proposal2$var)

res.poisson2 <- cbind(t(round(apply(res2$beta.post, 2, res.mcmc), 3)),
                      "exp(beta)" = round(exp(colMeans(res2$beta.post)), 5))
knitr::kable(res.poisson2)

(base_risk <- res.poisson2[1, "exp(beta)"]*10^4)
```

With 2.8 deadly or seriously injured children per 10000 at risk, the
estimated baseline risk is very similar to that from Model 1. Also the
estimated intervention effect is very similar in both models,
indicating a reduction of the risk by a factor of 0.69 in Model 2
(compared to 0.70 in Model 1). The monthly effects have rather wide
95\% HPD intervals that cover 0 for all months except for July and
August. For these two holiday months they are clearly negative,
indicating a considerable reduction of the risk.

```{r}
res2$accept
```

The acceptance rate is `r round(res1$accept, 2)` for the smaller Model
1 with three parameters, but only `r round(res2$accept, 2)` for Model
2, where `r ncol(X.large)` regression coefficient parameters have to
be estimated.

## Section 8.2.2: Negative binomial regression

### Example 8.9: Road safety data

Now we re-analyze the road safety data allowing for unobserved
heterogeneity. We first set up the two versions of the three-block
MH-within-Gibbs sampler.

Note that the negative binomial distribution in R is specified
as
$$
p(y|\alpha, p)={\alpha-1+y \choose \alpha-1} p^\alpha (1-p)^y
$$
or alternatively by the parameters $\alpha$ and its expected value $$\mu=\alpha (1-p)/p.$$  The expected value of 
$p(y|\alpha, \beta)$ is given as
$$
\mu=\alpha \frac{\frac{1}{1+\alpha/e \exp(- \mathbf{x}\boldsymbol{\beta})}}
{\frac{\alpha/e \exp(- \mathbf{x}\boldsymbol{\beta})}{1+\alpha/e \exp(- \mathbf{x}\boldsymbol{\beta})}} = e \exp( \mathbf{x}\boldsymbol{\beta}),
$$
and we will use $\alpha$ and $\mu$ to specify the negative binomial distribution.

Extra Functions for the sampling steps ?ß


```{r}
negbin <- function(y, X, e, b0 = 0, B0 = 100, qmean, qvar, pri.alpha,
                   full.gibbs = FALSE, burnin = 1000L, M = 50000L) {
  
  N <- length(y)
  d <- ncol(X)
  
  b0 <- rep(b0, length.out = d)
  B0 <- diag(rep(B0, length.out = d), nrow = d)

  beta.post <- matrix(ncol = d, nrow = M)
  colnames(beta.post) <- colnames(X)
  acc.beta <- numeric(length = M)
    
  alpha.post <- rep(NA_real_, M)
  acc.alpha <- rep(NA_real_, M)
  c_alpha <- 0.1
  
  # Set starting values
  beta <- as.vector(mvtnorm::rmvnorm(1, mean = qmean, sigma = qvar))
  alpha <- with(pri.alpha, shape / rate)
  phi <- rep(1, N)
  
  for (m in seq_len(burnin + M)){
     # Step (a): Draw beta  
     beta.old <- beta
     beta.proposed <- as.vector(mvtnorm::rmvnorm(1, mean = qmean, sigma = qvar))

     # Compute log proposal density at proposed and old value
     lq_proposed <- mvtnorm::dmvnorm(beta.proposed, mean = qmean, sigma = qvar, 
                                     log = TRUE)
     lq_old  <- mvtnorm::dmvnorm(beta.old, mean = qmean, sigma = qvar,
                                 log = TRUE)
            
     # Compute log prior of proposed and old value
     lpri_proposed <- mvtnorm::dmvnorm(beta.proposed, mean = b0, sigma = B0, 
                                       log = TRUE)
     lpri_old  <- mvtnorm::dmvnorm(beta.old,  mean = b0, sigma = B0, log = TRUE)

     # Compute log likelihood of proposed and old value
     lh_proposed <- dpois(y, e * exp(X %*% beta.proposed), log = TRUE)
     lh_old  <- dpois(y, e * exp(X %*% beta.old), log = TRUE)
    
     maxlik <- max(lh_old, lh_proposed)
     ll <- sum(lh_proposed - maxlik) - sum(lh_old - maxlik)
    
     # Compute acceptance probability and accept or not
     log_acc <- min(0, ll + lpri_proposed - lpri_old + lq_old - lq_proposed)

     if (log(runif(1)) < log_acc) {
       beta <- beta.proposed
       acc.b <- 1
     } else {
       beta <- beta.old
       acc.b <- 0
     }
     linpred <- X %*% beta

     # Step (b): Draw alpha
     alpha.old <- alpha
     alpha.proposed <- alpha.old * exp(c_alpha * rnorm(1))
     
     if (full.gibbs) {
       llik_alpha.proposed <- sum(dgamma(phi, shape = alpha.proposed,
                                         rate = alpha.proposed, log = TRUE))
       llik_alpha.old      <- sum(dgamma(phi, shape = alpha.old,
                                         rate = alpha.old, log = TRUE)) 
     } else {
       llik_alpha.proposed <- sum(dnbinom(y, size = alpha.proposed, 
                                          mu = e * exp(linpred), log = TRUE))
       llik_alpha.old      <- sum(dnbinom(y, size = alpha.old, 
                                          mu = e * exp(linpred), log = TRUE))
     }
     log_acc_alpha <- llik_alpha.proposed - llik_alpha.old +
         dgamma(alpha.proposed, shape = pri.alpha$shape,
                rate = pri.alpha$rate, log = TRUE) -
         dgamma(alpha.old, shape = pri.alpha$shape,
                rate = pri.alpha$rate, log = TRUE) +
         log(alpha.proposed) - log(alpha.old)

     if (log(runif(1)) < log_acc_alpha) {
       alpha <- alpha.proposed
       acc.a <- 1
     } else {
       alpha <- alpha.old
       acc.a <- 0
     }
   
    # Step (c) : Draw phi from its full conditional
    phi <- rgamma(N, shape = alpha + y, rate = alpha + e * exp(linpred))
    
    # Save the draws
    if (m > burnin) {
      beta.post[m - burnin, ] <- beta
      acc.beta[m - burnin] <- acc.b
        
      alpha.post[m - burnin] <- alpha
      acc.alpha[m - burnin] <- acc.a
    }
  }
  return(res = list(beta.post = beta.post, acc.beta = acc.beta,
                    alpha.post = alpha.post, acc.alpha = acc.alpha))
}
```

We use the same Normal prior as in the Poisson model for the
regression effects $\boldsymbol{\beta}$ and a Gamma prior
$\mathcal{G}(2 , 0.5)$ for $\alpha$ and run both samplers for
$M=50,000$ iterations after a burn-in of 1000.

```{r}
set.seed(1234)
pri.alpha <- data.frame(shape = 2, rate = 0.5)
M <- 50000L

# Full Gibbs sampler
res1 <- negbin(y, X, e, qmean = parms.proposal$mean, qvar = parms.proposal$var,
               pri.alpha = pri.alpha, full.gibbs = TRUE, M = M)

res.negbin.full <- rbind(t(apply(res1$beta.post, 2, res.mcmc)), 
                         alpha = res.mcmc(res1$alpha.post))

ESS.beta1 <- coda::effectiveSize(res1$beta.post)
ESS.alpha1 <- coda::effectiveSize(res1$alpha.post)
IF.res1 <-  M / c(ESS.beta1, ESS.alpha1)
res.negbin.full <- cbind(res.negbin.full, IF = IF.res1)
knitr::kable(round(res.negbin.full, 3))

c(mean(res1$acc.beta), mean(res1$acc.alpha))

# Partially marginalized sampler               
res2 <- negbin(y, X, e, qmean = parms.proposal$mean, qvar = parms.proposal$var,
               pri.alpha = pri.alpha, full.gibbs = FALSE, M = M)

res.negbin.partial <- rbind(t(apply(res2$beta.post, 2, res.mcmc)),
                            alpha = res.mcmc(res2$alpha.post))

ESS.beta2 <- coda::effectiveSize(res2$beta.post)
ESS.alpha2 <- coda::effectiveSize(res2$alpha.post)
IF.res2 <-  M / c(ESS.beta2, ESS.alpha2)
res.negbin.partial<- cbind(res.negbin.partial, IF = IF.res2)
knitr::kable(round(res.negbin.partial, 3))

c(mean(res2$acc.beta), mean(res2$acc.alpha))
```

Both samplers yield essentially the same estimation results, which is
to be expected, since both target the same posterior distribution.
The overdispersion parameter $\alpha$ has a posterior mean of $~12.3$,
which means that overdispersion is not very pronounced.

The two samplers differ, however, particularly w.r.t. the inefficiency
of $\alpha$ which has a value of 73.78 in the full sampler, but is
smaller with a value of 48.18 for the partially marginalized Gibbs
sampler.


## Section 8.2.3: Evaluating MCMC samplers
### Example 8.10 Veryfying the correctness of the full conditional MCMC samper

We extend the sampler in the scheme (a), (b), (c) by adding as a
further step sampling the data from the prior.

```{r}
negbin_check_abc <- function(X,e, b0 = 0, B0 = 100, qmean, qvar, pri.alpha,
                             full.gibbs = FALSE, burnin = 1000L, M = 50000L) {
  
  N <- nrow(X)
  d <- ncol(X)
  
  b0 <- rep(b0, length.out = d)
  B0 <- diag(rep(B0, length.out = d), nrow = d)    
  
  beta.post  <- matrix(ncol = d, nrow = M)
  colnames(beta.post) <- colnames(X)
  acc.beta <- numeric(length = M)
  
  alpha.post <- rep(NA_real_, M)
  acc.alpha <- rep(NA_real_, M)
  c_alpha <- 0.1
  
  # Set starting values
  phi <- rep(1, N)  
  beta <- as.vector(mvtnorm::rmvnorm(1, mean = qmean, sigma = qvar))
  alpha <- pri.alpha$shape/pri.alpha$rate
  
  for (m in seq_len(burnin + M)){
    # sample new data
    y <- rnbinom(N, size = alpha, mu = e * exp(X%*%beta))
    
    # Step (a): Draw beta  
    beta.old <- beta
    beta.proposed <- as.vector(mvtnorm::rmvnorm(1, mean = qmean, sigma = qvar))

    # Compute log proposal density at proposed and old value
    lq_proposed <- mvtnorm::dmvnorm(beta.proposed, mean = qmean, sigma = qvar, 
                                    log = TRUE)
    lq_old  <- mvtnorm::dmvnorm(beta.old, mean = qmean, sigma = qvar,
                                log = TRUE)
            
    # Compute log prior  of proposed and old value
    lpri_proposed <- mvtnorm::dmvnorm(beta.proposed, mean = b0, sigma = B0, 
                                      log = TRUE)
    lpri_old  <- mvtnorm::dmvnorm(beta.old,  mean = b0, sigma = B0, log = TRUE)
      
    # Compute log likelihood of proposed and old value
    lh_proposed <- dpois(y, e * exp(X %*% beta.proposed), log = TRUE)
    lh_old  <- dpois(y, e * exp(X %*% beta.old), log = TRUE)
    
    maxlik <- max(lh_old, lh_proposed)
    ll <- sum(lh_proposed - maxlik) - sum(lh_old - maxlik)
    
    # Compute acceptance probability and accept or not
    log_acc <- min(0, ll + lpri_proposed - lpri_old + lq_old - lq_proposed)

    if (log(runif(1)) < log_acc) {
      beta <- beta.proposed
      acc.b <- 1
    } else {
      beta <- beta.old
      acc.b <- 0
    }
    linpred <- X %*% beta

    # Step (b): Draw alpha
    alpha.old <- alpha
    alpha.proposed <- alpha.old * exp(c_alpha * rnorm(1))
     
    if (full.gibbs) {
      llik_alpha.proposed <- sum(dgamma(phi, shape = alpha.proposed,
                                        rate = alpha.proposed, log = TRUE))
      llik_alpha.old      <- sum(dgamma(phi, shape = alpha.old,
                                        rate = alpha.old, log = TRUE)) 
    } else {
      llik_alpha.proposed <- sum(dnbinom(y, size = alpha.proposed, 
                                         mu = e * exp(linpred), log = TRUE))
      llik_alpha.old      <- sum(dnbinom(y, size = alpha.old, 
                                         mu = e * exp(linpred), log = TRUE))
    }
    log_acc_alpha <- llik_alpha.proposed - llik_alpha.old +
        dgamma(alpha.proposed, shape = pri.alpha$shape,
               rate = pri.alpha$rate, log = TRUE) -
        dgamma(alpha.old, shape = pri.alpha$shape,
               rate = pri.alpha$rate, log = TRUE) +
        log(alpha.proposed) - log(alpha.old)

    if (log(runif(1)) < log_acc_alpha) {
      alpha <- alpha.proposed
      acc.a <- 1
    } else {
      alpha <- alpha.old
      acc.a <- 0
    }
   
    # Step (c) : Draw phi from its full conditional
    phi <- rgamma(N, shape = alpha + y, rate = alpha + e * exp(linpred))
    
    # Save the draws
    if (m > burnin) {
      beta.post[m - burnin, ] <- beta
      acc.beta[m - burnin] <- acc.b
        
      alpha.post[m - burnin] <- alpha
      acc.alpha[m - burnin] <- acc.a
    }
  }
  return(res = list(beta.post = beta.post, acc.beta = acc.beta,
                    alpha.post = alpha.post, acc.alpha = acc.alpha))
}
```

As the data are sampled from the prior distribution, we need a proper
prior. Moreover as we draw the regression effects from a tailored
Normal proposal, we avoid changing this proposal in each step and set
the prior mean to the mean of this proposal and for the prior variance
we use the diagonal elements of the proposal.  We generate also draws
from the prior distribution.

```{r}
pri.beta <- list(b0 = parms.proposal$mean, B0 = diag(parms.proposal$var))
```

We  then run the sampler and investigate the draws of intercept and heterogeneity
parameter via Q-Q plots of draws from the prior and the posterior.

```{r, echo = -c(1:3)}
if (pdfplots) {
  pdf("8-2_1.pdf", width = 8, height = 4)
}
par(mfrow = c(1, 2), mar = c(2.5, 2.5, 1.5, .1), mgp = c(1.5, .5, 0), lwd = 1.5)

set.seed(123)
res_check_abc <- negbin_check_abc(X, e, b0 = pri.beta$b0, B0 = pri.beta$B0, pri.alpha,
                                  qmean = parms.proposal$mean, qvar = parms.proposal$var,
                                  full.gibbs = TRUE, M = M)

beta0.prior <- rnorm(M, mean = pri.beta$b0[1], sd = sqrt(pri.beta$B0[1]))
alpha.prior <- rgamma(M, shape = pri.alpha$shape, rate = pri.alpha$rate)

qqplot(beta0.prior, res_check_abc$beta.post[, 1], xlab = "Prior",
       ylab = "Posterior", main = "Intercept")
abline(a = 0, b = 1)

qqplot(alpha.prior,res_check_abc$alpha.post, xlab = "Prior",
       ylab = "Posterior", main = "Heterogeneity parameter", 
       xlim = c(0, 30), ylim = c(0, 30))
abline(a = 0, b = 1)
```

We conclude that the sampler is correct.

We now change the order of the sampling steps to (c)-(b)-(a).

```{r}
negbin_check_cba <- function(X,e, b0 = 0, B0 = 100, qmean, qvar, pri.alpha,
                             full.gibbs = FALSE, burnin = 1000L, M = 50000L) {
  
  N <- nrow(X)
  d <- ncol(X)
  
  b0 <- rep(b0, length.out = d)
  B0 <- diag(rep(B0, length.out = d), nrow = d)
    
  beta.post  <- matrix(ncol = d, nrow = M)
  colnames(beta.post) <- colnames(X)
  acc.beta <- numeric(length = M)
  
  alpha.post <- rep(NA_real_, M)
  acc.alpha <- rep(NA_real_, M)
  c_alpha <- 0.1
  
  # Set starting values
  phi <- rep(1, N)  
  beta <- as.vector(mvtnorm::rmvnorm(1, mean = qmean, sigma = qvar))
  alpha <- pri.alpha$shape/pri.alpha$rate
  
  for (m in seq_len(burnin + M)){
    # sample new data
    linpred <- X %*% beta
    y <- rnbinom(N, size = alpha, mu = e * exp(linpred))
    
    # Step (c) : Draw phi from its full conditional
    phi <- rgamma(N, shape = alpha + y, rate = alpha + e * exp(linpred))
     
    # Step (b): Draw alpha
    alpha.old <- alpha
    alpha.proposed <- alpha.old * exp(c_alpha * rnorm(1))
     
    if (full.gibbs) {
      llik_alpha.proposed <- sum(dgamma(phi, shape = alpha.proposed,
                                        rate = alpha.proposed, log = TRUE))
      llik_alpha.old      <- sum(dgamma(phi, shape = alpha.old,
                                        rate = alpha.old, log = TRUE)) 
    } else {
      llik_alpha.proposed <- sum(dnbinom(y, size = alpha.proposed, 
                                         mu = e * exp(linpred), log = TRUE))
      llik_alpha.old      <- sum(dnbinom(y, size = alpha.old, 
                                         mu = e * exp(linpred), log = TRUE))
    }
    log_acc_alpha <- llik_alpha.proposed - llik_alpha.old +
        dgamma(alpha.proposed, shape = pri.alpha$shape,
               rate = pri.alpha$rate, log = TRUE) -
        dgamma(alpha.old, shape = pri.alpha$shape,
               rate = pri.alpha$rate, log = TRUE) +
                       log(alpha.proposed) - log(alpha.old)

    if (log(runif(1)) < log_acc_alpha) {
      alpha <- alpha.proposed
      acc.a <- 1
    } else {
      alpha <- alpha.old
      acc.a <- 0
    }
   
    # Step (a): Draw beta  
    beta.old <- beta
    beta.proposed <- as.vector(mvtnorm::rmvnorm(1, mean = qmean, sigma = qvar))

    # Compute log proposal density at proposed and old value
    lq_proposed <- mvtnorm::dmvnorm(beta.proposed, mean = qmean, sigma = qvar, 
                                    log = TRUE)
    lq_old  <- mvtnorm::dmvnorm(beta.old, mean = qmean, sigma = qvar,
                                log = TRUE)
            
    # Compute log prior  of proposed and old value
    lpri_proposed <- mvtnorm::dmvnorm(beta.proposed, mean = b0, sigma = B0, 
                                      log = TRUE)
    lpri_old  <- mvtnorm::dmvnorm(beta.old,  mean = b0, sigma = B0, log = TRUE)

    # Compute log likelihood of proposed and old value
    lh_proposed <- dpois(y, e * exp(X %*% beta.proposed), log = TRUE)
    lh_old  <- dpois(y, e * exp(X %*% beta.old), log = TRUE)
    
    maxlik <- max(lh_old, lh_proposed)
    ll <- sum(lh_proposed - maxlik) - sum(lh_old - maxlik)
    
    # Compute acceptance probability and accept or not
    log_acc <- min(0, ll + lpri_proposed - lpri_old + lq_old - lq_proposed)

    if (log(runif(1)) < log_acc) {
      beta <- beta.proposed
      acc.b <- 1
    } else {
      beta <- beta.old
      acc.b <- 0
    }
     
    # Save the draws
    if (m > burnin) {
      beta.post[m - burnin, ] <- beta
      acc.beta[m - burnin] <- acc.b
        
      alpha.post[m - burnin] <- alpha
      acc.alpha[m - burnin] <- acc.a
    }
  }
  return(res = list(beta.post = beta.post, acc.beta = acc.beta,
                    alpha.post = alpha.post,acc.alpha = acc.alpha))
}
```

We run the sampler under this scheme and show the Q-Q plots for the
intercept and the heterogeneity parameter.

```{r, echo = -c(1:3)}
if (pdfplots) {
  pdf("8-2_2.pdf", width = 8, height = 4)
}
par(mfrow = c(1, 2), mar = c(2.5, 2.5, 1.5, .1), mgp = c(1.5, .5, 0), lwd = 1.5)

set.seed(123)
res_check_cba <- negbin_check_cba(X, e, b0 = pri.beta$b0, B0 = pri.beta$B0, pri.alpha,
                                  qmean = parms.proposal$mean, qvar = parms.proposal$var,
                                  full.gibbs = TRUE, M = M)

qqplot(beta0.prior, res_check_cba$beta.post[, 1], xlab = "Prior",
       ylab = "Posterior", main = "Intercept")
abline(a = 0, b = 1)

qqplot(alpha.prior,res_check_cba$alpha.post, xlab = "Prior",
       ylab = "Posterior", main =  "Heterogeneity parameter",
       xlim = c(0, 30), ylim = c(0, 30))
abline(a = 0, b = 1)
```

## Example 8.11

We now analyze the partial marginalized Gibbs sampler, first in the order (a)-(b)-(c) 

```{r, echo = -c(1:3)}
if (pdfplots) {
  pdf("8-2_3.pdf", width = 8, height = 4)
}
par(mfrow = c(1, 2), mar = c(2.5, 2.5, 1.5, .1), mgp = c(1.5, .5, 0), lwd = 1.5)

set.seed(123)
# order (a)-(b)-(c)
res_check_abc <- negbin_check_abc(X, e, b0 = pri.beta$b0, B0 = pri.beta$B0, pri.alpha,
                                 qmean = parms.proposal$mean, qvar = parms.proposal$var,
                                 full.gibbs = FALSE, M = M)

qqplot(beta0.prior, res_check_abc$beta.post[, 1], xlab = "Prior",
       ylab = "Posterior", main = "Intercept")
abline(a = 0, b = 1)

qqplot(alpha.prior,res_check_abc$alpha.post, xlab = "Prior", 
       ylab = "Posterior", main  = "Heterogeneity parameter", 
       xlim = c(0, 30), ylim = c(0, 30))
abline(a = 0, b = 1)
```

and then in the order (c)-(b)-(a)

```{r, echo = -c(1:3)}
set.seed(1)
if (pdfplots) {
  pdf("8-2_4.pdf", width = 8, height = 4)
}
par(mfrow = c(1, 2), mar = c(2.5, 2.5, 1.5, .1), mgp = c(1.5, .5, 0), lwd = 1.5)

set.seed(123)
# order (c)- (b)-(a)
res_check_cba <- negbin_check_cba(X, e, b0 = pri.beta$b0, B0 = pri.beta$B0, pri.alpha,
                                  qmean = parms.proposal$mean, qvar = parms.proposal$var,
                                  full.gibbs = FALSE, M = M)

qqplot(beta0.prior, res_check_cba$beta.post[, 1], xlab = "Prior",
       ylab = "Posterior", main = "Intercept")
abline(a = 0, b = 1)

qqplot(alpha.prior,res_check_cba$alpha.post, xlab = "Prior", 
       ylab = "Posterior", main = "Heterogeneity parameter",
       xlim = c(0, 30), ylim = c(0, 30))
abline(a = 0, b = 1)
```

# Section 8.3: Beyond i.i.d. Gaussian error distributions

## Section 8.3.1: Regression analysis with heteroskedastic errors

### Example 8.12: Star cluster data

The bivariate data set of the star cluster CYG OB1 is available in
package *robustbase* and we load it from this package and visualize it
in a scatter plot:

```{r}
data("starsCYG", package = "robustbase")
plot(starsCYG, pch = 19, xlim = c(3, 5), ylim = c(3, 7),
     xlab = "log temperature", ylab = "log light intensity")
```

The four giant stars which can also be identified in the scatter plot
have the following indices in the data set:

```{r}
index <- c(11, 20, 30, 34)
```

We fit a standard Bayesian regression analysis under the improper
prior $p(\beta_0, \beta_1, \sigma^2) \propto 1 / \sigma^2$ and
determine the mean and pointwise 95%-HPD regions of the posterior
predictive distribution $p(y_i |x_i = x, \mathbf{y})$ using (a) the
full data set and (b) the data set where the observations $y_{11}$,
$y_{20}$, $y_{30}$, $y_{34}$ corresponding to the giant stars are
omitted.

The posterior predictive distribution for a single observation $i$
with covariate value $x_i$ given the sample $\mathbf{y}$ with model
matrix $\mathbf{X}$ is available in closed form when using the
improper prior and corresponds to the prediction intervals obtained
using OLS estimation.
```{r}
ols_all <- lm(log.light ~ log.Te, data = starsCYG)
xnew <- seq(3, 5, length.out = 100)
preds_all <- predict(ols_all, newdata = data.frame(log.Te = xnew),
                     interval = "prediction")
ols_subset <- lm(log.light ~ log.Te, data = starsCYG[-index, ])
preds_subset <- predict(ols_subset, newdata = data.frame(log.Te = xnew),
                        interval = "prediction")
```

We compare the expected values (full lines) and
the pointwise 95%-HPD regions in the following figure for the model
fit using all data (left) and only the subset without the giant stars
(right).

## Figure 8.9: Star cluster data
```{r, echo = -c(1:2)}
if (pdfplots) {
  pdf("8-3_1.pdf", width = 8, height = 4)
}
par(mfrow = c(1, 2), mar = c(2.5, 2.5, 1.5, .1), mgp = c(1.5, .5, 0), lwd = 1.5)
plot(starsCYG, pch = 19, xlim = c(3, 5), ylim = c(3, 7),
     xlab = "log temperature", ylab = "log light intensity")
lines(xnew, preds_all[, "fit"])
lines(xnew, preds_all[, "lwr"], lty = 2)
lines(xnew, preds_all[, "upr"], lty = 2)
plot(starsCYG, pch = 19, xlim = c(3, 5), ylim = c(3, 7),
     xlab = "log temperature", ylab = "log light intensity")
lines(xnew, preds_subset[, "fit"])
lines(xnew, preds_subset[, "lwr"], lty = 2)
lines(xnew, preds_subset[, "upr"], lty = 2)
```

### Example 8.13: Star cluster data - heteroskedastic regression analysis with known outliers

We define the binary indicator indicating outlying observations, i.e.,
in this case observations corresponding to giant stars.

```{r}
S <- rep(1L, nrow(starsCYG))
S[index] <- 2L
```

We prepare the model matrix and the vector of the response and define
the dimensions.

```{r}
X <- cbind(1, starsCYG$log.Te)
y <- starsCYG$log.light
N <- length(y)
d <- ncol(X)
```

For the heteroskedastic regression, we define weights depending on the
binary indicator which are either 1 or equal to $\phi \ll 1$.

```{r}
phi <- 0.001
w <- phi^(S - 1)
```

We include the weights and update the Gibbs sampling scheme for the
linear regression defined in Chapter 6 accordingly.

```{r}
set.seed(1)

# define prior parameters of semi-conjugate prior
B0.inv <- diag(rep(1 / 10000, d), nrow = d)
b0 <- rep(0, d)

c0 <- 2.5
C0 <- 1.5

# define quantities for the Gibbs sampler taking the weights into
# account
Xtilde <- sqrt(w) * X
ytilde <- sqrt(w) * y
wXX <- crossprod(Xtilde)
wXy <- t(Xtilde) %*% ytilde
cN <- c0 + N / 2

# define burnin and M
burnin <- 1000
M <- 20000

# prepare storing of results
betas <- matrix(NA_real_, nrow = M, ncol = d)
sigma2s <- rep(NA_real_, M)
colnames(betas) <- colnames(X)

# starting value for sigma2
sigma2 <- var(y) / 2

for (m in 1:(burnin + M)) {
    # sample beta from the full conditional
    BN <- solve(B0.inv + wXX / sigma2)
    bN <- BN %*% (B0.inv %*% b0 + wXy / sigma2)
    beta <- t(mvtnorm::rmvnorm(1, mean = bN, sigma = BN))

    # sample sigma^2 from its full conditional
    eps <- ytilde - Xtilde %*% beta
    CN <- C0 + crossprod(eps) / 2
    sigma2 <- rinvgamma(1, cN, CN)

    if (m > burnin) {
        betas[m - burnin, ] <- beta
        sigma2s[m - burnin] <- sigma2
    }
}
```

Based on the posterior draws of the parameters we determine draws from
the predictive distributions for new observations with `xnew` values:

```{r}
pred_hetero <- sapply(1:M, function(m) {
    rnorm(length(xnew), cbind(1, xnew) %*% betas[m, ], sqrt(sigma2s[m]))
})
plot(starsCYG, pch = 19, xlim = c(3, 5), ylim = c(3, 7),
     xlab = "log temperature", ylab = "log light intensity")
lines(xnew, rowMeans(pred_hetero))
lines(xnew, apply(pred_hetero, 1, quantile, 0.025), lty = 2)
lines(xnew, apply(pred_hetero, 1, quantile, 0.975), lty = 2)
```

## Section 8.3.2 Regression analysis with errors following a Gaussian mixture

### Example 8.14: Star cluster data - regression analysis with Gaussian two-component mixture errors

We now assume that the indices of the giant stars are not known. We
only assume that a two-component mixture is used as weight
distribution where the weights are given by $w_i = \phi^{S_i-1}$ and
the indicators $S_i$ are unknown. But we assume that the proportion of
giant stars is known and we assume this to correspond to 10\%.

```{r}
eta <- 0.1
phi <- 0.001
```

We use a more informative prior for the semi-conjugate prior on the
regression parameters.

```{r}
B0.inv <- diag(rep(1, d), nrow = d)
b0 <- coef(ols_subset)
```

We now modify the Gibbs sampling code to include a sampling step for
the mixture component indicators.

```{r}
set.seed(1)

# starting values for beta and sigma2
beta <- coef(ols_subset)
sigma2 <- var(y) / 2

for (m in seq_len(burnin + M)) {
    # mixture component indicator
    Xbeta <- X %*% beta
    posterior <- cbind((1 - eta) * dnorm(y, Xbeta, sqrt(sigma2)),
                       eta * dnorm(y, Xbeta, sqrt(sigma2 / phi)))
    posterior <- posterior / rowSums(posterior)
    S <- 1 + rbinom(nrow(posterior), prob = posterior[, 2], size = 1)

    # re-weight
    w <- phi^(S - 1)
    Xtilde <- sqrt(w) * X
    ytilde <- sqrt(w) * y
    wXX <- crossprod(Xtilde)
    wXy <- t(Xtilde) %*% ytilde

    # sample beta from the full conditional
    BN <- solve(B0.inv + wXX / sigma2)
    bN <- BN %*% (B0.inv %*% b0 + wXy / sigma2)
    beta <- t(mvtnorm::rmvnorm(1, mean = bN, sigma = BN))

    # sample sigma^2 from its full conditional
    eps <- ytilde - Xtilde %*% beta
    CN <- C0 + crossprod(eps) / 2
    sigma2 <- rinvgamma(1, cN, CN)

    if (m > burnin) {
        betas[m - burnin, ] <- beta
        sigma2s[m - burnin] <- sigma2
    }
}
```

```{r}
preds_mix_1 <- sapply(1:M, function(m) {
    rnorm(length(xnew), cbind(1, xnew) %*% betas[m, ], sqrt(sigma2s[m]))
})
```

We visualize again the mean and the 95%-HPD region together with the
data points and show that the fit now is robust to the outlying
observations.

```{r}
plot(starsCYG, pch = 19, xlim = c(3, 5), ylim = c(3, 7),
     xlab = "log temperature", ylab = "log light intensity")
lines(xnew, rowMeans(preds_mix_1))
lines(xnew, apply(preds_mix_1, 1, quantile, 0.025), lty = 2)
lines(xnew, apply(preds_mix_1, 1, quantile, 0.975), lty = 2)
```

We now assume that the indices of the giant stars are not known. We
only assume that a two-component mixture is used as weight
distribution where the weights are given by $w_i = \phi^{S_i-1}$ and
the indicators $S_i$ are unknown. Now, we also assume that the proportion of
giant stars is unknown.

We need to specify a prior for $\eta$. The usual prior is a beta prior
and we will use a prior which has mean 0.1 and corresponds to a prior
sample size of 10.

```{r}
a0 <- 1
d0 <- 9
```

We now modify the Gibbs sampling code to also include a sampling step
for the component size.

```{r}
set.seed(1)

# prepare storing of results
etas <- rep(NA_real_, M)

# starting values for eta, beta and sigma2
eta <- 0.1
beta <- coef(ols_subset)
sigma2 <- var(y) / 2

for (m in seq_len(burnin + M)) {
    # mixture component indicator
    Xbeta <- X %*% beta
    posterior <- cbind((1 - eta) * dnorm(y, Xbeta, sqrt(sigma2)),
                       eta * dnorm(y, Xbeta, sqrt(sigma2 / phi)))
    posterior <- posterior / rowSums(posterior)
    S <- 1 + rbinom(nrow(posterior), prob = posterior[, 2], size = 1)

    # sample eta
    aN <- a0 + sum(S == 2)
    dN <- d0 + sum(S == 1)
    eta <- rbeta(1, aN, dN)

    # re-weight
    w <- phi^(S - 1)
    Xtilde <- sqrt(w) * X
    ytilde <- sqrt(w) * y
    wXX <- crossprod(Xtilde)
    wXy <- t(Xtilde) %*% ytilde

    # sample beta from the full conditional
    BN <- solve(B0.inv + wXX / sigma2)
    bN <- BN %*% (B0.inv %*% b0 + wXy / sigma2)
    beta <- t(mvtnorm::rmvnorm(1, mean = bN, sigma = BN))

    # sample sigma^2 from its full conditional
    eps <- ytilde - Xtilde %*% beta
    CN <- C0 + crossprod(eps) / 2
    sigma2 <- rinvgamma(1, cN, CN)

    if (m > burnin) {
        betas[m - burnin, ] <- beta
        sigma2s[m - burnin] <- sigma2
        etas[m - burnin] <- eta
    }
}
```

```{r}
preds_mix_2 <- sapply(1:M, function(m) {
    rnorm(length(xnew), cbind(1, xnew) %*% betas[m, ], sqrt(sigma2s[m]))
})
plot(starsCYG, pch = 19, xlim = c(3, 5), ylim = c(3, 7),
     xlab = "log temperature", ylab = "log light intensity")
lines(xnew, rowMeans(preds_mix_2))
lines(xnew, apply(preds_mix_2, 1, quantile, 0.025), lty = 2)
lines(xnew, apply(preds_mix_2, 1, quantile, 0.975), lty = 2)
```

Finally, we visualize again the mean and the 95%-HPD region together
with the data points for the three modeling approaches: (1) a
heteroskedastic regression analysis with known outliers, (2) a
Gaussian two-component mixture with known component sizes and (3) a
Gaussian two-component mixture where the component size is unknown.


```{r, echo = -c(1:2)}
if (pdfplots) {
  pdf("8-3_2.pdf", width = 8, height = 3.5)
}
par(mfrow = c(1, 3), mar = c(2.5, 2.5, 1.5, .1), mgp = c(1.5, .5, 0), lwd = 1.5)
plot(starsCYG, pch = 19, xlim = c(3, 5), ylim = c(3, 7),
     xlab = "log temperature", ylab = "log light intensity")
lines(xnew, rowMeans(pred_hetero))
lines(xnew, apply(pred_hetero, 1, quantile, 0.025), lty = 2)
lines(xnew, apply(pred_hetero, 1, quantile, 0.975), lty = 2)
plot(starsCYG, pch = 19, xlim = c(3, 5), ylim = c(3, 7),
     xlab = "log temperature", ylab = "log light intensity")
lines(xnew, rowMeans(preds_mix_1))
lines(xnew, apply(preds_mix_1, 1, quantile, 0.025), lty = 2)
lines(xnew, apply(preds_mix_1, 1, quantile, 0.975), lty = 2)
plot(starsCYG, pch = 19, xlim = c(3, 5), ylim = c(3, 7),
     xlab = "log temperature", ylab = "log light intensity")
lines(xnew, rowMeans(preds_mix_2))
lines(xnew, apply(preds_mix_2, 1, quantile, 0.025), lty = 2)
lines(xnew, apply(preds_mix_2, 1, quantile, 0.975), lty = 2)
```

The plot indicates that all three modeling approaches result in a fit
that is robust to the outlying observations.

## Section 8.3.3 Regression analysis with Student-t errors

### Example 8.15: Star cluster data - Student-$t$ regression analysis

Now we analyze the Star cluster data assuming a Student-$t$
distribution for the errors. We fix the degrees of freedom to a low
value between 4 and 10, i.e., setting it to 7.
```{r}
nu <- 7
```

We perform the Gibbs sampling scheme with data augmentation as outlined
in Algorithm 8.3. We use the same prior specification for the
regression parameters as used for the mixture distribution.

```{r}
set.seed(1)

# prepare storing of results
ws <- matrix(NA_real_, nrow = M, ncol = N)

# starting value for sigma2 and w
sigma2 <- var(y) / 2
w <- rep(1, N)

for (m in 1:(burnin + M)) {
    # define quantities for the Gibbs sampler taking the weights into
    # account
    Xtilde <- sqrt(w) * X
    ytilde <- sqrt(w) * y
    wXX <- crossprod(Xtilde)
    wXy <- t(Xtilde) %*% ytilde

    # sample beta from the full conditional
    BN <- solve(B0.inv + wXX / sigma2)
    bN <- BN %*% (B0.inv %*% b0 + wXy / sigma2)
    beta <- t(mvtnorm::rmvnorm(1, mean = bN, sigma = BN))

    # sample sigma^2 from its full conditional
    eps <- ytilde - Xtilde %*% beta
    CN <- C0 + crossprod(eps) / 2
    sigma2 <- rinvgamma(1, cN, CN)

    # sample w
    r <- eps^2 / (w * sigma2)
    w <- rgamma(length(eps), (nu + 1) /2, (nu + r) / 2)

    if (m > burnin) {
        betas[m - burnin, ] <- beta
        sigma2s[m - burnin] <- sigma2
        ws[m - burnin, ] <- w
    }
}
```

```{r, echo = -c(1:3)}
if (pdfplots) {
  pdf("8-3_3.pdf", width = 8, height = 4)
}
par(mfrow = c(1, 2), mar = c(2.5, 2.5, 1.5, .1), mgp = c(1.5, .5, 0), lwd = 1.5)

preds_stud <- sapply(1:M, function(m) {
    rstudt(length(xnew), cbind(1, xnew) %*% betas[m, ], sqrt(sigma2s[m]), nu)
})
preds_norm <- sapply(1:M, function(m) {
    rnorm(length(xnew), cbind(1, xnew) %*% betas[m, ], sqrt(sigma2s[m]))
})

plot(starsCYG, pch = 19, xlim = c(3, 5), ylim = c(3, 7),
     xlab = "log temperature", ylab = "log light intensity")
lines(xnew, rowMeans(preds_stud))
lines(xnew, apply(preds_stud, 1, quantile, 0.025), lty = 2)
lines(xnew, apply(preds_stud, 1, quantile, 0.975), lty = 2)
lines(xnew, apply(preds_norm, 1, quantile, 0.025), lty = 3)
lines(xnew, apply(preds_norm, 1, quantile, 0.975), lty = 3)
boxplot(ws, col = 2 * (1:ncol(ws) %in% index))
```

### Example 8.16: CHF exchange rate data - Fitting a Student-$t$ with $\nu$ unknown 

We begin with loading the data.

```{r}
data("exrates", package = "stochvol")
y <- 100 * diff(log(exrates$USD / exrates$CHF))
X <- 1
N <- length(y)
```

Next we perform the Gibbs sampling scheme as above and
include sampling the degrees of freedom parameter $\nu$ via a log random walk
MH step. Note that we do not have any predictors in this case, so the design
matrix is simply a vector of ones.

```{r}
set.seed(1)

# fix tuning parameter for MH, the number of draws M, and burn-in length
cnu <- 0.3
M <- 5000
burnin <- 1000

# fix prior hyperparameters
B0.inv <- 1
b0 <- 0
c0 <- 2.5
C0 <- 1.5
lambda <- 1 / 7

# set up the design matrix
X <- matrix(1, nrow = N, ncol = 1)

# allocate space for storing the draws
nus <- mus <- sigma2s <- rep(NA_real_, M)
ws <- matrix(NA_real_, nrow = M, ncol = N)

# starting value for sigma2, log(nu), and w
sigma2 <- var(y) / 2
w <- rep(1, N)
nu <- 10

# pre-compute cN
cN <- c0 + N / 2

# log likelihood
loglik <- function(y, X, beta, sigma2, nu) {
  length(y) * (lgamma((nu + 1) / 2) - lgamma(nu / 2) - .5 * log(nu)) -
  (nu + 1) / 2 * sum(log(1 + (y - X %*% beta)^2 / nu / sigma2))
}

accepts <- 0L
for (m in 1:(burnin + M)) {
  Xtilde <- sqrt(w) * X
  ytilde <- sqrt(w) * y
  wXX <- crossprod(Xtilde)
  wXy <- crossprod(Xtilde, ytilde)

  # sample beta from the full conditional (M-DA)
  BN <- solve(B0.inv + wXX / sigma2)
  bN <- BN %*% (B0.inv %*% b0 + wXy / sigma2)
  beta <- t(mvtnorm::rmvnorm(1, mean = bN, sigma = BN))

  # sample sigma^2 from its full conditional (S-DA)
  eps <- ytilde - Xtilde %*% beta
  CN <- C0 + crossprod(eps) / 2
  sigma2 <- rinvgamma(1, cN, CN)

  # sample nu (N-MH)
  nuprop <- exp(rnorm(1, log(nu), cnu))
  logR <- loglik(y, X, beta, sigma2, nuprop) -
          loglik(y, X, beta, sigma2, nu) +
          dexp(nuprop, lambda, log = TRUE) -
          dexp(nu, lambda, log = TRUE) +
          log(nuprop) -
          log(nu)
  
  if (log(runif(1)) < logR) {
    nu <- nuprop
    if (m > burnin) accepts <- accepts + 1L
  }
  
  # sample w (W-DA)
  r <- eps^2 / (w * sigma2)
  w <- rgamma(length(eps), (nu + 1) /2, (nu + r) / 2)

  # store the results
  if (m > burnin) {
    mus[m - burnin] <- beta
    sigma2s[m - burnin] <- sigma2
    nus[m - burnin] <- nu
    ws[m - burnin, ] <- w
  }
}
```

We visualize some results.

```{r}
par(mfrow = c(2, 2), mgp = c(1.5, .5, 0), mar = c(2.5, 2.5, 1, .5))
ts.plot(mus, xlab = "Iteration", ylab = expression(mu), main = "Trace plot")
ts.plot(sqrt(sigma2s), xlab = "Iteration", ylab = expression(sigma))
ts.plot(nus, xlab = "Iteration", ylab = expression(nu))
mtext(paste0("Acceptance rate: ", round(100 * accepts / M), "%"), line = -1.1)
selecta <- sample.int(N, 1)
ts.plot(ws[, selecta], xlab = "Iteration", ylab = bquote(~omega[.(selecta)]))
```

```{r, echo = -c(1:2)}
if (pdfplots) {
  pdf("8-3_4.pdf", width = 6, height = 3)
}
par(mfrow = c(1, 3), mar = c(2.5, 2.5, 1.5, .5), mgp = c(1.5, .5, 0), lwd = 1.5)
grid <- seq(0, 20, by = .1)
hist(nus, breaks = 20, freq = FALSE, xlab = bquote(nu),
     main = "Histogram")
lines(grid, dexp(grid, lambda), col = 2, lty = 2)
ts.plot(nus, xlab = "Iteration", ylab = expression(nu))
title(paste0("Trace plot (AR: ", round(100 * accepts / M), "%)"))
acf(nus)
IF <- M / coda::effectiveSize(nus)
title(paste0("Empirical ACF (IF: ", round(IF), ")"))
```

## Section 8.3.4 Regression analysis with autocorrelated errors

We begin by loading the data and visualizing them as a time series plot.

```{r}
data(newcars, package = "BayesianLearningCode")
plot(newcars)
```

Seasonal patterns are evident, as is a trend. To model these, we set up an
appropriate design matrix. Leveraging the fact the the data is a `ts` object, we
can do this very easily using `time` and `cycle`. 
Note that `model.matrix` will, by default, use the first month
(January) as a baseline and automatically include an intercept.

```{r}
degree <- 4
tim <- time(newcars)
month <- factor(cycle(tim))
X <- model.matrix(~ month + poly(tim, degree))
head(X)
d <- ncol(X)
```

Before including an autoregressive parameter, we first fit a standard
regression model, where we assume that the residuals are uncorrelated.

```{r}
y <- newcars
# Define the burnin and the number of draws thereafter
burnin <- 100
M <- 1000

# Define the prior parameters
b0 <- rep(0, d)
B0inv <- diag(0, d)
c0 <- 0
C0 <- 0

# Pre-compute some quantities for the Gibbs sampler
XX <- crossprod(X)
Xy <- crossprod(X, y)
cN <- c0 + length(y) / 2
    
# Allocate space to store the results
betas <- matrix(NA_real_, nrow = M, ncol = d)
colnames(betas) <- colnames(X)
sigma2s <- rep(NA_real_,  M)

# Set the starting values
sigma2 <- var(y) / 2
   
for (m in 1:(burnin + M)) {
  # Sample beta from its full conditional
  BN <- solve(B0inv + XX / sigma2) 
  bN <- BN %*% (B0inv %*% b0 + Xy / sigma2)
  beta <- t(mvtnorm::rmvnorm(1, mean = bN, sigma = BN))
 
  # Sample sigma^2 from its full conditional
  eps <- y - X %*% beta
  CN <- C0 + crossprod(eps) / 2
  sigma2 <- rinvgamma(1, cN, CN)

  # Store the results
  if (m > burnin) {
    betas[m - burnin, ] <- beta
    sigma2s[m - burnin] <- sigma2
  }
}
colMeans(betas)
plot(y)
points(tim, colMeans(ypred), type = "l", col = 2)
```

We can compute draws of the predicted values and the residuals.

```{r}
ypreds <- tcrossprod(X, betas)
resids <- as.numeric(y) - ypred
```

We proceed with visualizing the data and the posterior mean of the predicted
values as well as the posterior mean of the residuals.

```{r, echo = -c(1:2)}
if (pdfplots) {
  pdf("8-3_5.pdf", width = 6, height = 3)
}
par(mfrow = c(2, 1), mar = c(2.5, 2.5, 1.5, .5), mgp = c(1.5, .5, 0), lwd = 1.5)
plot(y, ylab = "", main = "New car regristrations")
points(tim, rowMeans(ypreds), col = 2, type = 'l')
legend("topright", c("Data", "Posterior mean"), lty = 1, col = 1:2)
plot(tim, rowMeans(resids), type = 'l', main = "Mean residuals", xlab = "Time",
     ylab = "")
abline(h = 0, lty = 3)
```

Apart from some outliers (the most prominent ones being related to the
COVID-outbreak), we still see autocorrelation in the residuals. Thus, we move
forward by including an autoregressive coefficient.



