---
title: "Chapter 10: Bayesian Model Selection"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Chapter 10: Bayesian Model Selection}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6
)
knitr::opts_knit$set(global.par = TRUE)
pdfplots <- FALSE # default: FALSE; set this to TRUE only if you like pdf figures
```

# Section 10.1: The Foundations of Bayesian Model Selections
## Table 10.1: (Log) Bayes factor and model probabilities

```{r}
logBF <- -7:7
BF <- exp(logBF)
PrM1 <- BF / (1 + BF)
PrM2 <- 1 - PrM1
#xtable::xtable(cbind(BF, logBF, PrM1, PrM2), digits = c(0, 3, 1, 3, 3))
knitr::kable(cbind(BF, logBF, PrM1, PrM2))
```

# Section 10.2: Bayesian Testing of Hypotheses
## Example 10.5: Testing for zero mean in the CHF-USD log returns

Before producing Figure 10.1, we begin with a concrete example.
We load the data, compute the required parameters, and evaluate the two 
marginal likelihoods.

```{r}
data(exrates, package = "stochvol")
y <- 100 * diff(log(exrates$USD / exrates$CHF))

c0 <- 1
C0 <- 1
N0 <- 10^(-5:3)

N <- length(y)
cN <- c0 + N  / 2
CN_M1 <- C0 + sum(y^2) / 2
CN_M2 <- C0 + 0.5 * N * (mean((y - mean(y))^2) + N0 / (N0 + N) * mean(y)^2)

(logmarglikM1 <- lgamma(cN) + c0 * log(C0) -
   lgamma(c0) - cN * log(CN_M1) - 0.5 * N * log(2 * pi))
(logmarglikM2 <- lgamma(cN) + c0 * log(C0) + 0.5 * log(N0) -
   lgamma(c0) - cN * log(CN_M2) - 0.5 * N * log(2 * pi) - 0.5 * log(N0 + N))
```

The "direct" formula for the log Bayes factor is even simpler, and we can check
whether we get the same answer.

```{r}
(logBF <- 0.5 * (log(N0 + N) - log(N0)) + cN * (log(CN_M2) - log(CN_M1)))
all.equal(logBF, logmarglikM1 - logmarglikM2)
```

Now we are ready to investigate the sensitivity of the log Bayes factor with
respect to prior hyperparameter choices.

```{r, echo = -c(1:2)}
if (pdfplots) {
  pdf("10-1_1.pdf", width = 8, height = 4)
  par(mar = c(2.5, 2.5, 1.5, .5), mgp = c(1.6, .6, 0))
}
c0 <- c(-1, 0, 0.001, 1, 100)
C0 <- c(0, 0, 0.001, 1, 100)
N0 <- 10^seq(1, 10, by = 0.1)

logBF <- matrix(NA_real_, nrow = length(N0), ncol = length(c0))
for (i in seq_along(c0)) {
  cN <- c0[i] + N  / 2
  CN_M1 <- C0[i] + sum(y^2) / 2
  CN_M2 <- C0[i] + 0.5 * N * (mean((y - mean(y))^2) + N0 / (N0 + N) * mean(y)^2)
  logBF[, i] <- 0.5 * (log(N0 + N) - log(N0)) + cN * (log(CN_M2) - log(CN_M1))
}
matplot(N0, logBF, log = "x", type = "l",
        xlab = expression(paste(N[0], " [log scale]")), ylab = "log BF")
abline(h = 0, lty = 3)
legend("topright", paste0("IG(", c0, ",", C0, ")"), col = seq_along(c0),
       lty = seq_along(c0))
title("Log Bayes factor in favor of the zero-mean model")
```


## Example 10.6: Testing for heterogeneity of the no-income risk

First, we re-load the data from Chapter 3.

```{r}
data("labor", package = "BayesianLearningCode")
labor <- subset(labor,
                income_1997 != "zero" & female,
                c(income_1998, wcollar_1986))
labor <- with(labor,
              data.frame(unemployed = income_1998 == "zero",
                         wcollar = wcollar_1986))
```

Next, we compute the marginal likelihoods for the homogeneity model.

```{r}
N <- length(labor$unemployed)
SN <- sum(labor$unemployed)
hN <- SN / N
N0 <- c(2, 10)
a0 <- c(1, 0.5, hN * N0) 
b0 <- c(1, 0.5, N0 * (1 - hN))

aN <- a0 + sum(labor$unemployed)
bN <- b0 + N - sum(labor$unemployed)

logmarglikM1 <- lbeta(aN, bN) - lbeta(a0, b0)
```

And for the heterogeneity model.

```{r}
N1 <- with(labor, sum(wcollar))
SN1 <- with(labor, sum(wcollar & unemployed))
N2 <- with(labor, sum(!wcollar))
SN2 <- with(labor, sum(!wcollar & unemployed))
hN <- (SN1 + SN2) / (N1 + N2) # redundant, same as above
a01 <- a02 <- c(1, 0.5, hN * N0) # redundant, same as a0
b01 <- b02 <- c(1, 0.5, N0 * (1 - hN)) # redundant, same as b0

aN1 <- a01 + SN1
bN1 <- b01 + N1 - SN1
aN2 <- a02 + SN2
bN2 <- b02 + N2 - SN2

logmarglikM2 <- lbeta(aN1, bN1) + lbeta(aN2, bN2) -
   lbeta(a01, b01) - lbeta(a02, b02)
```

We can now compute the model probabilities and the log BFs.

```{r}
logBF <- logmarglikM1 - logmarglikM2
PrM2 <- 0.5 / (0.5 + 0.5 * exp(logBF))
PrM1 <- 1 - PrM2
#print(xtable::xtable(res, digits = c(0, 2, 2, 5, 5, 5), row.names = FALSE))
knitr::kable(res <- cbind(logmarglikM1, logmarglikM2, logBF, PrM1, PrM2))
```

## Example 10.7: Testing for heterogeneity of the mortality rate

We proceed exactly as above, just with different data.

```{r}
N1  <- 1668 # number at risk in City A
SN1 <- 2    # cancer deaths in City A
N2  <- 583  # number at risk in City B
SN2 <- 1    # cancer deaths in City B
```

We again begin with the homogeneity model.

```{r}
N <- N1 + N2
SN <- SN1 + SN2
hN <- SN / N
N0 <- c(2, 10)
a0 <- c(1, 0.5, hN * N0) 
b0 <- c(1, 0.5, N0 * (1 - hN))

aN <- a0 + SN
bN <- b0 + N - SN

logmarglikM1 <- lbeta(aN, bN) - lbeta(a0, b0)
```

Followed by the heterogeneity model.

```{r}
a01 <- a02 <- a0
b01 <- b02 <- b0

aN1 <- a01 + SN1
bN1 <- b01 + N1 - SN1
aN2 <- a02 + SN2
bN2 <- b02 + N2 - SN2

logmarglikM2 <- lbeta(aN1, bN1) + lbeta(aN2, bN2) -
   lbeta(a01, b01) - lbeta(a02, b02)
```

We can now compute the model probabilities and the log BFs.

```{r}
logBF <- logmarglikM1 - logmarglikM2
PrM2 <- 0.5 / (0.5 + 0.5 * exp(logBF))
PrM1 <- 1 - PrM2
#print(xtable::xtable(res, digits = c(0, 2, 2, 3, 3, 3), row.names = FALSE))
knitr::kable(res <- cbind(logmarglikM1, logmarglikM2, logBF, PrM1, PrM2))
```

## Example 10.8: Testing for a structural break in the road safety data

We load the data.

```{r}
data("accidents", package = "BayesianLearningCode")
```

We define the priors hyperparameters and compute the log marginal likelihood
for the exchangable model.

```{r}
y <- accidents[, "children_accidents"]
a0_tmp <- c(0.01, 0.1, 0.5, 1, 2)
m0_tmp <- c(1, mean(y), 5, 7, 10)
grid <- expand.grid(a0 = a0_tmp, m0 = m0_tmp)
a0 <- grid$a0
b0 <- grid$a0 / grid$m0
aN <- sum(y) + a0
bN <- length(y) + b0
logmarglikM1 <- matrix(a0 * log(b0) + lgamma(aN) -
                       aN * log(bN) - lgamma(a0) -
                       sum(lgamma(y + 1)),
                       nrow = length(a0_tmp), ncol = length(m0_tmp),
                       dimnames = list(a0 = a0_tmp, m0 = round(m0_tmp, 2)))
knitr::kable(round(logmarglikM1, 2))
```

And now the same for the model with structural break.

```{r}
accidents1 <- window(accidents, end = c(1994, 9))
accidents2 <- window(accidents, start = c(1994, 10))

a01 <- a02 <- a0
b01 <- b02 <- b0

aN1 <- a01 + sum(accidents1[, "children_accidents"])
aN2 <- a02 + sum(accidents2[, "children_accidents"])
bN1 <- b01 + length(accidents1[, "children_accidents"])
bN2 <- b02 + length(accidents2[, "children_accidents"])

logmarglikM2 <- matrix(a01 * log(b01) + lgamma(aN1) +
                       a02 * log(b02) + lgamma(aN2) -
                       aN1 * log(bN1) - lgamma(a01) -
                       aN2 * log(bN2) - lgamma(a02) -
                       sum(lgamma(y + 1)),
                       nrow = length(a0_tmp), ncol = length(m0_tmp),
                       dimnames = list(a0 = a0_tmp, m0 = round(m0_tmp, 2)))
knitr::kable(round(logmarglikM2, 2))
```

We can now compute log Bayes factors and corresponding model probabilities
(under uniform prior probabilities).

```{r}
logBF <- logmarglikM1[, 2] - logmarglikM2[, 2]
PrM2 <- 0.5 / (0.5 + 0.5 * exp(logBF))
PrM1 <- 1 - PrM2
knitr::kable(round(rbind(PrM1, PrM2), 3))
```

## Example 10.9: Testing normal versus t

After loading the data, we define prior hyperparameters and compute log
marginal likelihoods. This is straightforward for the normal model.

```{r}
nu <- 7
y <- 100 * diff(log(exrates$USD / exrates$CHF))
N <- length(y)

c10 <- c30 <- 1
C10 <- 2
C30 <- (nu - 2) / nu * C10

c1N <- c10 + N / 2
C1N <- C10 + sum(y^2) / 2

logmarglikM1 <- lgamma(c1N) + c10 * log(C10) -
   lgamma(c10) - c1N * log(C1N) - 0.5 * N * log(2 * pi)
```

For the Student $t$ model, we need, e.g., numerical integration.

```{r}
integrand_nonvec <- function(sigma2, y, c0, C0, nu, const = 0) {
  N <- length(y)
  logint <- -(N/2 + c0 + 1) * log(sigma2) -
            0.5 * (nu + 1) * sum(log(1 + y^2 / (nu * sigma2))) -
            C0 / sigma2
  exp(logint + const)
}
integrand <- Vectorize(integrand_nonvec, "sigma2")

resolution <- 1000
grid <- seq(0.01, 1, length.out = resolution + 1)

tmp <- integrand(grid, y, c30, C30, nu)
const <- -log(max(tmp))
const <- 0
tmp <- integrand(grid, y, c30, C30, nu, const = const)
logarea <- log(sum(diff(grid) * .5 * (head(tmp, -1) + tail(tmp, -1)))) - const

logmarglikM3 <- c30 * log(C30) +
                N * lgamma((nu + 1) / 2) -
                lgamma(c30) -
                N * lgamma(nu / 2) -
                N / 2 * log(nu * pi) +
                logarea
```

We can now compute log Bayes factors and corresponding model probabilities
(under uniform prior probabilities).

```{r}
(logBF <- logmarglikM1 - logmarglikM3)
PrM2 <- 0.5 / (0.5 + 0.5 * exp(logBF))
PrM1 <- 1 - PrM2
knitr::kable(round(rbind(PrM1, PrM2), 3))
```
