---
title: "Chapter 11: Bayesian Model Selection in Regression and Time Series Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Chapter 11: Bayesian Model Selection in Regression and Time Series Analysis}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6
)
knitr::opts_knit$set(global.par = TRUE)
pdfplots <- FALSE # default: FALSE; set this to TRUE only if you like pdf figures
mcmcspeedup <- 10 # default: 1; set this to something higher for fewer draws
seed <- 42
```

# Section 11.4: Model Selection Problems in Time Series Analysis

## Section 11.4.1: Selecting the model order in an AR(p) model

Akin to Chapter 7, we load the US GDP data, restrict our analysis to the time
before the COVID outbreak, and compute log returns.

```{r}
data("gdp", package = "BayesianLearningCode")
dat <- gdp[1:which(names(gdp) == "2019-10-01")]
logret <- log(dat[-1]) - log(dat[-length(dat)])
logret <- ts(logret, start = c(1947, 2), end = c(2019, 4),
             frequency = 4)
```

We re-use the regression function defined in Chapter 7, again using the tools
developed in Chapter 6. In addition, we store the parameters of the conditional
posteriors to compute Savage-Dickey density ratios and marginal likelihoods at
a later stage.

```{r}
library("BayesianLearningCode")
library("mvtnorm")

regression <- function(y, X, b0, B0, c0, C0,
                       nburn = 1000L, M = 100000L / mcmcspeedup) {
  
  N <- nrow(X)
  d <- ncol(X)
  
  if (length(b0) == 1L) b0 <- rep(b0, d)
  
  if (!is.matrix(B0)) {
    if (length(B0) == 1L) {
      B0 <- diag(rep(B0, d))
    } else {
      B0 <- diag(B0)
    }
  }
  
  # Precompute some values
  B0inv <- solve(B0)
  B0invb0 <- B0inv %*% b0
  cN <- c0 + N / 2
  XX <- crossprod(X)
  Xy <- crossprod(X, y)
    
  # Prepare memory to store the draws
  betas <- matrix(NA_real_, nrow = M, ncol = d)
  sigma2s <- rep(NA_real_, M)
  colnames(betas) <- colnames(X)
  
  # Prepare memory to store the parameters
  paras <- list(bN = matrix(NA_real_, nrow = M, ncol = d),
                BN = array(NA_real_, dim = c(M, d, d)),
                cN = rep(NA_real_, M),
                CN = rep(NA_real_, M))
    
  # Set the starting value for beta
  sigma2 <- var(y) / 2
    
  # Run the Gibbs sampler
  for (m in seq_len(nburn + M)) {
    # Sample beta from its full conditional
    BN <- solve(B0inv + XX / sigma2) 
    bN <- BN %*% (B0invb0 + Xy / sigma2)
    beta <- rmvnorm(1, mean = bN, sigma = BN)
    
    # Sample sigma^2 from its full conditional
    eps <- y - tcrossprod(X, beta)
    CN <- C0 + crossprod(eps) / 2
    sigma2 <- rinvgamma(1, cN, CN)
      
    # Store the results
    if (m > nburn) {
      betas[m - nburn, ] <- beta
      sigma2s[m - nburn] <- sigma2
      paras[["bN"]][m - nburn, ] <- bN
      paras[["BN"]][m - nburn, , ] <- BN
      paras[["cN"]][m - nburn] <- cN
      paras[["CN"]][m - nburn] <- CN
    }
  }
  list(betas = betas, sigma2s = sigma2s, paras = paras, X = X, y = y,
       b0 = b0, B0 = B0, c0 = c0, C0 = C0)
}
```

We also need a function to create the AR-specific design matrix (also re-used
from Chapter 7, now with the added option to condition on more initial data
than then bare minimum).

```{r}
ARdesignmatrix <- function(dat, p = 1, conditioninglength = p) {
  d <- p + 1
  N <- length(dat) - p

  Xy <- matrix(NA_real_, N, d)
  Xy[, 1] <- 1
  for (i in seq_len(p)) {
    Xy[, i + 1] <- dat[(p + 1 - i) : (length(dat) - i)]
  }
  Xy[(1 + conditioninglength - p):N, , drop = FALSE]
}
```

Now we are ready to reproduce the results in the book.

## Example 11.9: US GDP data: Qualitative model-order selection

We obtain draws and parameters for four AR models under the semi-conjugate
prior.

```{r}
set.seed(seed)
b0 <- 0
B0 <- 1
res <- vector("list", 5)
for (i in seq_along(res)) {
  p <- i - 1
  if (p > 0) y <- tail(logret, -p) else y <- logret
  Xy <- ARdesignmatrix(logret, p)
  res[[i]] <- regression(y, Xy, b0 = b0, B0 = B0, c0 = 2, C0 = 0.001)
}
```

## Figure 11.2: Graphical assessment of the Savage-Dickey density ratio

Now we plot the draws for the leading coefficient, i.e., the coefficient
corresponding to the highest lag in each of the models, along with the marginal
prior.

```{r, echo = -c(1:2)}
if (pdfplots) {
  pdf("11-4_1.pdf", width = 6, height = 3)
}
par(mfrow = c(1, 4), mar = c(2.5, 1.5, 1.5, .1), mgp = c(1.5, .5, 0))
mymin <- -.75
mymax <- .75
mybreaks <- seq(mymin, mymax, by = .02)
myats <- seq(mymin, mymax, length.out = 200)
for (i in 2:5) {
  p <- i - 1
  hist(res[[i]]$betas[, p + 1], freq = FALSE, main = bquote(AR(.(p))),
       xlab = bquote(phi[.(p)]), ylab = "", breaks = mybreaks, border = NA)
  abline(v = 0, lty = 3)
  abline(h = 0, lty = 3)
  lines(myats, dnorm(myats, b0, sqrt(B0)), lwd = 1.5)
}
```

## Example 11.10: US GDP data: Quantitative model-order selection

After visualizing the Savage-Dickey density ratio, we now move towards
computing its numerical value by Rao-Blackwellization. We first define
a numerically stable version of the log-sum-exp (log-mean-exp) function.

```{r}
logsumexp <- function(x) {
  maxx <- max(x)
  maxx + log(sum(exp(x - maxx)))
}

logmeanexp <- function(x) {
  logsumexp(x) - log(length(x))
}
```

Now we are ready to compute.

```{r}
logSD <- rep(NA_real_, length(res) - 1)
for (i in seq_len(length(res) - 1)) {
  p <- i
  means <- res[[i + 1]]$paras$bN[, p + 1]
  sds <- sqrt(res[[i + 1]]$paras$BN[, p + 1, p + 1])
  logSD[i] <- logmeanexp(dnorm(0, means, sds, log = TRUE)) -
    dnorm(0, b0, sqrt(B0), log = TRUE)
}
names(logSD) <- c("0 vs 1", "1 vs 2", "2 vs 3", "3 vs 4")
round(logSD, 2)
```

Note that the first value of logSD is numerically rather unstable.

To reproduce this using Chib's method, we need to be careful to compare models
which were estimated conditional on the same initial data. Thus, we re-estimate
conditioning on one extra initial observation.

```{r}
set.seed(seed)
res2 <- vector("list", 5)
for (i in seq_along(res2)) {
  p <- i - 1
  y <- tail(logret, -(p + 1))
  Xy <- ARdesignmatrix(logret, p, conditioninglength = p + 1)
  res2[[i]] <- regression(y, Xy, b0 = b0, B0 = B0, c0 = 2, C0 = 0.001)
}
```

And now, we always condition on the first four observations (irrespectively
of the model order). This makes it possible to compare all AR models via
model probabilities.

```{r}
set.seed(seed)
res3 <- vector("list", 5)
for (i in seq_along(res3)) {
  p <- i - 1
  y <- tail(logret, -4)
  Xy <- ARdesignmatrix(logret, p, conditioninglength = 4)
  res3[[i]] <- regression(y, Xy, b0 = b0, B0 = B0, c0 = 2, C0 = 0.001)
}
```

We now compute Chib's estimator, evaluated at the posterior median.

```{r}
allres <- list(res, res2, res3)
logML <- matrix(NA_real_, length(allres), length(res))
for (i in seq_along(allres)) {
  for (j in seq_len(length(res))) {
    myres <- allres[[i]][[j]]
    beta_med <- apply(myres$betas, 2, median)
    sigma2_med <- median(myres$sigma2s)
    BN_med <- solve(solve(myres$B0) + crossprod(myres$X) / sigma2_med)
    bN_med <- BN_med %*% (solve(myres$B0) %*% myres$b0 + 
                            crossprod(myres$X, myres$y) / sigma2_med)
    logML[i, j] <-
      sum(dnorm(myres$y, myres$X %*% beta_med, sqrt(sigma2_med), log = TRUE)) +
      dmvnorm(beta_med, myres$b0, myres$B0, log = TRUE) +
      dinvgamma(sigma2_med, myres$c0, myres$C0, log = TRUE) -
      dmvnorm(beta_med, bN_med, BN_med, log = TRUE) -
      logmeanexp(dinvgamma(sigma2_med, myres$paras$cN, myres$paras$CN, log = TRUE))
  }
}
dimnames(logML) <- list("Conditioning set length" = c("p", "p + 1", "4"),
                        "Lag order" = seq_len(ncol(logML)))
knitr::kable(round(logML, 2))
```

The values below should be equal to the logSD values from above.

```{r}
pairwiselogML <- logML[2, 1:4] - logML[1, 2:5]
names(pairwiselogML) <- names(logSD)
round(pairwiselogML, 2)
```

## Example 11.11: US GDP data: Choosing the model order via marginal likelihoods

To compute model probabilities, we use the uniform and the penalty prior.

```{r}
# Uniform prior:
stableML <- exp(logML[3, ] - max(logML[3, ]))
probs_unif <- stableML / sum(stableML)

# Penalty prior:
tmp <- logML[3, ] - log(seq_along(stableML)) - log(seq_along(stableML) + 1)
stablepost <- exp(tmp - max(tmp))
probs_pen <- stablepost / sum(stablepost)

knitr::kable(t(round(cbind(unif = probs_unif, penalized = probs_pen), 3)))
```

## Section 11.4.2: Bayesian unit root testing

### Example 11.12: CHF exchange rate data: Exploring stationarity

Let us check for stationarity of the exchange rate data. First, we load the
data and visualize it as well as its empirical ACF. We do the same for the
absolute returns.

```{r, echo = -c(1:2), fig.width = 8, fig.height = 3}
if (pdfplots) {
  pdf("7-3_1.pdf", width = 8, height = 4)
}
par(mfrow = c(2, 2), mar = c(2.5, 2.5, 1.2, .5), mgp = c(1.5, .5, 0), lwd = 2)
data("exrates", package = "stochvol")
dat <- exrates$USD / exrates$CHF
ret <- diff(dat)
plot(dat, type = "l", main = "CHF-USD exchange rate", xlab = "Days",
     ylab = "CHF in USD")
plot(ret, type = "l", main = "CHF-USD daily returns", xlab = "Days",
     ylab = "CHF-USD")
acf(dat)
acf(ret)
```

This clearly hints at non-stationarity of the exchange rate series and at
(first order) stationarity of the returns.


### Example 11.13: CHF exchange rate data: Exploring unit roots

To check more formally, we fit AR(p) models to both.

```{r}
ardat <- arret <- list()
for (p in 1:4) {
  y <- tail(dat, -p)
  Xy <- ARdesignmatrix(dat, p)
  ardat[[p]] <- regression(y, Xy, b0 = 0, B0 = 1, c0 = 2, C0 = 0.001)
  y <- tail(ret, -p)
  Xy <- ARdesignmatrix(ret, p)
  arret[[p]] <- regression(y, Xy, b0 = 0, B0 = 1, c0 = 2, C0 = 0.001)
}
```

Now we can graphically investigate stationarity as above.

```{r, echo = -c(1:2), fig.width = 8, fig.height = 3}
if (pdfplots) {
  pdf("7-3_2.pdf", width = 8, height = 4)
}
par(mfrow = c(2, 2), mar = c(2.5, 2.5, 1.5, .1), mgp = c(1.5, .5, 0), lwd = 2)
draws <- list(ardat[[2]]$betas[, 2:3], arret[[2]]$betas[, 2:3])
eigenvalues <- matrix(NA_complex_, nrow(draws[[1]]), ncol(draws[[1]]))
mains <- c("AR(2) on the raw series", "AR(2) on the returns")
for (i in seq_along(draws)) {
  plot(draws[[i]], main = mains[i], xlab = bquote(phi[1]),
     ylab = bquote(phi[2]), xlim = c(-2, 2), ylim = c(-1, 1),
     col = rgb(0, 0, 0, .05), pch = 16)
  polygon(c(-2, 0, 2, -2), c(-1, 1, -1, -1), border = 2)

  for (m in seq_len(nrow(draws[[i]]))) {
    Phi <- matrix(c(draws[[i]][m, ], c(1, 0)), byrow = TRUE, nrow = 2)
    eigenvalues[m, ] <- eigen(Phi, only.values = TRUE)$values
  }
  plot(eigenvalues, xlim = c(-1, 1), ylim = c(-1, 1), asp = 1,
       main = mains[i], col = rgb(0, 0, 0, .05), pch = 16)
  symbols(0, 0, 1, add = TRUE, fg = 2, inches = FALSE)
}
```

To explore whether the nonstationarity of the raw series could be caused by
a unit root, we investigate the posterior of $1 - \phi_1 - \dots - \phi_p$ for
$p = 1, \dots, 4$.

```{r, echo = -c(1:2), fig.width = 8, fig.height = 3}
if (pdfplots) {
  pdf("7-3_3.pdf", width = 7, height = 2.5)
}
par(mfrow = c(1, 4), mar = c(2.5, 2.5, 1.5, .1), mgp = c(1.5, .5, 0), lwd = 2)
toplot <- matrix(NA_real_, ndraws, 4)
for (p in 1:4)
  toplot[, p] <- rowSums(ardat[[p]]$betas[, 2:(p + 1), drop = FALSE]) - 1
for (p in 1:4) {
  hist(toplot[, p], freq = FALSE,
       breaks = seq(min(toplot), max(toplot), length.out = 20),
       main = paste0("AR(", p, ")"), xlab = expression(delta), ylab = "")
  abline(v = 0, lty = 2, col = 2)
}
```

We do the same for the inflation data.

```{r}
ardat <- arret <- list()
for (p in 1:4) {
  y <- tail(inflation, -p)
  Xy <- ARdesignmatrix(inflation, p)
  ardat[[p]] <- regression(y, Xy, b0 = 0, B0 = 1, c0 = 2, C0 = 0.001)
}
```

```{r, echo = -c(1:2), fig.width = 8, fig.height = 3}
if (pdfplots) {
  pdf("7-3_4.pdf", width = 7, height = 2.5)
}
par(mfrow = c(1, 4), mar = c(2.5, 2.5, 1.5, .1), mgp = c(1.5, .5, 0), lwd = 2)
for (p in 1:4)
  toplot[, p] <- rowSums(ardat[[p]]$betas[, 2:(p + 1), drop = FALSE]) - 1
for (p in 1:4) {
  hist(toplot[, p], freq = FALSE,
       breaks = seq(min(toplot), max(toplot), length.out = 20),
       main = paste0("AR(", p, ")"), xlab = expression(delta), ylab = "")
  abline(v = 0, lty = 2, col = 2)
}
```

## Section 11.4.3: Bayesian testing for first-order Markov Chain models

## Example 11.18: Application to Austrian wage mobility data - homogeneity versus grouped model

We load the data and reduce the observations to workers from the birth cohort
1946-1960. 

```{r}
data("labor", package = "BayesianLearningCode")
labor <- subset(labor, birthyear >= 1946 & birthyear <= 1960)
```

We extract the income columns and transform the data to obtain for
each worker the matrix which contains the number of transitions from
one class to the other, i.e., the matrix with values $N_{i,hk}$.


```{r}
income <- labor[, grepl("^income", colnames(labor))]
income <- sapply(income, as.integer)
colnames(income) <- gsub("income_", "", colnames(income))
getTransitions <- function(x, classes) {
    transitions <- matrix(0, length(classes), length(classes))
    for (i in seq_len(length(x) - 1)) {
        transitions[x[i], x[i + 1]] <- transitions[x[i], x[i + 1]] + 1
    }
    dimnames(transitions) <- list(from = classes, to = classes)
    transitions
}
income_transitions <-
    lapply(seq_len(nrow(income)),
           function(i) getTransitions(income[i, ], classes = 0:5))
```

We determine the marginal likelihood of the first-order Markov chain
model assuming homogeneity and grouping by gender.

```{r}
N_hk <- Reduce("+", income_transitions)
Ng_hk <- list(male = Reduce("+", income_transitions[!labor$female]),
              female = Reduce("+", income_transitions[labor$female]))
gammas <- c(1, 4)
K <- nrow(N_hk)
G <- length(Ng_hk)
logmarglikMH <- logmarglikMG <- numeric(2)
for (i in 1:2) {
    gamma <- gammas[i]
    logmarglikMH[i] <- K * (lgamma(K * gamma) - K * lgamma(gamma)) +
        sum(lgamma(gamma + N_hk)) -
        sum(lgamma(rowSums(gamma + N_hk)))
    logmarglikMG[i] <- G * K * (lgamma(K * gamma) - K * lgamma(gamma)) +
        sum(lgamma(gamma + Ng_hk[["male"]])) +
        sum(lgamma(gamma + Ng_hk[["female"]])) -
        sum(lgamma(rowSums(gamma + Ng_hk[["male"]]))) -
        sum(lgamma(rowSums(gamma + Ng_hk[["female"]])))
}
```

We also calculate the log BF and summarize results.

```{r}
res <- rbind(MH = c(logmarglikMH, NA, NA),
             MG = c(logmarglikMG, logmarglikMG - logmarglikMH))
colnames(res) <- c("gamma = 1", "gamma = 4",
                   "gamma = 1", "gamma = 4")
knitr::kable(res)
```

## Example 11.19: Application to Austrian wage mobility data - restricted models

We continue to compare restricted models to the grouped model. We
first calculate the log BFs for the restricted models.

```{r}
logBF_R1G <- logBF_R2G <- numeric(2)
for (i in 1:2) {
    gamma <- gammas[i]
    logBF_R1G[i] <- lbeta(gamma, gamma) -
        lbeta(gamma + Ng_hk[["male"]]["1", "0"],
              gamma + Ng_hk[["male"]]["1", "2"])
    aN1 <- gamma + Ng_hk[["male"]]["1", "0"]
    aN2 <- gamma + Ng_hk[["female"]]["1", "0"]
    bN1 <- sum(gamma + Ng_hk[["male"]]["1", -1])
    bN2 <- sum(gamma + Ng_hk[["female"]]["1", -1])
    logBF_R2G[i] <- lbeta(aN1 + aN2 - 1, bN1 + bN2 - 1) +
        2 * lbeta(gamma, (K - 1) * gamma) -
        lbeta(aN1, bN1) - lbeta(aN2, bN2) -
        lbeta(2 * gamma - 1, 2 * (K - 1) * gamma - 1)
}
```

We obtain the log marginal likelihoods for the restricted models from
the log marginal likelihoods of the grouped models and the log BFs. We
then also summarize the results.

```{r}
logmarglikR1 <- logmarglikMG + logBF_R1G
logmarglikR2 <- logmarglikMG + logBF_R2G
res <- rbind(MR1 = c(logmarglikR1, logBF_R1G),
             MR2 = c(logmarglikR2, logBF_R2G))
colnames(res) <- c("gamma = 1", "gamma = 4",
                   "gamma = 1", "gamma = 4")
knitr::kable(res)
```
