---
title: "Chapter 11: Bayesian Model Selection in Regression and Time Series Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Chapter 11: Bayesian Model Selection in Regression and Time Series Analysis}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6
)
knitr::opts_knit$set(global.par = TRUE)
pdfplots <- FALSE # default: FALSE; set this to TRUE only if you like pdf figures
mcmcspeedup <- 10 # default: 1; set this to something higher for fewer draws
seed <- 42
```

# Section 11.1: Marginal likelihood computations in regression analysis

### Example 11.1: Marginal likelihoods when the error variance is known

[no code]

### Example 11.2: Marginal likelihood estimation under a semi-conjugate prior

[no code]

# Section 11.2: Variable selection in standard regression analysis

## Section 11.2.1: Bayesian variable selection

### Example 11.3: Movie data: Full enumeration marginal likelihoods

We  write a function that computes the marginal likelihoods for models defined by indicator vectors.
```{r}
logmarlik_reg <- function(y,X, a0, A0, B0, c0, C0, gammas){
  
  N <- length(y)
  
  cN <- c0 + N / 2
  const <- (- N / 2) * log(2*pi) + lgamma(cN) - lgamma(c0) + c0 * log(c0)
  
  nmod <- dim(gammas)[1]
  p <- dim(gammas)[2]
  
  lmarlik <- rep(NA,nmod)
  
  for (i in 1:nmod){
       
    gamma <- gammas[i,]
    k_gamma <- sum(gamma)
    
    if (k_gamma==0){
       X_gamma <-  as.matrix(rep(1, N))
       b0v <- t(as.vector(a0))
       B0v.inv<- as.matrix(1/A0)
    }else{
       X_gamma <- as.matrix(cbind(rep(1, N), X[,gamma==1] ))
       b0v <- c(a0,rep(0,k_gamma))
       B0v.inv <- diag(c(1/A0,rep(1/B0,k_gamma)) )
    }
    BNv.inv <- B0v.inv + crossprod(X_gamma)
    BNv <- solve(BNv.inv)

    bNv <- BNv %*% (crossprod(X_gamma, y) + B0v.inv%*% b0v )
    SSE <- as.numeric(crossprod(y) +  t(b0v) %*% B0v.inv %*% b0v -
                    t(bNv) %*% BNv.inv %*% bNv)
    
   CN <- C0 + SSE / 2
  lmarlik[i]  <- const   - cN * log(CN) +
                  0.5 *( log(det(BNv)) + log(det(B0v.inv)) ) 
  }
  return(lmarlik)
}
```
We load the data and choose covariates Budget, Screens and Comedy which we center at their means, and define the indicators for all $2^3=8$ models 
```{r}
# Example 11.2
library("BayesianLearningCode")
data("movies", package = "BayesianLearningCode")

y <- movies[, "OpenBoxOffice"]

covs <- c("Budget", "Screens","Comedy")
covs.cen <- scale(movies[, covs], scale = FALSE) # center the covariates
gammas <- cbind(c(rep(0,4),rep(1,4)),rep(c(rep(0,2), rep(1,2)),2),
                rep(c(0,1),4)) 
```
Next we choose the prior parameters and compute the log marginal likelihoods.
```{r}
a0=0
A0=10000

B0=1
c0 = 2.5
C0 = 1.5
logmarliks <- logmarlik_reg(y, covs.cen, a0, A0, B0, c0, C0, gammas) 
```

Finally, we compute the model probabilities for a uniform prior on the model space. 

```{r}
prob <- 1 / dim(gammas)[1]
h <- exp(logmarliks - max(logmarliks))
cM <- prob * sum(h)
model_probs <- prob * h / cM

knitr::kable(cbind(gammas, logmarliks, model_probs),
             digits = c(rep(0, ncol(gammas)), 1, 3))
```

We next perform model selection also for the standardized covariates.

```{r}
covs.std <- scale(movies[, covs], scale = TRUE) 
logmarliks <- logmarlik_reg(y, covs.std, a0, A0, B0, c0, C0, gammas) 

h <- exp(logmarliks - max(logmarliks))
cM <- prob * sum(h)
model_probs <- prob * h / cM

knitr::kable(cbind(gammas, logmarliks, model_probs),
             digits = c(rep(0, ncol(gammas)), 1, 3))
```

## Section 11.2.2: Model space MCMC

### Example 11.4: Movie data: Model space MCMC
We start by implementing the model space MH algorithm and a function
that computes the frequencies different models are visited. 

```{r}
modelspace_mh <- function(y,X, a0, A0, B0, c0, C0,
                          burnin = 1000L, M = 5000L){
  p<-dim(X)[2] # change to X later

  gamma_post <- matrix(ncol = p, nrow = M)
  acc <- numeric(length = M)

  gamma <- matrix(rep(1,p),nrow=1)
  lmarlik_old <-logmarlik_reg(y,X, a0, A0, B0, c0, C0, gamma)

  for (m in seq_len(burnin + M)) {
    gamma_proposed <- gamma_old <- gamma

    j <- sample(1:p, size=1)
    gamma_proposed[j] <- 1-gamma_old[j]
    lmarlik_proposed <-logmarlik_reg(y,X, a0, A0, B0, c0, C0, gamma_proposed)
    
    # compute acceptance probability and decide on acceptance
    log_acc <- lmarlik_proposed - lmarlik_old

    if (log(runif(1)) < log_acc) {
       gamma <- gamma_proposed
       lmarlik_old <- lmarlik_proposed
       accept <- 1
    } else {
       gamma <- gamma_old
       accept <- 0
    }

    if (m > burnin) {
        gamma_post[m-burnin, ] <- gamma
        acc[m-burnin] <- accept
    }
  }
  return(list(gamma_post=gamma_post,acc=acc))
}

number_draws<- function(gamma_post, models){

  nmod <- dim(models)[1]
  freq <- rep(NA,nmod)

  for (j in (1:nmod)){
      freq[j] <- sum(apply(gamma_post, 1, 
                           function(x) identical(x, models[j,])))
  }
  return(freq)
}
```

We run the MH algorithm and compute how often each model is visited.

```{r}
set.seed(seed)
M <- 50000
res <- modelspace_mh(y, covs.cen, a0, A0, B0, c0, C0, M = M)

model_freq  <- number_draws(res$gamma_post, gammas)
knitr::kable(cbind(gammas, model_freq, model_freq / M),
            digits = c(rep(0, ncol(gammas)), 0, 3))
```

Finally, we compute the PIPs.

```{r}
p <- dim(gammas)[2]

PIP <- rep(NA,p)
 for (j in (1:p)){
   PIP[j]=sum(model_freq[gammas[,j]==1])/M
 }

print(PIP)
```

We run model space MCMC and compute PIPs also for the standardized covariates.

```{r}
set.seed(seed)
res <- modelspace_mh(y, covs.std, a0, A0, B0, c0, C0, M = M)

model_freq  <- number_draws(res$gamma_post, gammas)
knitr::kable(cbind(gammas, model_freq, model_freq / M),
            digits = c(rep(0, ncol(gammas)), 0, 3))

p <- dim(gammas)[2]

PIP <- rep(NA,p)
 for (j in (1:p)){
   PIP[j]=sum(model_freq[gammas[,j]==1])/M
 }

print(PIP)
```
### Example 11.5: Marginal likelihoods when the error variance is known

[no code]


## Section 11.2.3: Benchmark priors for comparing regression models

We  again start writing  a function that 
```{r}
varselreg_gunif<- function(y, X, pri=NA, burnin=1000L, M=50000L){
  N=length(y)

  p <- dim(X)[2]
  g <- N

  gamma_post <- matrix(ncol = p, nrow = M)
  kgamma_post <- numeric(length = M)
  acc <- numeric(length = M)

  gamma_old <- matrix(rep(1,p),nrow=1)
  k_gamma=sum(gamma_old)
  X_gamma <- X

  BN_gamma <- solve( ((1+g)/g) * crossprod(X_gamma) )
  R2_gamma <- t(y)%*%X_gamma%*%BN_gamma%*%t(X_gamma)%*%y / (N*var(y))

  lmarlik_old <- (N-k_gamma-1)/2*log(1+g) -(N/2) * log(1+g*(1-R2_gamma) )

  for (m in seq_len(burnin + M)) {

    gamma_proposed <- gamma_old
    j <- sample(1:p, size=1)
    gamma_proposed[j] <- 1-gamma_old[j]

    kgamma <- sum(gamma_proposed)

   if (kgamma==0){ 
      R2_gamma <- 0 
     
   }else{
    X_gamma <- X[,gamma_proposed==1]
    
    BNinv_gamma <- ((1+g)/g) * crossprod(X_gamma)
    BN_gamma <- solve( BNinv_gamma)
    R2_gamma <- t(y) %*% X_gamma %*% BN_gamma %*% t(X_gamma) %*% y / (N*var(y))
   }

    lmarlik_proposed <- (N-kgamma-1)/2*log(1+g) -
                           (N-1)/2 * log(1+g*(1-R2_gamma) )

    # compute acceptance probability and accept or not
    log_acc <- lmarlik_proposed - lmarlik_old

    if (log(runif(1)) < log_acc) {
      
      gamma_old <- gamma_proposed
      lmarlik_old <- lmarlik_proposed
      accept=1
      
    } else {
      accept=0
    }

    if (m > burnin) {
      gamma_post[m-burnin, ] <- gamma_old
      kgamma_post[m-burnin] <- sum(gamma_old)
      acc[m-burnin] <- accept
    }

  }
  return(list(gamma_post=gamma_post,acc=acc))
}
```
As a  first exercise we apply the g-prior to the model with the centered covariates used above.

```{r}
set.seed(seed)
res_cen<-varselreg_gunif(y, covs.cen, M=100000)
print(colMeans(res_cen$gamma_post))

gammas_unique<-unique(res_cen$gamma_post)
freq_gammas<-number_draws(res_cen$gamma_post,gammas_unique) 
print(cbind(gammas_unique,freq_gammas))

res_std <- varselreg_gunif(y, covs.std, M=100000)
print(colMeans(res_std$gamma_post))

gammas_unique <- unique(res_std$gamma_post)
freq_gammas<-number_draws(res_std$gamma_post,gammas_unique) 
print(cbind(gammas_unique,freq_gammas))
```
We see that results are very similar as the g-prior takes into account the 
scale of the covariates. 
 
### Example 11.6: Movie data: Applying the g-prior

We now  use all covariates from example 6.7  except  the categorical variable MMPA rating.
```{r}
covs <- c("Comedy", "Thriller", "Budget", "Weeks", "Screens",
          "S-4-6", "S-1-3", "Vol-4-6", "Vol-1-3")
covs.cen <- scale(movies[, covs], scale = FALSE)

M <- 100000
p <- dim(covs.cen)[2]

set.seed(seed)
res_gunif <- varselreg_gunif(y, covs.cen,M=M )
```
We show the PIPs in a barplot and compute the estimated posterior probabilities of all models visited during MCMC.

```{r}
if (pdfplots) {
  pdf("11-2_2a.pdf", width = 3, height = 3)
}
par(mfrow = c(1, 1), mar = c(2.5, 2.5, 1.5, .5), mgp = c(1.5, .5, 0))
barplot(colMeans(res_gunif$gamma_post), col="blue",names.arg=1:p, 
        xlab="Covariate", ylab="PIP")

gammas_unique <- unique(res_gunif$gamma_post)
nmod <- dim(gammas_unique)[1]
print(nmod)

freq_gammas<-number_draws(res_gunif$gamma_post,gammas_unique)
io <- order(freq_gammas, decreasing=TRUE)

knitr::kable(cbind(gammas_unique,freq_gammas/M)[io,],
             digits=cbind(rep(0,p),4))

#xtable(cbind(h,freqs)[io,],digits=c(rep(0, p+1),4))
```

We determine the posterior distribution of the size of the models and plot the MCMC draws and their distribution.

```{r}
if (pdfplots) {
  pdf("11-2_1.pdf", width = 6, height = 3)
}
par(mfrow = c(1, 2), mar = c(2.5, 2.5, 1.5, .5), mgp = c(1.5, .5, 0))
options(scipen = 999)

k_gamma=rowSums(res_gunif$gamma_post)
plot(k_gamma, type="l", xlab="Draw", ylab=expression(k[gamma]), xaxt="n",
          ylim=c(0,p)) 
axis(side = 1, at = seq(from=0, to=M, by=M/5),labels = T)

barplot(tabulate(k_gamma),  col ="blue",xlab=expression(k[gamma]),
        ylab="Frequencies", ylim=c(0,50000),  names.arg=1:p)
```

## Section 11.2.4: Priors on the model space

### Figure 11.3
```{r}
if (pdfplots) {
  pdf("11-2_3.pdf", width = 8, height = 5)
}
px=10
xx=0:px
par(mfrow = c(1, 2))
barplot(rbind(dbinom(xx, size=px, prob=0.05),
              dbinom(xx, size=px, prob=0.5)), beside=TRUE,
              col=c("darkblue","darkred"),ylim=c(0,0.6), names.arg=xx,
              xlab = "x", ylab = "P(X=x)",cex.axis=1.2,cex.names=1.2,
              legend.text=c(expression(paste(pi, " = 0.05")),
                            expression(paste(pi, " = 0.5"))  ) )
px=100
xx=0:px
barplot(rbind(dbinom(xx, size=px, prob=0.05),
              dbinom(xx, size=px, prob=0.5)), beside=TRUE,
              col=c("darkblue","darkred"),ylim=c(0,0.2), names.arg=xx,
              xlab = "x", ylab = "P(X=x)",cex.axis=1,cex.names=1,
              border=NA,
              legend.text=c(expression(paste(pi, " = 0.05")),
                            expression(paste(pi, " = 0.5"))  ) )
```
### Example 11.8: Movie Data: Hierarchical prior on the model space

We now implement variable selection with the g-prior under a hierarchical prior.
```{r}
varselreg_ghier<- function(y, X, 
                           prior_model=list(type="hier", a0=1,b0=1),
                           burnin=1000L, M=50000L){
  N=length(y)

  p <- dim(X)[2]
  g <- N

  gamma_post <- matrix(ncol = p, nrow = M)
  
  gamma <- matrix(rep(1,p),nrow=1)

  for (m in seq_len(burnin + M)) {

    # Step a: sample gamma
    j <- sample(1:p, size=1)
    p0 <- sum(gamma[-j])
    
    # compute prior "odds of the model
    if (prior_model$type=="hier"){
        pri.odds <- (prior_model$a0+p0)/(prior_model$b0+p-p0-1)
    }else if(prior_model$type=="unif"){
      pri.odds<- 1
    }else {stop("not implemented")}
    
   # determine the logmarginal likelihoods
    gamma[j]=0
    k_gamma <- sum(gamma)

    if (k_gamma==0){
      R2_gamma <- 0
    }else{
      X_gamma<-X[,gamma==1]
      BN_gamma <- solve( ((1+g)/g) * crossprod(X_gamma) )
      R2_gamma <- t(y) %*% X_gamma %*% BN_gamma %*% t(X_gamma) %*% y /
        (N*var(y))
    }

    lmarlik0 <- (N-k_gamma-1)/2*log(1+g) -(N/2) * log(1+g*(1-R2_gamma) )

    gamma[j]=1
    k_gamma=k_gamma+1
    X_gamma<-X[,gamma==1]
    BN_gamma <- solve( ((1+g)/g) * crossprod(X_gamma) )
    R2_gamma <- t(y) %*% X_gamma %*% BN_gamma %*% t(X_gamma) %*% y /
      (N*var(y))

    lmarlik1 <- (N-k_gamma-1)/2*log(1+g) -(N/2) * log(1+g*(1-R2_gamma) )
    Oj<- exp(lmarlik1-lmarlik0)*pri.odds

    # compute acceptance probability and accept or not
    gamma[j] <- 1*(-log(runif(1)) <= log(1+Oj))

    if (m > burnin) {
      gamma_post[m-burnin, ] <- gamma
    }
  }
  return(gamma_post)
}
```
We now  estimate the model under the hierarchical prior on the model space and show the PIPs.

```{r}
if (pdfplots) {
  pdf("11-2_2b.pdf", width = 3, height = 3)
}
set.seed(seed)
gamma_post_hier <- varselreg_ghier(y, covs.cen, M=M)

par(mfrow = c(1, 1), mar = c(2.5, 2.5, 1.5, .5), mgp = c(1.5, .5, 0))
barplot(colMeans(gamma_post_hier), col="blue",names.arg=1:p, 
        xlab="Covariate",ylab="PIP")

gammas_unique_hier <- unique(gamma_post_hier)
num_mod_hier <- dim(gammas_unique_hier)[1]
print(num_mod_hier)

freq_gammas_hier<- number_draws(gamma_post_hier,gammas_unique_hier)
io_hier <- order(freq_gammas_hier, decreasing=TRUE)

knitr::kable(cbind(gammas_unique_hier,freq_gammas_hier/M)[io_hier,],
             digits=cbind(rep(0, p),4))

#xtable(cbind(h,freqs)[io,],digits=c(rep(0,p+1),4))
```

We determine the distribution of the model space complexity 
```{r}
if (pdfplots) {
  pdf("11-2_2c.pdf", width = 5, height = 6)
}
k_gamma_hier=rowSums(gamma_post_hier)

barplot(tabulate(k_gamma_hier),  col="blue",xlab=expression(k[gamma]),
              ylab="Frequencies", ylim=c(0,50000), names.arg=1:p)

# TEST
#prior_model <- list(type="unif")
#gamma_post1<-varselreg_ghier(y, covs.cen,prior_model, M=50000)
```
and the posterior of $\pi$.


```{r}
if (pdfplots) {
  pdf("11-2_2d.pdf", width = 5, height = 6)
}

b0 <- a0 <- 1

p <- dim(covs.cen)[2]
xx=seq(from=0, to=1, by=0.001)
post_pi <- matrix(NA, ncol=M, nrow=length(xx))

for (m in (1:M)){
  post_pi[,m] <- dbeta(xx, a0+k_gamma_hier[m] , b0+p-k_gamma_hier[m])
}
plot(xx,rowMeans(post_pi) , type="l", col="blue", xlab=expression(pi),
     ylab=expression(paste("p(", pi,")")))
lines(xx,dbeta(xx, a0, b0),col="black")
legend("topleft",legend=c("posterior","prior"), col=c("blue","black"),lty=1)
```

## Section 11.2.5: Bayesian model averaging

### Example 11.9: Movie data: Bayesian model averaging
As we have sampled  the model indicators we can now sample the parameters
of the regression model, using the frequencies of the different models. We write 
a function for this task. An alternative would be to implement this step in
the function that performs variable selection.
```{r}
estimates_gprior<- function(y,X, gammas_unique, freq_gammas){
  
  N<- length(y)
  p<- dim(X)[2] 
  nmod <- length(freq_gammas)
  
  M <- sum(freq_gammas)
  g <- N
  cN<- (N-1)/2
  
   beta_post <-  matrix(0, ncol=p, nrow=M)
   sigma2_post <- rep(NA,M)
   
   m <- 0
   for (imod in 1:nmod){
     
     freq_imod <- freq_gammas[imod]
     ind_gammas <- gammas_unique[imod,]==1
     p_imod <- sum(ind_gammas)
     ind<-m + (1:freq_imod)
     
     X_gamma <- X[ ,ind_gammas==1]
     BN <- g/(1+g)*solve(crossprod(X_gamma))
     bN <- BN %*% crossprod(X_gamma,y)
     
     CN <- (N*var(y) - t(bN)%*% solve(BN)%*%bN)/2
     sigma2 <- rinvgamma(freq_imod, cN, CN)
     u <- mvtnorm::rmvnorm(freq_imod, mean=rep(0,p_imod),sigma = BN)
     
     beta_post[ind,ind_gammas] <- t(matrix(rep(bN,freq_imod), nrow=p_imod)) +
                             u*matrix(rep(sqrt(sigma2),p_imod), nrow=freq_imod)
     sigma2_post[ind] <- sigma2
     m <- m+freq_imod
   }
   return(list(sigma2_post=sigma2_post,beta_post=beta_post))
}
```
 and apply it to the results under both priors.
 
```{r}
set.seed(seed)
estimates_unif <- estimates_gprior(y,covs.cen,gammas_unique, freq_gammas)
estimates_hier <- estimates_gprior(y,covs.cen,gammas_unique_hier,freq_gammas_hier)
```
We determine the mean and the 2.5\% and 97.\% quantiles 
under the uniform and the hierarchical model space  g-prior and provide results in  a table
(left: uniform prior, right: hierarchical prior )
```{r}
res_mcmc <- function(x, lower = 0.025, upper = 0.975){
  res <- c(quantile(x, lower), mean(x), quantile(x, upper))
  names(res) <- c(paste0(lower * 100, "%"), "Posterior mean", 
                  paste0(upper *   100, "%"))
  res
}
res_unif<-t(apply(cbind(estimates_unif$beta_post,estimates_unif$sigma2_post),2,res_mcmc))
rownames(res_unif)<- c(covs, "sigma2")
res_hier<-t(apply(cbind(estimates_hier$beta_post,estimates_hier$sigma2_post),2,res_mcmc))

knitr::kable(round(cbind(res_unif,res_hier),3))
#xtable(cbind(res_unif,res_hier),digits=rep(2,7))
```
# Section 11.3: Model selection beyond standard regression analysis

## Section 11.3.1: Marginal likelihoods under transformed outcome variables

### Example 11.7: Movie data: Model comparison under log transformations

[TODO]<br />[TODO]<br />[TODO]<br />[TODO]<br />[TODO]<br />[TODO]<br />[TODO]<br />[TODO]<br />[TODO]<br />[TODO]<br />

## Section 11.3.2: Bayesian variable selection for a probit model

### Example 11.8: Labor market data: Bayesian variable selection

[TODO]<br />[TODO]<br />[TODO]<br />[TODO]<br />[TODO]<br />[TODO]<br />[TODO]<br />[TODO]<br />[TODO]<br />[TODO]<br />

# Section 11.4: Model selection problems in time series analysis

## Section 11.4.1: Selecting the model order in an AR(p) model

Akin to Chapter 7, we load the US GDP data, restrict our analysis to the time
before the COVID outbreak, and compute log returns.

```{r}
data("gdp", package = "BayesianLearningCode")
dat <- gdp[1:which(names(gdp) == "2019-10-01")]
logret <- log(dat[-1]) - log(dat[-length(dat)])
logret <- ts(logret, start = c(1947, 2), end = c(2019, 4),
             frequency = 4)
```

We re-use the regression function defined in Chapter 7, again using the tools
developed in Chapter 6. In addition, we store the parameters of the conditional
posteriors to compute Savage-Dickey density ratios and marginal likelihoods at
a later stage.

```{r}
library("BayesianLearningCode")
library("mvtnorm")

regression <- function(y, X, b0, B0, c0, C0,
                       nburn = 1000L, M = 100000L / mcmcspeedup) {
  
  N <- nrow(X)
  d <- ncol(X)
  
  if (length(b0) == 1L) b0 <- rep(b0, d)
  
  if (!is.matrix(B0)) {
    if (length(B0) == 1L) {
      B0 <- diag(rep(B0, d))
    } else {
      B0 <- diag(B0)
    }
  }
  
  # Precompute some values
  B0inv <- solve(B0)
  B0invb0 <- B0inv %*% b0
  cN <- c0 + N / 2
  XX <- crossprod(X)
  Xy <- crossprod(X, y)
    
  # Prepare memory to store the draws
  betas <- matrix(NA_real_, nrow = M, ncol = d)
  sigma2s <- rep(NA_real_, M)
  colnames(betas) <- colnames(X)
  
  # Prepare memory to store the parameters
  paras <- list(bN = matrix(NA_real_, nrow = M, ncol = d),
                BN = array(NA_real_, dim = c(M, d, d)),
                cN = rep(NA_real_, M),
                CN = rep(NA_real_, M))
    
  # Set the starting value for beta
  sigma2 <- var(y) / 2
    
  # Run the Gibbs sampler
  for (m in seq_len(nburn + M)) {
    # Sample beta from its full conditional
    BN <- solve(B0inv + XX / sigma2) 
    bN <- BN %*% (B0invb0 + Xy / sigma2)
    beta <- rmvnorm(1, mean = bN, sigma = BN)
    
    # Sample sigma^2 from its full conditional
    eps <- y - tcrossprod(X, beta)
    CN <- C0 + crossprod(eps) / 2
    sigma2 <- rinvgamma(1, cN, CN)
      
    # Store the results
    if (m > nburn) {
      betas[m - nburn, ] <- beta
      sigma2s[m - nburn] <- sigma2
      paras[["bN"]][m - nburn, ] <- bN
      paras[["BN"]][m - nburn, , ] <- BN
      paras[["cN"]][m - nburn] <- cN
      paras[["CN"]][m - nburn] <- CN
    }
  }
  list(betas = betas, sigma2s = sigma2s, paras = paras, X = X, y = y,
       b0 = b0, B0 = B0, c0 = c0, C0 = C0)
}
```

We also need a function to create the AR-specific design matrix (also re-used
from Chapter 7, now with the added option to condition on more initial data
than then bare minimum).

```{r}
ARdesignmatrix <- function(dat, p = 1, conditioninglength = p) {
  d <- p + 1
  N <- length(dat) - p

  Xy <- matrix(NA_real_, N, d)
  Xy[, 1] <- 1
  for (i in seq_len(p)) {
    Xy[, i + 1] <- dat[(p + 1 - i) : (length(dat) - i)]
  }
  Xy[(1 + conditioninglength - p):N, , drop = FALSE]
}
```

Now we are ready to reproduce the results in the book.

### Example 11.9: US GDP data: Qualitative model-order selection

We obtain draws and parameters for four AR models under the semi-conjugate
prior.

```{r}
set.seed(seed)
b0 <- 0
B0 <- 1
res <- vector("list", 5)
for (i in seq_along(res)) {
  p <- i - 1
  if (p > 0) y <- tail(logret, -p) else y <- logret
  Xy <- ARdesignmatrix(logret, p)
  res[[i]] <- regression(y, Xy, b0 = b0, B0 = B0, c0 = 2, C0 = 0.001)
}
```

Now we plot the draws for the leading coefficient, i.e., the coefficient
corresponding to the highest lag in each of the models, along with the marginal
prior.

```{r, echo = -c(1:2)}
if (pdfplots) {
  pdf("11-4_1.pdf", width = 6, height = 3)
}
par(mfrow = c(1, 4), mar = c(2.5, 1.5, 1.5, .1), mgp = c(1.5, .5, 0))
mymin <- -.75
mymax <- .75
mybreaks <- seq(mymin, mymax, by = .02)
myats <- seq(mymin, mymax, length.out = 200)
for (i in 2:5) {
  p <- i - 1
  hist(res[[i]]$betas[, p + 1], freq = FALSE, main = bquote(AR(.(p))),
       xlab = bquote(phi[.(p)]), ylab = "", breaks = mybreaks, border = NA)
  abline(v = 0, lty = 3)
  abline(h = 0, lty = 3)
  lines(myats, dnorm(myats, b0, sqrt(B0)), lwd = 1.5)
}
```

### Example 11.10: US GDP data: Quantitative model-order selection

After visualizing the Savage-Dickey density ratio, we now move towards
computing its numerical value by Rao-Blackwellization. We first define
a numerically stable version of the log-sum-exp (log-mean-exp) function.

```{r}
logsumexp <- function(x) {
  maxx <- max(x)
  maxx + log(sum(exp(x - maxx)))
}

logmeanexp <- function(x) {
  logsumexp(x) - log(length(x))
}
```

Now we are ready to compute.

```{r}
logSD <- rep(NA_real_, length(res) - 1)
for (i in seq_len(length(res) - 1)) {
  p <- i
  means <- res[[i + 1]]$paras$bN[, p + 1]
  sds <- sqrt(res[[i + 1]]$paras$BN[, p + 1, p + 1])
  logSD[i] <- logmeanexp(dnorm(0, means, sds, log = TRUE)) -
    dnorm(0, b0, sqrt(B0), log = TRUE)
}
names(logSD) <- c("0 vs 1", "1 vs 2", "2 vs 3", "3 vs 4")
round(logSD, 2)
```

Note that the first value of logSD is numerically rather unstable.

To reproduce this using Chib's method, we need to be careful to compare models
which were estimated conditional on the same initial data. Thus, we re-estimate
conditioning on one extra initial observation.

```{r}
set.seed(seed)
res2 <- vector("list", 5)
for (i in seq_along(res2)) {
  p <- i - 1
  y <- tail(logret, -(p + 1))
  Xy <- ARdesignmatrix(logret, p, conditioninglength = p + 1)
  res2[[i]] <- regression(y, Xy, b0 = b0, B0 = B0, c0 = 2, C0 = 0.001)
}
```

And now, we always condition on the first four observations (irrespectively
of the model order). This makes it possible to compare all AR models via
model probabilities.

```{r}
set.seed(seed)
res3 <- vector("list", 5)
for (i in seq_along(res3)) {
  p <- i - 1
  y <- tail(logret, -4)
  Xy <- ARdesignmatrix(logret, p, conditioninglength = 4)
  res3[[i]] <- regression(y, Xy, b0 = b0, B0 = B0, c0 = 2, C0 = 0.001)
}
```

We now compute Chib's estimator, evaluated at the posterior median.

```{r}
allres <- list(res, res2, res3)
logML <- matrix(NA_real_, length(allres), length(res))
for (i in seq_along(allres)) {
  for (j in seq_len(length(res))) {
    myres <- allres[[i]][[j]]
    beta_med <- apply(myres$betas, 2, median)
    sigma2_med <- median(myres$sigma2s)
    BN_med <- solve(solve(myres$B0) + crossprod(myres$X) / sigma2_med)
    bN_med <- BN_med %*% (solve(myres$B0) %*% myres$b0 + 
                            crossprod(myres$X, myres$y) / sigma2_med)
    logML[i, j] <-
      sum(dnorm(myres$y, myres$X %*% beta_med, sqrt(sigma2_med), log = TRUE)) +
      dmvnorm(beta_med, myres$b0, myres$B0, log = TRUE) +
      dinvgamma(sigma2_med, myres$c0, myres$C0, log = TRUE) -
      dmvnorm(beta_med, bN_med, BN_med, log = TRUE) -
      logmeanexp(dinvgamma(sigma2_med, myres$paras$cN, myres$paras$CN, log = TRUE))
  }
}
dimnames(logML) <- list("Conditioning set length" = c("p", "p + 1", "4"),
                        "Lag order" = seq_len(ncol(logML)))
knitr::kable(round(logML, 2))
```

The values below should be equal to the logSD values from above.

```{r}
pairwiselogML <- logML[2, 1:4] - logML[1, 2:5]
names(pairwiselogML) <- names(logSD)
round(pairwiselogML, 2)
```

### Example 11.11: US GDP data: Choosing the model order via marginal likelihoods

For convenience, we write a small function which converts log marginal
likelihoods to posterior model probabilities.

```{r}
logML2post <- function(logML, prior = rep(1, length(logML))) {
  logpost_unnorm <- logML + log(prior)
  post_unnorm <- exp(logpost_unnorm - max(logpost_unnorm))
  post_unnorm / sum(post_unnorm)
}
```

To compute model probabilities, we use the uniform and the penalty prior.

```{r}
# Uniform prior:
probs_unif <- logML2post(logML[3, ])

# Penalty prior:
prior <- 1 / (seq_along(logML[3, ]) * (seq_along(logML[3, ]) + 1))
probs_pen <- logML2post(logML[3, ], prior)

knitr::kable(t(round(cbind(unif = probs_unif, penalized = probs_pen), 3)))
```

## Section 11.4.2: Bayesian unit root testing

### Example 11.12: CHF exchange rate data: Exploring stationarity

Let us check for stationarity of the exchange rate data. First, we load the
data and visualize it as well as its empirical ACF. We do the same for the
absolute returns.

```{r, echo = -c(1:2), fig.width = 8, fig.height = 3}
if (pdfplots) {
  pdf("11-4_2.pdf", width = 8, height = 4)
}
par(mfrow = c(2, 2), mar = c(2.5, 2.5, 1.2, .5), mgp = c(1.5, .5, 0), lwd = 2)
data("exrates", package = "stochvol")
dat <- exrates$USD / exrates$CHF
ret <- diff(dat)
plot(dat, type = "l", main = "CHF-USD exchange rate", xlab = "Days",
     ylab = "CHF in USD")
plot(ret, type = "l", main = "CHF-USD daily returns", xlab = "Days",
     ylab = "CHF-USD")
acf(dat)
acf(ret)
```

This clearly hints at non-stationarity of the exchange rate series and at
(first order) stationarity of the returns.


### Example 11.13: CHF exchange rate data: Exploring unit roots

To check more formally, we fit AR(p) models to both.

```{r}
ardat <- arret <- list()
for (p in 1:4) {
  y <- tail(dat, -p)
  Xy <- ARdesignmatrix(dat, p)
  ardat[[p]] <- regression(y, Xy, b0 = 0, B0 = 1, c0 = 2, C0 = 0.001)
  y <- tail(ret, -p)
  Xy <- ARdesignmatrix(ret, p)
  arret[[p]] <- regression(y, Xy, b0 = 0, B0 = 1, c0 = 2, C0 = 0.001)
}
```

Now we can graphically investigate stationarity as above.

```{r, echo = -c(1:2), fig.width = 8, fig.height = 3}
if (pdfplots) {
  pdf("11-4_3.pdf", width = 8, height = 4)
}
par(mfrow = c(2, 2), mar = c(2.5, 2.5, 1.5, .1), mgp = c(1.5, .5, 0), lwd = 2)
draws <- list(ardat[[2]]$betas[, 2:3], arret[[2]]$betas[, 2:3])
eigenvalues <- matrix(NA_complex_, nrow(draws[[1]]), ncol(draws[[1]]))
mains <- c("AR(2) on the raw series", "AR(2) on the returns")
for (i in seq_along(draws)) {
  plot(draws[[i]], main = mains[i], xlab = bquote(phi[1]),
     ylab = bquote(phi[2]), xlim = c(-2, 2), ylim = c(-1, 1),
     col = rgb(0, 0, 0, .05), pch = 16)
  polygon(c(-2, 0, 2, -2), c(-1, 1, -1, -1), border = 2)

  for (m in seq_len(nrow(draws[[i]]))) {
    Phi <- matrix(c(draws[[i]][m, ], c(1, 0)), byrow = TRUE, nrow = 2)
    eigenvalues[m, ] <- eigen(Phi, only.values = TRUE)$values
  }
  plot(eigenvalues, xlim = c(-1, 1), ylim = c(-1, 1), asp = 1,
       main = mains[i], col = rgb(0, 0, 0, .05), pch = 16)
  symbols(0, 0, 1, add = TRUE, fg = 2, inches = FALSE)
}
```

To explore whether the nonstationarity of the raw series could be caused by
a unit root, we investigate the posterior of $1 - \phi_1 - \dots - \phi_p$ for
$p = 1, \dots, 4$.

```{r, echo = -c(1:2), fig.width = 8, fig.height = 3}
if (pdfplots) {
  pdf("11-4_4.pdf", width = 7, height = 2.5)
}
par(mfrow = c(1, 4), mar = c(2.5, 2.5, 1.5, .1), mgp = c(1.5, .5, 0), lwd = 2)
toplot <- matrix(NA_real_, nrow(ardat[[1]]$betas), length(ardat))
for (p in 1:4) {
  toplot[, p] <- rowSums(ardat[[p]]$betas[, 2:(p + 1), drop = FALSE]) - 1
}
for (p in 1:4) {
  hist(toplot[, p], freq = FALSE,
       breaks = seq(min(toplot), max(toplot), length.out = 20),
       main = paste0("AR(", p, ")"), xlab = expression(delta), ylab = "")
  abline(v = 0, lty = 2, col = 2)
}
```

We do the same for the inflation data (currently not included in the book).

```{r}
ardat <- arret <- list()
for (p in 1:4) {
  y <- tail(inflation, -p)
  Xy <- ARdesignmatrix(inflation, p)
  ardat[[p]] <- regression(y, Xy, b0 = 0, B0 = 1, c0 = 2, C0 = 0.001)
}
```

```{r, echo = -c(1:2), fig.width = 8, fig.height = 3}
if (pdfplots) {
  pdf("11-4_4a.pdf", width = 7, height = 2.5)
}
par(mfrow = c(1, 4), mar = c(2.5, 2.5, 1.5, .1), mgp = c(1.5, .5, 0), lwd = 2)
for (p in 1:4) {
  toplot[, p] <- rowSums(ardat[[p]]$betas[, 2:(p + 1), drop = FALSE]) - 1
}
for (p in 1:4) {
  hist(toplot[, p], freq = FALSE,
       breaks = seq(min(toplot), max(toplot), length.out = 20),
       main = paste0("AR(", p, ")"), xlab = expression(delta), ylab = "")
  abline(v = 0, lty = 2, col = 2)
}
```

### Example 11.XX: CHF exchange rate data: Testing for a unit root using the Savage-Dickey density ratio

We start by defining covariates and response, running the regression thereafter.

```{r}
set.seed(seed + 123)
deltay <- tail(diff(dat), -1)       # remove delta y_1
ylagged <- head(tail(dat, -1), -1)  # remove y_1 and y_T
deltaylagged <- head(diff(dat), -1) # remove delta y_T
Xy <- matrix(c(ylagged, deltaylagged, rep(1, length(ylagged))), ncol = 3)

c0 <- 2
C0 <- 0.001
b0 <- rep(0, 3)
A0 <- F0 <- 1
D0 <- 0.01^2
B0 <- diag(c(D0, F0, A0))

res <- regression(deltay, Xy, b0 = b0, B0 = B0, c0 = c0, C0 = C0)
```

Now we can visualize the Savage-Dickey density ratio.

```{r, echo = -c(1:2), fig.width = 8, fig.height = 3}
if (pdfplots) {
  pdf("11-4_5.pdf", width = 7, height = 4)
}
par(mfrow = c(1, 1), mar = c(2.5, 1.5, 1.5, .1), mgp = c(1.5, .5, 0), lwd = 1)
breaks <- seq(0.001 * floor(1000 * min(res$beta[,1])),
              0.001 * ceiling(1000 * max(res$beta[,1])),
              length.out = 40)
hist(res$betas[,1], probability = TRUE, breaks = breaks,
     xlab = expression(delta), ylab = "", border = NA,
     main = "Augmented Dickey Fuller regression")
lines(breaks, dnorm(breaks, b0[1], sqrt(B0[1, 1])), lwd = 1.5)
abline(v = 0, lty = 3)
abline(h = 0, lty = 3)
```

To compute the numerical value of the SD density ratio, we again use
Rao-Blackwellization.

```{r}
means <- res$paras$bN[, 1]
sds <- sqrt(res$paras$BN[, 1, 1])
logpostordinate <- logmeanexp(dnorm(0, means, sds, log = TRUE))
logSD <- logpostordinate - dnorm(0, b0[1], sqrt(B0[1, 1]), log = TRUE)
(round(exp(logSD), 2))
(round(sqrt(2 * pi * B0[1, 1]) * exp(logpostordinate), 2)) # equivalent
```


### Example 11.XX: Inflation data: Bayesian unit root testing with unknown model order

To compare all models, we need to condition on the same set of initial
observations.

```{r}
set.seed(seed - 123)
pu <- 4
D0 <- 1
A0 <- F0 <- 1
resG <- resDF <- vector("list", pu)

# Set up response and covariates:
deltay <- tail(diff(inflation), -(pu - 1))        # rm dy1 ... dy(pu - 1)
ylagged <- head(tail(inflation, -(pu - 1)), -1)   # rm y1 ... y(pu - 1), yT
for (p in seq_len(pu)) {
  deltaylagged <- matrix(NA_real_, nrow = length(deltay), ncol = p - 1)
  # if p == 1:  empty
  # if p == 2:  keep              dy(pu - 2) ... dy(T - 1)
  # if p == 3:  keep additionally dy(pu - 3) ... dy(T - 2)
  #      ...
  # if p == pu: keep additionally dy1        ... dy(T - pu - 1)
  for (j in seq_len(p - 1)) {
    deltaylagged[, j] <- diff(inflation)[seq(pu - j, length(inflation) - 1 - j)]
  }
  
  # Restricted model MG:
  Xy <- matrix(c(deltaylagged, rep(1, length(ylagged))), ncol = p)
  b0 <- rep(0, p)
  B0 <- diag(c(rep(F0, p - 1), A0))
  resG[[p]] <- regression(deltay, Xy, b0 = b0, B0 = B0, c0 = 2, C0 = 0.001)
  
  # Unrestricted model MDF:
  Xy <- matrix(c(ylagged, deltaylagged, rep(1, length(ylagged))), ncol = p + 1)
  b0 <- rep(0, p + 1)
  B0 <- diag(c(D0, rep(F0, p - 1), A0))
  resDF[[p]] <- regression(deltay, Xy, b0 = b0, B0 = B0, c0 = 2, C0 = 0.001)
}
```

We now compute Chib's estimator, evaluated at the posterior median, exactly
as before.

```{r}
allres <- list(resG, resDF)
logML <- matrix(NA_real_, length(allres), length(resG))
for (i in seq_along(allres)) {
  for (j in seq_len(length(resG))) {
    myres <- allres[[i]][[j]]
    beta_med <- apply(myres$betas, 2, median)
    sigma2_med <- median(myres$sigma2s)
    BN_med <- solve(solve(myres$B0) + crossprod(myres$X) / sigma2_med)
    bN_med <- BN_med %*% (solve(myres$B0) %*% myres$b0 + 
                            crossprod(myres$X, myres$y) / sigma2_med)
    logML[i, j] <-
      sum(dnorm(myres$y, myres$X %*% beta_med, sqrt(sigma2_med), log = TRUE)) +
      dmvnorm(beta_med, myres$b0, myres$B0, log = TRUE) +
      dinvgamma(sigma2_med, myres$c0, myres$C0, log = TRUE) -
      dmvnorm(beta_med, bN_med, BN_med, log = TRUE) -
      logmeanexp(dinvgamma(sigma2_med, myres$paras$cN, myres$paras$CN, log = TRUE))
  }
}
```

Now we can compute and print the table.

```{r}
prob <- logML2post(logML)
tab <- cbind(logML[1, ], prob[1, ], logML[2, ], prob[2, ])
dimnames(tab) <- list("Lag order" = seq_len(ncol(logML)), NULL)
knitr::kable(tab, digits = 2)
```

## Section 11.4.3: Bayesian testing for first-order Markov chain models

### Example 11.19: Wage mobility data: two-group versus homogeneous Markov chain model

We return to the wage mobility data considered in Example 7.16 and
prepare the data analogously.  We load the data and reduce the
observations to workers from the birth cohort 1946-1960.

```{r}
data("labor", package = "BayesianLearningCode")
labor <- subset(labor, birthyear >= 1946 & birthyear <= 1960)
```

We extract the income columns and transform the data to obtain for
each worker the matrix which contains the number of transitions from
one class to the other, i.e., the matrix with values $N_{i,hk}$.

```{r}
income <- labor[, grepl("^income", colnames(labor))]
income <- sapply(income, as.integer)
colnames(income) <- gsub("income_", "", colnames(income))
getTransitions <- function(x, classes) {
    transitions <- matrix(0, length(classes), length(classes))
    for (i in seq_len(length(x) - 1)) {
        transitions[x[i], x[i + 1]] <- transitions[x[i], x[i + 1]] + 1
    }
    dimnames(transitions) <- list(from = classes, to = classes)
    transitions
}
income_transitions <-
    lapply(seq_len(nrow(income)),
           function(i) getTransitions(income[i, ], classes = 0:5))
```

We determine the aggregate transition matrix across all workers as
well as separate transition matrices for male and female workers.

```{r}
N_hk <- Reduce("+", income_transitions)
Ng_hk <- list(male = Reduce("+", income_transitions[!labor$female]),
              female = Reduce("+", income_transitions[labor$female]))
```

We want to compare the following two models: the homogeneous Markov
chain model $\mathcal{M}_H$ and the two-group Markov chain model
$\mathcal{M}_G$ where we assume that the wage mobility of male and
female workers differs in the Austrian labor market.

We determine the marginal likelihoods of the two models, considering
also two different prior choices, namely $\gamma_{0,hk} = 1$ and
$\gamma_{0,hk} = 4$.

```{r}
gammas <- c(1, 4)
K <- nrow(N_hk)
G <- length(Ng_hk)
logmarglikMH <- logmarglikMG <- numeric(2)
for (i in 1:2) {
    gamma <- gammas[i]
    logmarglikMH[i] <- K * (lgamma(K * gamma) - K * lgamma(gamma)) +
        sum(lgamma(gamma + N_hk)) -
        sum(lgamma(rowSums(gamma + N_hk)))
    logmarglikMG[i] <- G * K * (lgamma(K * gamma) - K * lgamma(gamma)) +
        sum(lgamma(gamma + Ng_hk[["male"]])) +
        sum(lgamma(gamma + Ng_hk[["female"]])) -
        sum(lgamma(rowSums(gamma + Ng_hk[["male"]]))) -
        sum(lgamma(rowSums(gamma + Ng_hk[["female"]])))
}
```

We also calculate the log Bayes factor and summarize results.

```{r}
res <- rbind(MG = c(logmarglikMG, NA, NA),
             MH = c(logmarglikMH, logmarglikMH - logmarglikMG))
colnames(res) <- paste("gamma =", rep(gammas, 2))
res
```

There is evidence for the two-group Markov chain model when using
$\gamma_{0,hk} = 1$, whereas the homogeneous model is preferred for
$\gamma_{0,hk} = 4$.

### Example 11.20: Wage mobility data: testing low-income male workers

We impose the restriction $\xi_{g,10} = \xi_{g,12}$ for
group $g$ equal to male workers and determine the log Bayes factor to
the unrestricted two-group model using the two priors for
$\gamma_{0,hk}$.

```{r}
logBF_R1G <- numeric(2)
for (i in 1:2) {
    gamma <- gammas[i]
    logBF_R1G[i] <- lbeta(gamma, gamma) -
        lbeta(gamma + Ng_hk[["male"]]["1", "0"],
              gamma + Ng_hk[["male"]]["1", "2"])
}
```

We obtain the log marginal likelihood of the restricted model by
adding the log Bayes factor to the log marginal likelihood of the
unrestricted two-group model:

```{r}
logmarglikR1 <- logmarglikMG + logBF_R1G
res <- rbind(res,
             MR1 = c(logmarglikR1, logBF_R1G))
res["MR1", ]
```

Results suggest overwhelming evidence for the restricted model,
regardless of the prior on $\gamma_{0,hk}$.

### Example 11.20: Wage mobility data: comparing no-income risk for male and female workers

We continue to compare another restricted model to the two-group
model. In particular, we assume under the restricted model that the
risk of moving to the no-income class is the same for male and female
workers in the lowest income class.

We first calculate the log Bayes factor for this restricted model
compared to the unrestricted two-group model for the two different
priors on $\gamma_{0,hk}$.

```{r}
logBF_R2G <- numeric(2)
for (i in 1:2) {
    gamma <- gammas[i]
    aN1 <- gamma + Ng_hk[["male"]]["1", "0"]
    aN2 <- gamma + Ng_hk[["female"]]["1", "0"]
    bN1 <- sum(gamma + Ng_hk[["male"]]["1", -1])
    bN2 <- sum(gamma + Ng_hk[["female"]]["1", -1])
    logBF_R2G[i] <- lbeta(aN1 + aN2 - 1, bN1 + bN2 - 1) +
        2 * lbeta(gamma, (K - 1) * gamma) -
        lbeta(aN1, bN1) - lbeta(aN2, bN2) -
        lbeta(2 * gamma - 1, 2 * (K - 1) * gamma - 1)
}
```

We obtain the log marginal likelihoods for the restricted model from
the log marginal likelihood of the two-group model and the log Bayes
factor. 

```{r}
logmarglikR2 <- logmarglikMG + logBF_R2G
res <- rbind(res,
             MR2 = c(logmarglikR2, logBF_R2G))
res["MR2", ]
```

Results suggest that the unrestricted model is preferred, regardless
of the prior on $\gamma_{0,hk}$.

We also summarize the results.
```{r}
knitr::kable(res)
```

