---
title: "Chapter 11: Bayesian Model Selection in Regression and Time Series Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Chapter 11: Bayesian Model Selection in Regression and Time Series Analysis}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6
)
knitr::opts_knit$set(global.par = TRUE)
pdfplots <- FALSE # default: FALSE; set this to TRUE only if you like pdf figures
```

# Section 11.4: Model Selection Problems in Time Series Analysis 

## Example 11.18: Application to Austrian wage mobility data - homogeneity versus grouped model

We load the data and reduce the observations to workers from the birth cohort
1946-1960. 

```{r}
data("labor", package = "BayesianLearningCode")
labor <- subset(labor, birthyear >= 1946 & birthyear <= 1960)
```

We extract the income columns and transform the data to obtain for
each worker the matrix which contains the number of transitions from
one class to the other, i.e., the matrix with values $N_{i,hk}$.


```{r}
income <- labor[, grepl("^income", colnames(labor))]
income <- sapply(income, as.integer)
colnames(income) <- gsub("income_", "", colnames(income))
getTransitions <- function(x, classes) {
    transitions <- matrix(0, length(classes), length(classes))
    for (i in seq_len(length(x) - 1)) {
        transitions[x[i], x[i + 1]] <- transitions[x[i], x[i + 1]] + 1
    }
    dimnames(transitions) <- list(from = classes, to = classes)
    transitions
}
income_transitions <-
    lapply(seq_len(nrow(income)),
           function(i) getTransitions(income[i, ], classes = 0:5))
```

We determine the marginal likelihood of the first-order Markov chain
model assuming homogeneity and grouping by gender.

```{r}
N_hk <- Reduce("+", income_transitions)
Ng_hk <- list(male = Reduce("+", income_transitions[!labor$female]),
              female = Reduce("+", income_transitions[labor$female]))
gammas <- c(1, 4)
K <- nrow(income_transitions_all)
G <- length(Ng_hk)
logmarglikMH <- logmarglikMG <- numeric(2)
for (i in 1:2) {
    gamma <- gammas[i]
    logmarglikMH[i] <- K * (lgamma(K * gamma) - K * lgamma(gamma)) +
        sum(lgamma(gamma + N_hk)) -
        sum(lgamma(rowSums(gamma + N_hk)))
    logmarglikMG[i] <- G * K * (lgamma(K * gamma) - K * lgamma(gamma)) +
        sum(lgamma(gamma + Ng_hk[["male"]])) +
        sum(lgamma(gamma + Ng_hk[["female"]])) -
        sum(lgamma(rowSums(gamma + Ng_hk[["male"]]))) -
        sum(lgamma(rowSums(gamma + Ng_hk[["female"]])))
}
```

We also calculate the log BF and summarize results.

```{r}
res <- rbind(MH = c(logmarglikMH, NA, NA),
             MG = c(logmarglikMG, logmarglikMG - logmarglikMH))
colnames(res) <- c("gamma = 1", "gamma = 4",
                   "gamma = 1", "gamma = 4")
knitr::kable(res)
```

## Example 11.19: Application to Austrian wage mobility data - restricted models

We continue to compare restricted models to the grouped model. We
first calculate the log BFs for the restricted models.

```{r}
logBF_R1G <- logBF_R2G <- numeric(2)
for (i in 1:2) {
    gamma <- gammas[i]
    logBF_R1G[i] <- lbeta(gamma, gamma) -
        lbeta(gamma + Ng_hk[["male"]]["1", "0"],
              gamma + Ng_hk[["male"]]["1", "2"])
    aN1 <- gamma + Ng_hk[["male"]]["1", "0"]
    aN2 <- gamma + Ng_hk[["female"]]["1", "0"]
    bN1 <- sum(gamma + Ng_hk[["male"]]["1", -1])
    bN2 <- sum(gamma + Ng_hk[["female"]]["1", -1])
    logBF_R2G[i] <- lbeta(aN1 + aN2 - 1, bN1 + bN2 - 1) +
        2 * lbeta(gamma, (K - 1) * gamma) -
        lbeta(aN1, bN1) - lbeta(aN2, bN2) -
        lbeta(2 * gamma - 1, 2 * (K - 1) * gamma - 1)
}
```

We obtain the log marginal likelihoods for the restricted models from
the log marginal likelihoods of the grouped models and the log BFs. We
then also summarize the results.

```{r}
logmarglikR1 <- logmarglikMG + logBF_R1G
logmarglikR2 <- logmarglikMG + logBF_R2G
res <- rbind(MR1 = c(logmarglikR1, logBF_R1G),
             MR2 = c(logmarglikR2, logBF_R2G))
colnames(res) <- c("gamma = 1", "gamma = 4",
                   "gamma = 1", "gamma = 4")
knitr::kable(res)
```
